---
title: "Correlations tables with COWG's data"
author: "CLETZ Laura"
date: "03/07/2025"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Packages needed:
```{r}
library(readr)
library(mgcv)
library(rlang)
library(ggplot2)
library(Hmisc)
library(GGally)
library(dplyr)
library(FactoMineR)
library(factoextra)
here::i_am("Correlations_and_Predictions/corr_table_COWG.Rmd")
library(here)
```

```{r}
data_small <- read.table(here("processed", "COWG_gamm_2004_2011.csv"), sep=";", dec=".", header=T) %>% 
  mutate_at(18:23, as.factor)
data_small$Dates <- as.Date(data_small$Dates, format = "%Y-%m-%d")
data_small$Corrected_Dates <- as.Date(data_small$Corrected_Dates, format = "%Y-%m-%d")
```

```{r}
data_big <- read.table(here("processed", "COWG_gamm_2004_2018.csv"), sep=";", dec=".", header=T) %>% 
  mutate_at(18:23, as.factor)
data_big$Dates <- as.Date(data_big$Dates, format = "%Y-%m-%d")
data_big$Corrected_Dates <- as.Date(data_big$Corrected_Dates, format = "%Y-%m-%d")
```

```{r}
custom_gam_plot <- function(data, mapping, ...) {
  
  x_var <- rlang::as_label(mapping$x)
  y_var <- rlang::as_label(mapping$y)

  formula <- substitute( y ~ s(x, bs='cr'), list(y = as.name(y_var), x = as.name(x_var)) )
  
  gam_model <- gamm(formula, correlation = corARMA(form = ~1, p=1), data = data, na.action = na.exclude)

  pred_data <- data.frame(x = seq(min(data[[x_var]], na.rm = TRUE),
                                    max(data[[x_var]], na.rm = TRUE),
                                    length.out = 100))
  names(pred_data) <- x_var
  
  pred_data$y <- predict(gam_model$gam, newdata = pred_data)

  ggplot(data, mapping = mapping) +
    geom_point() +
    geom_line(data = pred_data, 
              mapping = aes_string(x = x_var, y = "y"), 
              color = "blue") +
    labs(title = paste0("GAM fit:", y_var, "~ s(", x_var, ")"), 
         x = x_var, y = y_var) +
    theme_minimal()
}
```

```{r, warning=FALSE, echo=FALSE}
g <- ggpairs(
  data = data_small[, c(3:23)],
  lower = list(continuous = custom_gam_plot, combo = 'facetdensity'),
  upper = list(continuous = wrap('cor', size = 3, hjust = 0.8)),
  progress = FALSE
)
```

```{r, warning=FALSE, echo=FALSE}
g
```

Rainfall and Dates:
```{r}
ggplot(data_small, aes(x = Numeric_Dates)) +
  geom_point(aes(y = Rainfall), col = "darkred") +
  geom_smooth(aes(y = Rainfall), method = 'gam', formula = y ~ s(x, bs='re'), col = "red") +
  geom_point(aes(y = Rainfall_lag_1), col = "darkgreen") +
  geom_smooth(aes(y = Rainfall_lag_1), method = 'gam', formula = y ~ s(x, bs='re'), col = "green") +
  theme_minimal()
```

Rainfall at one lag and O2:
```{r}
ggplot(data_small, aes(x = Rainfall_lag_1, y = O2, colour = O2_method)) +
  geom_point() +
  geom_smooth(method = 'gam', formula = y ~ s(x, bs='re')) +
  theme_minimal()
```

Dates and O2:
```{r}
ggplot(data_small, aes(x = Numeric_Dates, y = O2, colour = O2_method)) +
  geom_point() +
  geom_smooth(method = 'gam', formula = y ~ s(x, bs='re')) +
  theme_minimal()
```

Rainfall at one lag and Copper:
```{r}
ggplot(data_small, aes(x = Rainfall_lag_1, y = Cuivre, colour = Cuivre_method)) +
  geom_point() +
  geom_smooth(method = 'gam', formula = y ~ s(x, bs='re')) +
  theme_minimal()
```

FeII and Zinc:
```{r}
ggplot(data_small, aes(x = Ion_Ferreux, y = Zinc, colour = Zinc_method)) +
  geom_point() +
  geom_smooth(method = 'gam', formula = y ~ s(x, bs='re')) +
  theme_minimal()
```

Arsenic and O2:
```{r}
ggplot(data_small, aes(x = Arsenic_Total, y = O2, colour = O2_method)) +
  geom_point() +
  geom_smooth(method = 'gam', formula = y ~ s(x, bs='re')) +
  theme_minimal()
```

Lead and Copper:
```{r}
ggplot(data_small, aes(x = Cuivre, y = Plomb, colour = Plomb_method)) +
  geom_point() +
  geom_smooth(method = 'gam', formula = y ~ s(x, bs='re')) +
  theme_minimal()
```

Measurement methods seem to add some precision at predicting confidence intervals in some case only. There are methods like Colorimétrie for O2 that don't help at all, giving straight lines instead of smooth curves. We did not remove NAs because they are too many to avoid, we need to count them in.

