---
title: "ARIMA and GAMMs predictions at station COWG"
author: "CLETZ Laura"
date: "08/07/2025"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Packages needed:
```{r}
library(dplyr)
library(mgcv)
library(gam)
library(forecast)
here::i_am("Correlations_and_Predictions/COWG_predictions.Rmd")
library(here)
```

```{r}
#data <- read.table(here('processed', 'COWG_gamm_2004_2022.csv'), dec = '.', sep = ';', header = T) %>%
#  mutate_at(1:2, function(x) as.Date(x, format="%Y-%m-%d"))

data <- read.table(here('processed', 'COWG_gamm_2004_2022_twice.csv'), dec = '.', sep = ';', header = T) %>%
  mutate_at(1:2, function(x) as.Date(x, format="%Y-%m-%d"))
```

```{r}
data_small <- read.table(here('processed', 'COWG_gamm_2004_2010.csv'), dec = '.', sep = ';', header = T) %>%
  mutate_at(1:2, function(x) as.Date(x, format="%Y-%m-%d"))
```

```{r}
new_data <- data.frame(
  Dates = as.Date("2025-04-24", format = "%Y-%m-%d"),
  Temperature = 17,
  Rainfall = 107.4,
  Rainfall_lag_1 = 243.1,
  Conductivité = 3250,
  pH = 3.92,
  O2 = 8.62,
  Ion_Ferreux = 801.09,
  Sulfate = (770363.27 * 96) / 32000,
  Zinc = 13294.42,
  Arsenic_Total = 67612.73,
  Plomb = 879.14
  ) %>%
  mutate(Numeric_Dates = as.numeric(Dates)) 
```

```{r}
test <- read.table(here('processed','COWG.csv'), dec = '.', sep = ";", header = TRUE)[76:213,] %>%
  mutate_at(1:2, function(x) as.Date(x, format="%Y-%m-%d")) %>%
  mutate(Numeric_Dates = as.numeric(Corrected_Dates)) %>%
  select(names(new_data))
```

# Temperature
## Whole data

```{r}
temp <- ts(data$Temperature, start = c(2004,11), end = c(2019,1), frequency = 12)

fit_temp <- arima(temp, order = c(1, 1, 1),
      seasonal = list(order = c(1, 1, 1), period = 12), method = "CSS-ML")

pred_temp <- predict(fit_temp, n.ahead = 51) # because difftime("2026-06-01","2022-04-01")/30 = 51
```

```{r}
plot.ts(temp, type='l', xlab='Temps', ylab="Température (C°)",
        xlim = c(2004,2026), xaxt = 'n')

ts <- ts(pred_temp$pred, start = c(2019,1), end = c(2026,6), frequency = 12)
ts_se <- ts(pred_temp$se, start = c(2019,1), end = c(2026,6), frequency = 12)
x_coords <- c(time(ts), rev(time(ts)))
y_coords <- c(ts + ts_se, rev(ts - ts_se))
polygon(x_coords, y_coords, col = 'cyan3', border = "cyan")

points(ts, type='l', col='blue')
points(ts(test$Temperature, start = c(2011,2), end = c(2019,1), frequency = 12),
       pch = 19, col = "darkred")

points(2025.33, new_data$Temperature, col='red', pch=19)

abline(h=10, col='red')
abline(h=18, col='red')

legend(x = "topright",
       legend=c("COWG", "Intervalle de prédiction", "Points non-imputés", "Point test"),
       col=c("black", "cyan3", "darkred", "red"),
       pch=c(NA, NA, 19, 19),
       lty=c(1, 1, NA, NA),
       lwd=c(1, 5, 1, 1),
       cex=c(0.6, 0.6, 0.6, 0.6),
       bty = "n")

axis(1, at = seq(2004, 2026, by = 1), las = 2)
```

## Small data

```{r}
temp <- ts(data_small$Temperature, start = c(2004,11), end = c(2011,1), frequency = 12)

fit_temp <- arima(temp, order = c(1, 1, 1),
      seasonal = list(order = c(1, 1, 1), period = 12), method = "CSS-ML")

pred_temp <- predict(fit_temp, n.ahead = 187) # because difftime("2026-06-01","2011-01-01")/30 = 187
```

```{r}
plot.ts(temp, type='l', xlab='Temps', ylab="Température (C°)",
        xlim = c(2004,2026), xaxt = 'n')

ts <- ts(pred_temp$pred, start = c(2011,1), end = c(2026,6), frequency = 12)
ts_se <- ts(pred_temp$se, start = c(2011,1), end = c(2026,6), frequency = 12)
x_coords <- c(time(ts), rev(time(ts)))
y_coords <- c(ts + ts_se, rev(ts - ts_se))
polygon(x_coords, y_coords, col = 'cyan3', border = "cyan")

points(ts, type='l', col='blue')
points(ts(test$Temperature, start = c(2011,2), end = c(2022,4), frequency = 12),
       pch = 19, col = "darkred")

points(2025.33, new_data$Temperature, col='red', pch=19)

abline(h=13, col='red')
abline(h=18, col='red')

axis(1, at = seq(2004, 2026, by = 1), las = 2)
```

Less precise but still works.

## Mixed Data

```{r}
temp <- ts(data$Temperature, start = c(2004,11), end = c(2011,1), frequency = 12)

fit_temp <- arima(temp, order = c(1, 1, 1),
      seasonal = list(order = c(1, 1, 1), period = 12), method = "CSS-ML")

pred_temp <- predict(fit_temp, n.ahead = 51) # because difftime("2026-06-01","2022-04-01")/30 = 51
```

```{r}
plot.ts(temp, type='l', xlab='Temps', ylab="Température (C°)",
        xlim = c(2004,2026), xaxt = 'n')

ts <- ts(pred_temp$pred, start = c(2011,1), end = c(2026,6), frequency = 12)
ts_se <- ts(pred_temp$se, start = c(2011,1), end = c(2026,6), frequency = 12)
x_coords <- c(time(ts), rev(time(ts)))
y_coords <- c(ts + ts_se, rev(ts - ts_se))
polygon(x_coords, y_coords, col = 'cyan3', border = "cyan")

points(ts, type='l', col='blue')
points(ts(test$Temperature, start = c(2011,2), end = c(2019,1), frequency = 12),
       pch = 19, col = "darkred")

points(2025.33, new_data$Temperature, col='red', pch=19)

abline(h=9, col='red')
abline(h=18, col='red')

legend(x = "topright",
       legend=c("COWG", "Intervalle de prédiction", "Points non-imputés", "Point test"),
       col=c("black", "cyan3", "darkred", "red"),
       pch=c(NA, NA, 19, 19),
       lty=c(1, 1, NA, NA),
       lwd=c(1, 5, 1, 1),
       cex=c(0.6, 0.6, 0.6, 0.6),
       bty = "n")

axis(1, at = seq(2004, 2026, by = 1), las = 2)
```

# O2
## Whole data

```{r}
o2 <- ts(data$O2, start = c(2004,11), end = c(2022,4), frequency = 12)

fit_o2 <- arima(data$O2, order = c(1, 1, 1),
      seasonal = list(order = c(1, 1, 2), period = 12), method = "CSS-ML")

pred_o2 <- predict(fit_o2, n.ahead = 51)
```

```{r}
plot.ts(o2, type='l', xlab='Temps', ylab="O2",
        xlim = c(2004,2026), xaxt = 'n')

ts <- ts(pred_o2$pred, start = c(2022,4), end = c(2026,6), frequency = 12)
ts_se <- ts(pred_o2$se, start = c(2022,4), end = c(2026,6), frequency = 12)
x_coords <- c(time(ts), rev(time(ts)))
y_coords <- c(ts + ts_se, rev(ts - ts_se))
polygon(x_coords, y_coords, col = 'cyan3', border = "cyan")

points(ts, type='l', col='blue')
points(2025.33, new_data$O2, col="red", pch=19)
points(ts(test$O2, start = c(2011,2), end = c(2022,4), frequency = 12),
       pch = 19, col = "darkred")

axis(1, at = seq(2004, 2026, by = 1), las = 2)
```

Near but not clear enough.

## Small data

```{r}
o2 <- ts(data_small$O2, start = c(2004,11), end = c(2011,1), frequency = 12)

fit_o2 <- arima(o2, order = c(1, 1, 1),
      seasonal = list(order = c(1, 1, 1), period = 12), method = "CSS-ML")

pred_o2 <- predict(fit_o2, n.ahead = 187)
```

```{r}
plot.ts(o2, type='l', xlab='Temps', ylab="O2",
        xlim = c(2004,2026), xaxt = 'n')

ts <- ts(pred_o2$pred, start = c(2011,1), end = c(2026,6), frequency = 12)
ts_se <- ts(pred_o2$se, start = c(2011,1), end = c(2026,6), frequency = 12)
x_coords <- c(time(ts), rev(time(ts)))
y_coords <- c(ts + ts_se, rev(ts - ts_se))
polygon(x_coords, y_coords, col = 'cyan3', border = "cyan")

points(ts, type='l', col='blue')
points(ts(test$O2, start = c(2011,2), end = c(2019,1), frequency = 12),
       pch = 19, col = "darkred")

points(2025.33, new_data$O2, col='red', pch=19)

abline(h=13, col='red')
abline(h=18, col='red')

axis(1, at = seq(2004, 2026, by = 1), las = 2)
```

This could work but feels worse.

# pH
## Whole data

```{r}
ph <- ts(data$pH, start = c(2004,11), end = c(2022,4), frequency = 12)

fit_ph <- arima(data$pH, order = c(1, 1, 1),
      seasonal = list(order = c(1, 1, 1), period = 12), method = "CSS-ML")

pred_ph <- predict(fit_ph, n.ahead = 51)
```

```{r}
plot.ts(ph, type='l', xlab='Temps', ylab="pH",
        xlim = c(2004,2026), ylim = c(0,7), xaxt = 'n')

ts <- ts(pred_ph$pred, start = c(2022,4), end = c(2026,6), frequency = 12)
ts_se <- ts(pred_ph$se, start = c(2022,4), end = c(2026,6), frequency = 12)
x_coords <- c(time(ts), rev(time(ts)))
y_coords <- c(ts + ts_se, rev(ts - ts_se))
polygon(x_coords, y_coords, col = 'cyan3', border = "cyan")

points(ts, type='l', col='blue')
points(2025.33, new_data$pH, col="red", pch=19)
points(ts(test$pH, start = c(2011,2), end = c(2022,4), frequency = 12),
       pch = 19, col = "darkred")

axis(1, at = seq(2004, 2026, by = 1), las = 2)
```

Really good prediction!

## Small data

```{r}
ph <- ts(data_small$pH, start = c(2004,11), end = c(2011,1), frequency = 12)

fit_ph <- arima(ph, order = c(1, 1, 1),
      seasonal = list(order = c(1, 1, 1), period = 12), method = "CSS-ML")

pred_ph <- predict(fit_ph, n.ahead = 187)
```

```{r}
plot.ts(ph, type='l', xlab='Temps', ylab="pH",
        xlim = c(2004,2026), xaxt = 'n')

ts <- ts(pred_ph$pred, start = c(2011,1), end = c(2026,6), frequency = 12)
ts_se <- ts(pred_ph$se, start = c(2011,1), end = c(2026,6), frequency = 12)
x_coords <- c(time(ts), rev(time(ts)))
y_coords <- c(ts + ts_se, rev(ts - ts_se))
polygon(x_coords, y_coords, col = 'cyan3', border = "cyan")

points(ts, type='l', col='blue')
points(ts(test$pH, start = c(2011,2), end = c(2019,1), frequency = 12),
       pch = 19, col = "darkred")

points(2025.33, new_data$pH, col='red', pch=19)

abline(h=13, col='red')
abline(h=18, col='red')

axis(1, at = seq(2004, 2026, by = 1), las = 2)
```

The first prediction is better.

# Arsenic_Total
## Whole data

```{r}
ars <- ts(data$Arsenic_Total, start = c(2004,11), end = c(2022,4), frequency = 12)

fit_ars <- arima(data$Arsenic_Total, order = c(1, 1, 1),
      seasonal = list(order = c(1, 1, 1), period = 12), method = "CSS-ML")

pred_ars <- predict(fit_ars, n.ahead = 51)
```

```{r}
plot.ts(ars, type='l', xlab='Temps', ylab="Concentration en Arsenic (ppb)",
        xlim = c(2004,2026), xaxt = 'n')

ts <- ts(pred_ars$pred, start = c(2022,4), end = c(2026,6), frequency = 12)
ts_se <- ts(pred_ars$se, start = c(2022,4), end = c(2026,6), frequency = 12)
x_coords <- c(time(ts), rev(time(ts)))
y_coords <- c(ts + ts_se, rev(ts - ts_se))
polygon(x_coords, y_coords, col = 'cyan3', border = "cyan")

points(ts, type='l', col='blue')
points(2025.33, new_data$Arsenic_Total, col="red", pch=19)
points(ts(test$Arsenic_Total, start = c(2011,2), end = c(2022,4), frequency = 12),
       pch = 19, col = "darkred")

legend(x = "topright",
       legend=c("S1", "Intervalle de prédiction", "Points non-imputés", "Points tests"),
       col=c("black", "cyan3", "darkred", "red"),
       pch=c(NA, NA, 19, 19),
       lty=c(1, 1, NA, NA),
       lwd=c(1, 5, 1, 1),
       cex=c(0.6, 0.6, 0.6, 0.6),
       bty = "n")

abline(h=150000, col='red')
abline(h=50000, col='red')

axis(1, at = seq(2004, 2026, by = 1), las = 2)
```

It might be a good prediction.

## Small data

```{r}
ars <- ts(data_small$Arsenic_Total, start = c(2004,11), end = c(2011,1), frequency = 12)

fit_ars <- arima(ars, order = c(1, 1, 1),
      seasonal = list(order = c(1, 1, 1), period = 12), method = "CSS-ML")

pred_ars <- predict(fit_ars, n.ahead = 187)
```

```{r}
plot.ts(ars, type='l', xlab='Temps', ylab="Concentration en arsenic (ppb)",
        xlim = c(2004,2026), xaxt = 'n')

ts <- ts(pred_ars$pred, start = c(2011,1), end = c(2026,6), frequency = 12)
ts_se <- ts(pred_ars$se, start = c(2011,1), end = c(2026,6), frequency = 12)
x_coords <- c(time(ts), rev(time(ts)))
y_coords <- c(ts + ts_se, rev(ts - ts_se))
polygon(x_coords, y_coords, col = 'cyan3', border = "cyan")

points(ts, type='l', col='blue')
points(ts(test$Arsenic_Total, start = c(2011,2), end = c(2022,4), frequency = 12),
       pch = 19, col = "darkred")

points(2025.33, new_data$Arsenic_Total, col='red', pch=19)

abline(h=13, col='red')
abline(h=18, col='red')

axis(1, at = seq(2004, 2026, by = 1), las = 2)
```

The prediction interval is too wide.

```{r}
ars <- ts(data$Arsenic_Total, start = c(2004,11), end = c(2011,1), frequency = 12)

fit_ars <- arima(ars, order = c(1, 1, 1),
      seasonal = list(order = c(1, 1, 1), period = 12), method = "CSS-ML")

pred_ars <- predict(fit_ars, n.ahead = 51) 
```

```{r}
plot.ts(ars, type='l', xlab='Temps', ylab="Concentration en arsenic (ppb)",
        xlim = c(2004,2026), xaxt = 'n')

ts <- ts(pred_ars$pred, start = c(2011,1), end = c(2026,6), frequency = 12)
ts_se <- ts(pred_ars$se, start = c(2011,1), end = c(2026,6), frequency = 12)
x_coords <- c(time(ts), rev(time(ts)))
y_coords <- c(ts + ts_se, rev(ts - ts_se))
polygon(x_coords, y_coords, col = 'cyan3', border = "cyan")

points(ts, type='l', col='blue')
points(ts(test$Arsenic_Total, start = c(2011,2), end = c(2019,1), frequency = 12),
       pch = 19, col = "darkred")

points(2025.33, new_data$Arsenic_Total, col='red', pch=19)

abline(h=9, col='red')
abline(h=18, col='red')

legend(x = "topleft",
       legend=c("COWG", "Intervalle de prédiction", "Points non-imputés", "Point test"),
       col=c("black", "cyan3", "darkred", "red"),
       pch=c(NA, NA, 19, 19),
       lty=c(1, 1, NA, NA),
       lwd=c(1, 5, 1, 1),
       cex=c(0.6, 0.6, 0.6, 0.6),
       bty = "n")

axis(1, at = seq(2004, 2026, by = 1), las = 2)
```

## GAMM and whole data

```{r}
ars <- ts(data$Arsenic_Total, start = c(2004,11), end = c(2022,4), frequency = 12)

plot.ts(ars, type='l', xlab='Temps', ylab="Concentration en Arsenic (ppb)",
        xlim = c(2004,2026), xaxt = 'n')

gam_ars <- gamm(Arsenic_Total ~ s(Rainfall_lag_1, bs='re') + s(O2, bs='re') + s(Sulfate, bs='re') + Ion_Ferreux + 
                 Numeric_Dates + Plomb + Rainfall_lag_1:O2 + Conductivité^2 + Zinc + 
                 Ion_Ferreux:Plomb + O2:Ion_Ferreux + Zinc:pH + Zinc:Sulfate + Sulfate:Ion_Ferreux, data = data, 
               correlation = corARMA(form = ~1, p=1))

pred_ars <- predict.gam(gam_ars$gam, newdata = test, se.fit = TRUE)
points(ts(pred_ars$fit, start = c(2011,2), end = c(2022,4), frequency = 12),
       pch = 15, cex = 0.5, col = "cyan3")
points(spline(ts(pred_ars$fit, start = c(2011,2), end = c(2022,4), frequency = 12)),
       type='l', col = "cyan3")
points(ts(test$Arsenic_Total, start = c(2011,2), end = c(2022,4), frequency = 12),
       pch = 19, col = "darkred")

upper <- spline(ts(pred_ars$fit + pred_ars$se.fit, start = c(2011,2), end = c(2022,4), frequency = 12))
lower <- spline(ts(pred_ars$fit - pred_ars$se.fit, start = c(2011,2), end = c(2022,4), frequency = 12))
points(upper, type='l', col = "cyan4")
points(lower, type='l', col = "cyan4")

pred_ars <- predict.gam(gam_ars$gam, newdata = new_data, se.fit = TRUE)
points(2025.33, pred_ars$fit, pch=19, col='cyan3')
points(2025.33, pred_ars$fit + pred_ars$se.fit, pch=15, col='cyan4')
points(2025.33, pred_ars$fit - pred_ars$se.fit, pch=15, col='cyan4')

points(2025.33, new_data$Arsenic_Total, col="red", pch=19)

axis(1, at = seq(2004, 2026, by = 1), las = 2)
```

Still not conclusive...

## GAMM and small data

```{r}
ars <- ts(data_small$Arsenic_Total, start = c(2004,11), end = c(2011,1), frequency = 12)

plot.ts(ars, type='l', xlab='Temps', ylab="Concentration en Arsenic",
        xlim = c(2004,2026), xaxt = 'n')

gam_ars <- gamm(Arsenic_Total ~ s(Rainfall_lag_1, bs='re') + s(O2, bs='re') + s(Sulfate, bs='re') + Ion_Ferreux + 
                 Numeric_Dates + Plomb + Rainfall_lag_1:O2 + Conductivité^2 + Zinc + 
                 Ion_Ferreux:Plomb + O2:Ion_Ferreux + Zinc:pH + Zinc:Sulfate + Sulfate:Ion_Ferreux, 
                data = data_small, correlation = corARMA(form = ~1, p=1))

pred_ars <- predict.gam(gam_ars$gam, newdata = test, se.fit = TRUE)
points(ts(pred_ars$fit, start = c(2011,2), end = c(2019,1), frequency = 12),
       pch = 15, col = "cyan3")
points(spline(ts(pred_ars$fit, start = c(2011,2), end = c(2019,1), frequency = 12)),
       type='l', col = "cyan3")
points(ts(test$Arsenic_Total, start = c(2011,2), end = c(2019,1), frequency = 12),
       pch = 19, col = "darkred")

upper <- spline(ts(pred_ars$fit + pred_ars$se.fit, start = c(2011,2), end = c(2019,1), frequency = 12))
lower <- spline(ts(pred_ars$fit - pred_ars$se.fit, start = c(2011,2), end = c(2019,1), frequency = 12))
points(upper, type='l', col = "cyan4")
points(lower, type='l', col = "cyan4")

pred_ars <- predict.gam(gam_ars$gam, newdata = new_data, se.fit = TRUE)
points(2025.33, pred_ars$fit, pch=19, col='cyan3')
points(2025.33, pred_ars$fit + pred_ars$se.fit, pch=15, col='cyan4')
points(2025.33, pred_ars$fit - pred_ars$se.fit, pch=15, col='cyan4')

points(2025.33, new_data$Arsenic_Total, col="red", pch=19)

axis(1, at = seq(2004, 2026, by = 1), las = 2)
```

The prediction is mostly accurate but it's hard to determine the prediction interval precisely this way.

# Zinc
## Whole data

```{r}
zinc <- ts(data$Zinc, start = c(2004,11), end = c(2022,4), frequency = 12)

fit_zinc <- arima(data$Zinc, order = c(1, 1, 1),
      seasonal = list(order = c(1, 1, 1), period = 12), method = "CSS-ML")

pred_zinc <- predict(fit_zinc, n.ahead = 51)
```

```{r}
plot.ts(zinc, type='l', xlab='Temps', ylab="Concentration en Zinc (ppb)",
        xlim = c(2004,2026), xaxt = 'n')

ts <- ts(pred_zinc$pred, start = c(2022,4), end = c(2026,6), frequency = 12)
ts_se <- ts(pred_zinc$se, start = c(2022,4), end = c(2026,6), frequency = 12)
x_coords <- c(time(ts), rev(time(ts)))
y_coords <- c(ts + ts_se, rev(ts - ts_se))
polygon(x_coords, y_coords, col = 'cyan3', border = "cyan")

points(ts, type='l', col='blue')
points(2025.33, new_data$Zinc, col="red", pch=19)
points(ts(test$Zinc, start = c(2011,2), end = c(2022,4), frequency = 12),
       pch = 19, col = "darkred")

axis(1, at = seq(2004, 2026, by = 1), las = 2)
```

Good prediction.

## Small data

```{r}
zinc <- ts(data_small$Zinc, start = c(2004,11), end = c(2011,1), frequency = 12)

fit_zinc <- arima(zinc, order = c(1, 1, 1),
      seasonal = list(order = c(1, 1, 1), period = 12), method = "CSS-ML")

pred_zinc <- predict(fit_zinc, n.ahead = 187)
```

```{r}
plot.ts(zinc, type='l', xlab='Temps', ylab="Concentration en Zinc (ppb)",
        xlim = c(2004,2026), xaxt = 'n')

ts <- ts(pred_zinc$pred, start = c(2011,1), end = c(2026,6), frequency = 12)
ts_se <- ts(pred_zinc$se, start = c(2011,1), end = c(2026,6), frequency = 12)
x_coords <- c(time(ts), rev(time(ts)))
y_coords <- c(ts + ts_se, rev(ts - ts_se))
polygon(x_coords, y_coords, col = 'cyan3', border = "cyan")

points(ts, type='l', col='blue')
points(ts(test$Zinc, start = c(2011,2), end = c(2022,4), frequency = 12),
       pch = 19, col = "darkred")

points(2025.33, new_data$Zinc, col='red', pch=19)

abline(h=13, col='red')
abline(h=18, col='red')

axis(1, at = seq(2004, 2026, by = 1), las = 2)
```

The prediction interval is too wide.

## GAMM and whole data

```{r}
zinc <- ts(data$Zinc, start = c(2004,11), end = c(2022,4), frequency = 12)

plot.ts(zinc, type='l', xlab='Temps', ylab="Concentration en Zinc (ppb)",
        xlim = c(2004,2026), xaxt = 'n')

gam_zinc <- gamm(Zinc ~ s(Sulfate, bs='re') + Ion_Ferreux + Plomb + Conductivité^2 + Arsenic_Total +
               Sulfate:Ion_Ferreux + Sulfate:Plomb + Sulfate:Conductivité, data = data, 
               correlation = corARMA(form = ~1, p=1))

pred_zinc <- predict.gam(gam_zinc$gam, newdata = test, se.fit = TRUE)
points(ts(pred_zinc$fit, start = c(2011,2), end = c(2022,4), frequency = 12),
       pch = 15, cex = 0.5, col = "cyan3")
points(spline(ts(pred_zinc$fit, start = c(2011,2), end = c(2022,4), frequency = 12)),
       type='l', col = "cyan3")
points(ts(test$Zinc, start = c(2011,2), end = c(2022,4), frequency = 12),
       pch = 19, col = "darkred")

upper <- spline(ts(pred_zinc$fit + pred_zinc$se.fit, start = c(2011,2), end = c(2022,4), frequency = 12))
lower <- spline(ts(pred_zinc$fit - pred_zinc$se.fit, start = c(2011,2), end = c(2022,4), frequency = 12))
points(upper, type='l', col = "cyan4")
points(lower, type='l', col = "cyan4")

pred_zinc <- predict.gam(gam_zinc$gam, newdata = new_data, se.fit = TRUE)
points(2025.33, pred_zinc$fit, pch=19, col='cyan3')
points(2025.33, pred_zinc$fit + pred_zinc$se.fit, pch=15, col='cyan4')
points(2025.33, pred_zinc$fit - pred_zinc$se.fit, pch=15, col='cyan4')

points(2025.33, new_data$Zinc, col="red", pch=19)

axis(1, at = seq(2004, 2026, by = 1), las = 2)
```

Still not conclusive...

## GAMM and small data

```{r}
zinc <- ts(data_small$Zinc, start = c(2004,11), end = c(2011,1), frequency = 12)

plot.ts(zinc, type='l', xlab='Temps', ylab="Concentration en Zinc",
        xlim = c(2004,2026), xaxt = 'n')

gam_zinc <- gamm(Zinc ~ s(Sulfate, bs='re') + Ion_Ferreux + Plomb + Conductivité^2 + Arsenic_Total +
               Sulfate:Ion_Ferreux + Sulfate:Plomb + Sulfate:Conductivité, 
                data = data_small, correlation = corARMA(form = ~1, p=1))

pred_zinc <- predict.gam(gam_zinc$gam, newdata = test, se.fit = TRUE)
points(ts(pred_zinc$fit, start = c(2011,2), end = c(2019,1), frequency = 12),
       pch = 15, col = "cyan3")
points(spline(ts(pred_zinc$fit, start = c(2011,2), end = c(2019,1), frequency = 12)),
       type='l', col = "cyan3")
points(ts(test$Zinc, start = c(2011,2), end = c(2019,1), frequency = 12),
       pch = 19, col = "darkred")

upper <- spline(ts(pred_zinc$fit + pred_zinc$se.fit, start = c(2011,2), end = c(2019,1), frequency = 12))
lower <- spline(ts(pred_zinc$fit - pred_zinc$se.fit, start = c(2011,2), end = c(2019,1), frequency = 12))
points(upper, type='l', col = "cyan4")
points(lower, type='l', col = "cyan4")

pred_zinc <- predict.gam(gam_zinc$gam, newdata = new_data, se.fit = TRUE)
points(2025.33, pred_zinc$fit, pch=19, col='cyan3')
points(2025.33, pred_zinc$fit + pred_zinc$se.fit, pch=15, col='cyan4')
points(2025.33, pred_zinc$fit - pred_zinc$se.fit, pch=15, col='cyan4')

points(2025.33, new_data$Zinc, col="red", pch=19)

axis(1, at = seq(2004, 2026, by = 1), las = 2)
```

Almost!

# Plomb
## Whole data

```{r}
lead <- ts(data$Plomb, start = c(2004,11), end = c(2022,4), frequency = 12)

fit_lead <- arima(data$Plomb, order = c(1, 1, 1),
      seasonal = list(order = c(1, 1, 1), period = 12), method = "CSS-ML")

pred_lead <- predict(fit_lead, n.ahead = 51)
```

```{r}
plot.ts(lead, type='l', xlab='Temps', ylab="Concentration en Plomb (ppb)",
        xlim = c(2004,2026), xaxt = 'n')

ts <- ts(pred_lead$pred, start = c(2022,4), end = c(2026,6), frequency = 12)
ts_se <- ts(pred_lead$se, start = c(2022,4), end = c(2026,6), frequency = 12)
x_coords <- c(time(ts), rev(time(ts)))
y_coords <- c(ts + ts_se, rev(ts - ts_se))
polygon(x_coords, y_coords, col = 'cyan3', border = "cyan")

points(ts, type='l', col='blue')
points(2025.33, new_data$Plomb, col="red", pch=19)
points(ts(test$Plomb, start = c(2011,2), end = c(2022,4), frequency = 12),
       pch = 19, col = "darkred")

axis(1, at = seq(2004, 2026, by = 1), las = 2)
```

Really good prediction!

## Small data

```{r}
lead <- ts(data_small$Plomb, start = c(2004,11), end = c(2011,1), frequency = 12)

fit_lead <- arima(lead, order = c(1, 1, 1),
      seasonal = list(order = c(1, 1, 1), period = 12), method = "CSS-ML")

pred_lead <- predict(fit_lead, n.ahead = 187)
```

```{r}
plot.ts(lead, type='l', xlab='Temps', ylab="Concentration en Plomb (ppb)",
        xlim = c(2004,2026), xaxt = 'n')

ts <- ts(pred_lead$pred, start = c(2011,1), end = c(2026,6), frequency = 12)
ts_se <- ts(pred_lead$se, start = c(2011,1), end = c(2026,6), frequency = 12)
x_coords <- c(time(ts), rev(time(ts)))
y_coords <- c(ts + ts_se, rev(ts - ts_se))
polygon(x_coords, y_coords, col = 'cyan3', border = "cyan")

points(ts, type='l', col='blue')
points(ts(test$Plomb, start = c(2011,2), end = c(2022,4), frequency = 12),
       pch = 19, col = "darkred")

points(2025.33, new_data$Plomb, col='red', pch=19)

abline(h=13, col='red')
abline(h=18, col='red')

axis(1, at = seq(2004, 2026, by = 1), las = 2)
```

Nope.

## GAMM and whole data

```{r}
lead <- ts(data$Plomb, start = c(2004,11), end = c(2022,4), frequency = 12)

plot.ts(lead, type='l', xlab='Temps', ylab="Concentration en Plomb (ppb)",
        xlim = c(2004,2026), xaxt = 'n')

gam_lead <- gamm(Plomb ~ s(Rainfall_lag_1, bs='re') + s(O2, bs='re') + Zinc + Ion_Ferreux + pH + 
                   Conductivité + Arsenic_Total + Conductivité:Arsenic_Total + Rainfall_lag_1:Conductivité +
                   Arsenic_Total:Ion_Ferreux, data = data, 
               correlation = corARMA(form = ~1, p=1))

pred_lead <- predict.gam(gam_lead$gam, newdata = test, se.fit = TRUE)
points(ts(pred_lead$fit, start = c(2011,2), end = c(2022,4), frequency = 12),
       pch = 15, cex = 0.5, col = "cyan3")
points(spline(ts(pred_lead$fit, start = c(2011,2), end = c(2022,4), frequency = 12)),
       type='l', col = "cyan3")
points(ts(test$Plomb, start = c(2011,2), end = c(2022,4), frequency = 12),
       pch = 19, col = "darkred")

upper <- spline(ts(pred_lead$fit + pred_lead$se.fit, start = c(2011,2), end = c(2022,4), frequency = 12))
lower <- spline(ts(pred_lead$fit - pred_lead$se.fit, start = c(2011,2), end = c(2022,4), frequency = 12))
points(upper, type='l', col = "cyan4")
points(lower, type='l', col = "cyan4")

pred_lead <- predict.gam(gam_lead$gam, newdata = new_data, se.fit = TRUE)
points(2025.33, pred_lead$fit, pch=19, col='cyan3')
points(2025.33, pred_lead$fit + pred_lead$se.fit, pch=15, col='cyan4')
points(2025.33, pred_lead$fit - pred_lead$se.fit, pch=15, col='cyan4')

points(2025.33, new_data$Plomb, col="red", pch=19)

axis(1, at = seq(2004, 2026, by = 1), las = 2)
```

Nope.

## GAMM and small data

```{r}
lead <- ts(data_small$Plomb, start = c(2004,11), end = c(2011,1), frequency = 12)

plot.ts(lead, type='l', xlab='Temps', ylab="Concentration en Plomb",
        xlim = c(2004,2026), xaxt = 'n')

gam_lead <- gamm(Plomb ~ s(Rainfall_lag_1, bs='re') + s(O2, bs='re') + Zinc + Ion_Ferreux + pH + 
                   Conductivité + Arsenic_Total + Conductivité:Arsenic_Total + Rainfall_lag_1:Conductivité +
                   Arsenic_Total:Ion_Ferreux, 
                data = data_small, correlation = corARMA(form = ~1, p=1))

pred_lead <- predict.gam(gam_lead$gam, newdata = test, se.fit = TRUE)
points(ts(pred_lead$fit, start = c(2011,2), end = c(2019,1), frequency = 12),
       pch = 15, col = "cyan3")
points(spline(ts(pred_lead$fit, start = c(2011,2), end = c(2019,1), frequency = 12)),
       type='l', col = "cyan3")
points(ts(test$Plomb, start = c(2011,2), end = c(2019,1), frequency = 12),
       pch = 19, col = "darkred")

upper <- spline(ts(pred_lead$fit + pred_lead$se.fit, start = c(2011,2), end = c(2019,1), frequency = 12))
lower <- spline(ts(pred_lead$fit - pred_lead$se.fit, start = c(2011,2), end = c(2019,1), frequency = 12))
points(upper, type='l', col = "cyan4")
points(lower, type='l', col = "cyan4")

pred_lead <- predict.gam(gam_lead$gam, newdata = new_data, se.fit = TRUE)
points(2025.33, pred_lead$fit, pch=19, col='cyan3')
points(2025.33, pred_lead$fit + pred_lead$se.fit, pch=15, col='cyan4')
points(2025.33, pred_lead$fit - pred_lead$se.fit, pch=15, col='cyan4')

points(2025.33, new_data$Plomb, col="red", pch=19)

axis(1, at = seq(2004, 2026, by = 1), las = 2)
```

Maybe.

# Ion_Ferreux
## Whole data

```{r}
fer <- ts(data$Ion_Ferreux, start = c(2004,11), end = c(2022,4), frequency = 12)

fit_fer <- arima(data$Ion_Ferreux, order = c(1, 1, 1),
      seasonal = list(order = c(1, 1, 1), period = 12), method = "CSS-ML")

pred_fer <- predict(fit_fer, n.ahead = 51)
```

```{r}
plot.ts(fer, type='l', xlab='Temps', ylab="Concentration en Ion_Ferreux (ppb)",
        xlim = c(2004,2026), xaxt = 'n')

ts <- ts(pred_fer$pred, start = c(2022,4), end = c(2026,6), frequency = 12)
ts_se <- ts(pred_fer$se, start = c(2022,4), end = c(2026,6), frequency = 12)
x_coords <- c(time(ts), rev(time(ts)))
y_coords <- c(ts + ts_se, rev(ts - ts_se))
polygon(x_coords, y_coords, col = 'cyan3', border = "cyan")

points(ts, type='l', col='blue')
points(2025.33, new_data$Ion_Ferreux, col="red", pch=19)
points(ts(test$Ion_Ferreux, start = c(2011,2), end = c(2022,4), frequency = 12),
       pch = 19, col = "darkred")

axis(1, at = seq(2004, 2026, by = 1), las = 2)
```

Really good prediction!

## Small data

```{r}
fer <- ts(data_small$Ion_Ferreux, start = c(2004,11), end = c(2011,1), frequency = 12)

fit_fer <- arima(fer, order = c(1, 1, 1),
      seasonal = list(order = c(1, 1, 1), period = 12), method = "CSS-ML")

pred_fer <- predict(fit_fer, n.ahead = 187)
```

```{r}
plot.ts(fer, type='l', xlab='Temps', ylab="Concentration en Ion_Ferreux (ppb)",
        xlim = c(2004,2026), xaxt = 'n')

ts <- ts(pred_fer$pred, start = c(2011,1), end = c(2026,6), frequency = 12)
ts_se <- ts(pred_fer$se, start = c(2011,1), end = c(2026,6), frequency = 12)
x_coords <- c(time(ts), rev(time(ts)))
y_coords <- c(ts + ts_se, rev(ts - ts_se))
polygon(x_coords, y_coords, col = 'cyan3', border = "cyan")

points(ts, type='l', col='blue')
points(ts(test$Ion_Ferreux, start = c(2011,2), end = c(2022,4), frequency = 12),
       pch = 19, col = "darkred")

points(2025.33, new_data$Ion_Ferreux, col='red', pch=19)

abline(h=13, col='red')
abline(h=18, col='red')

axis(1, at = seq(2004, 2026, by = 1), las = 2)
```

Nope.

## GAMM and whole data

```{r}
fer <- ts(data$Ion_Ferreux, start = c(2004,11), end = c(2022,4), frequency = 12)

plot.ts(fer, type='l', xlab='Temps', ylab="Concentration en Ion_Ferreux (ppb)",
        xlim = c(2004,2026), xaxt = 'n')

gam_fer <- gamm(Ion_Ferreux ~ s(Rainfall, bs='re') + s(O2, bs='re') + Plomb + Conductivité + 
               sqrt(Arsenic_Total) + Plomb:Conductivité + Rainfall:O2, data = data, 
               correlation = corARMA(form = ~1, p=1))

pred_fer <- predict.gam(gam_fer$gam, newdata = test, se.fit = TRUE)
points(ts(pred_fer$fit, start = c(2011,2), end = c(2022,4), frequency = 12),
       pch = 15, cex = 0.5, col = "cyan3")
points(spline(ts(pred_fer$fit, start = c(2011,2), end = c(2022,4), frequency = 12)),
       type='l', col = "cyan3")
points(ts(test$Ion_Ferreux, start = c(2011,2), end = c(2022,4), frequency = 12),
       pch = 19, col = "darkred")

upper <- spline(ts(pred_fer$fit + pred_fer$se.fit, start = c(2011,2), end = c(2022,4), frequency = 12))
lower <- spline(ts(pred_fer$fit - pred_fer$se.fit, start = c(2011,2), end = c(2022,4), frequency = 12))
points(upper, type='l', col = "cyan4")
points(lower, type='l', col = "cyan4")

pred_fer <- predict.gam(gam_fer$gam, newdata = new_data, se.fit = TRUE)
points(2025.33, pred_fer$fit, pch=19, col='cyan3')
points(2025.33, pred_fer$fit + pred_fer$se.fit, pch=15, col='cyan4')
points(2025.33, pred_fer$fit - pred_fer$se.fit, pch=15, col='cyan4')

points(2025.33, new_data$Ion_Ferreux, col="red", pch=19)

axis(1, at = seq(2004, 2026, by = 1), las = 2)
```

Nope.

## GAMM and small data

```{r}
fer <- ts(data_small$Ion_Ferreux, start = c(2004,11), end = c(2011,1), frequency = 12)

plot.ts(fer, type='l', xlab='Temps', ylab="Concentration en Ion_Ferreux",
        xlim = c(2004,2026), xaxt = 'n')

gam_fer <- gamm(Ion_Ferreux ~ s(Rainfall, bs='re') + s(O2, bs='re') + Plomb + Conductivité + 
               sqrt(Arsenic_Total) + Plomb:Conductivité + Rainfall:O2, 
                data = data_small, correlation = corARMA(form = ~1, p=1))

pred_fer <- predict.gam(gam_fer$gam, newdata = test, se.fit = TRUE)
points(ts(pred_fer$fit, start = c(2011,2), end = c(2019,1), frequency = 12),
       pch = 15, col = "cyan3")
points(spline(ts(pred_fer$fit, start = c(2011,2), end = c(2019,1), frequency = 12)),
       type='l', col = "cyan3")
points(ts(test$Ion_Ferreux, start = c(2011,2), end = c(2019,1), frequency = 12),
       pch = 19, col = "darkred")

upper <- spline(ts(pred_fer$fit + pred_fer$se.fit, start = c(2011,2), end = c(2019,1), frequency = 12))
lower <- spline(ts(pred_fer$fit - pred_fer$se.fit, start = c(2011,2), end = c(2019,1), frequency = 12))
points(upper, type='l', col = "cyan4")
points(lower, type='l', col = "cyan4")

pred_fer <- predict.gam(gam_fer$gam, newdata = new_data, se.fit = TRUE)
points(2025.33, pred_fer$fit, pch=19, col='cyan3')
points(2025.33, pred_fer$fit + pred_fer$se.fit, pch=15, col='cyan4')
points(2025.33, pred_fer$fit - pred_fer$se.fit, pch=15, col='cyan4')

points(2025.33, new_data$Ion_Ferreux, col="red", pch=19)

axis(1, at = seq(2004, 2026, by = 1), las = 2)
```

Maybe.

