---
title: "MFA with COWG's data"
author: "CLETZ Laura"
date: "04/07/2025"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Packages needed:
```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggpubr)
library(stringr)
library(lubridate)
library(FactoMineR)
library(factoextra)
here::i_am("Correlations_and_Predictions/MFA_S1.Rmd")
library(here)
```

```{r}
data_small <- read.table(here("processed","COWG_gamm_2004_2010.csv"), sep=";", dec=".", header=T)[c(1,3:4,6:9,12,15:18)] %>%
  mutate_at(1, function(x) as.Date(x, format="%Y-%m-%d"))

data_big <- read.table(here("processed","COWG_gamm_2004_2022.csv"), sep=";", dec=".", header=T)[c(1,3:4,6:9,12,15:18)] %>%
  mutate_at(1, function(x) as.Date(x, format="%Y-%m-%d"))

data_diff <- data_big %>% anti_join(data_small)
```

# From 2004 to 2010

```{r}
res.mfa <- MFA(data_small[-1], group=c(6,5), graph=F)
barplot(res.mfa$eig[,2])
```

The first two components seem the most importants, but the third and the forth components have near or above 10% of the total of inertia.

```{r}
res.mfa$quanti.var$contrib
```
Conductivité, Arsenic_Total, Zinc and Ion_Ferreux contribute the most to the first component.
Rainfall, pH, O2 and Plomb contribute the most to the second one.
Rainfall_lag_1, Temperature and **pH** contribute the most to the third one.
Rainfall and **Temperature** contribute the most to the forth one, that means the forth and the second components are somewhat alike.
Rainfall, **Temperature**, **pH**, **O2** contribute the most to the fifth component that we will not consider.

Most of the time, physicochemical elements are the most contributing to any axis. It also seems that when metalloids contribute, they're strongly correlated to dilution and thus to physicochemical measures.

The contributions on the first two axis are the same as in `MFA_S1.Rmd`.

```{r}
gr_names <- rownames(res.mfa$partial.axes$coord)
gr_names <- ifelse(gr_names %in% gr_names[c(1:5)],
                        str_replace(gr_names, "Gr1", "Élém. ph-chimiques + temps"),
                        str_replace(gr_names, "Gr2", "Élém. traces"))
rownames(res.mfa$partial.axes$coord) <- gr_names

fviz_mfa_axes(res.mfa,
            repel = TRUE, 
            ggtheme = theme_minimal(),
            title = "MFA Partial Axis Plot")
```

The first dimension is well balanced when it comes to groups contribution but their positions are perpendicular, meaning they don't fit the first dimension of a regular PCA.
The second dimension is well balanced between groups and seem to fit the second dimension.
The third to fifth dimension are negligible.

Let's analyse axis 1 and 2 remembering their contributors.

```{r}
var_names <- rownames(res.mfa$quanti.var$coord)
custom_groups <- ifelse(var_names %in% var_names[c(1:6)],
                        "Élém. ph-chimiques + temps",
                        "Élém. traces")

a <- fviz_mfa_var(res.mfa, 
             axes = c(1,2), 
             repel = TRUE, 
             col.var = custom_groups,  
             ggtheme = theme_minimal(),
             title = "") +
  labs(color='')  +
  theme(legend.position = "bottom")

b <- fviz_mfa_ind(res.mfa,
            axes = c(1,2),
            repel = TRUE,
            habillage = factor(month(data_small$Corrected_Dates)),
            ggtheme = theme_minimal(),
            title = "") +
  labs(color='Months')

ggarrange(a, b, labels = c("MFA Variables Plot", "MFA Individuals Plot"))
```

```{r}
library(plotly)
ggplotly(b) %>% 
  layout(hovermode = "x unified")
```

# From 2011 to 2022

```{r}
res.mfa <- MFA(data_diff[-1], group=c(6,5), graph=F)
barplot(res.mfa$eig[,2])
```
The first component seems the most important, but the components 2, 3 and 4 might not be negligible.

```{r}
res.mfa$quanti.var$contrib
```
**Rainfall**, **O2**, Conductivité, Arsenic_Total, **Sulfate** and Ion_Ferreux contribute the most to the first component.
**pH** and **Plomb** contribute the most to the second one.
Rainfall, Rainfall_lag_1 and **Conductivité** contribute the most to the third one.
Temperature contributes the most to the forth one.
Rainfall**_lag_1**, O2 and **Zinc** contribute the most to the fifth one.

The variables have been distributed differently from station S1.

```{r}
gr_names <- rownames(res.mfa$partial.axes$coord)
gr_names <- ifelse(gr_names %in% gr_names[c(1:5)],
                        str_replace(gr_names, "Gr1", "Élém. ph-chimiques + temps"),
                        str_replace(gr_names, "Gr2", "Élém. traces"))
rownames(res.mfa$partial.axes$coord) <- gr_names

fviz_mfa_axes(res.mfa,
            repel = TRUE, 
            ggtheme = theme_minimal(),
            title = "MFA Partial Axis Plot")
```

The dimension 1 is not fitted by any group, the dimension 1 for each PCA contributes oppositely.
That goes the same way with the dimension 2.
The other dimensions are negligible.

Let's analyse the axis 1 and 2.

```{r}
a <- fviz_mfa_var(res.mfa, 
             axes = c(1,2), 
             repel = TRUE, 
             col.var = custom_groups,  
             ggtheme = theme_minimal(),
             title = "") +
  labs(color='')  +
  theme(legend.position = "bottom")

b <- fviz_mfa_ind(res.mfa,
            axes = c(1,2),
            repel = TRUE,
            habillage = factor(month(data_diff$Corrected_Dates)),
            ggtheme = theme_minimal(),
            title = "") +
  labs(color='Months')

ggarrange(a, b, labels = c("MFA Variables Plot", "MFA Individuals Plot"))
```

```{r}
ggplotly(b) %>% 
  layout(hovermode = "x unified")
```
