---
title: "MFA with S1's data"
author: "CLETZ Laura"
date: "04/07/2025"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Packages needed:
```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggpubr)
library(stringr)
library(lubridate)
library(FactoMineR)
library(factoextra)
here::i_am("Correlations_and_Predictions/MFA_S1.Rmd")
library(here)
```

```{r}
data_small <- read.table(here("processed","S1_gamm_2004_2010.csv"), sep=";", dec=".", header=T)[c(1,3:4,6:9,12,15:18)] %>%
  mutate_at(1, function(x) as.Date(x, format="%Y-%m-%d"))

data_big <- read.table(here("processed","S1_gamm_2004_2019.csv"), sep=";", dec=".", header=T)[c(1,3:4,6:9,12,15:18)] %>%
  mutate_at(1, function(x) as.Date(x, format="%Y-%m-%d"))

data_diff <- data_big %>% anti_join(data_small)
```

# From 2004 to 2010

```{r}
res.mfa <- MFA(data_small[-1], group=c(6,5), graph=F)
barplot(res.mfa$eig[,2])
```

The first two components seem the most importants, but the third and the forth components have near or above 10% of the total of inertia.

```{r}
res.mfa$quanti.var$contrib
```
Conductivité, Arsenic_Total, Zinc and Ion_Ferreux contribute the most to the first component.
Rainfall, pH, O2 and Plomb contribute the most to the second one.
Rainfall_lag_1, Temperature contribute the most to the third one.
pH and O2 contribute the most to the forth one, that means the forth and the second components are somewhat alike.
Rainfall, Rainfall_lag_1, Sulfate and Ion_Ferreux contribute the most to the fifth component that we will not consider.

Most of the time, physicochemical elements are the most contributing to any axis. It also seems that when metalloids contribute, they're strongly correlated to dilution and thus to physicochemical measures.

```{r}
gr_names <- rownames(res.mfa$partial.axes$coord)
gr_names <- ifelse(gr_names %in% gr_names[c(1:5)],
                        str_replace(gr_names, "Gr1", "Élém. ph-chimiques + temps"),
                        str_replace(gr_names, "Gr2", "Élém. traces"))
rownames(res.mfa$partial.axes$coord) <- gr_names

fviz_mfa_axes(res.mfa,
            repel = TRUE, 
            ggtheme = theme_minimal(),
            title = "MFA Partial Axis Plot")
```

The first dimension is well balanced when it comes to groups contribution but their positions are perpendicular, meaning they don't fit the first dimension of a regular PCA.
The second dimension is mostly due to physicochemicals parameters.
The third dimension is mostly due to metalloid variables.
The forth and fifth dimensions are negligible.

Let's analyse axis 1, 2 and 3 remembering their contributors.

## Axis 1 and 2

```{r}
var_names <- rownames(res.mfa$quanti.var$coord)
custom_groups <- ifelse(var_names %in% var_names[c(1:6)],
                        "Élém. ph-chimiques + temps",
                        "Élém. traces")

a <- fviz_mfa_var(res.mfa, 
             axes = c(1,2), 
             repel = TRUE, 
             col.var = custom_groups,  
             ggtheme = theme_minimal(),
             title = "") +
  labs(color='')  +
  theme(legend.position = "bottom")

b <- fviz_mfa_ind(res.mfa,
            axes = c(1,2),
            repel = TRUE,
            habillage = factor(month(data_small$Corrected_Dates)),
            ggtheme = theme_minimal(),
            title = "") +
  labs(color='Months')

ggarrange(a, b, labels = c("MFA Variables Plot", "MFA Individuals Plot"))
```

```{r}
library(plotly)
ggplotly(b) %>% 
  layout(hovermode = "x unified")
```

## Axis 1 and 3

```{r}
a <- fviz_mfa_var(res.mfa, 
             axes = c(1,3), 
             repel = TRUE, 
             col.var = custom_groups,  
             ggtheme = theme_minimal(),
             title = "") +
  labs(color='')  +
  theme(legend.position = "bottom")

b <- fviz_mfa_ind(res.mfa,
            axes = c(1,3),
            repel = TRUE,
            habillage = factor(month(data_small$Corrected_Dates)),
            ggtheme = theme_minimal(),
            title = "") +
  labs(color='Months')

ggarrange(a, b, labels = c("MFA Variables Plot", "MFA Individuals Plot"))
```

```{r}
library(plotly)
ggplotly(b) %>% 
  layout(hovermode = "x unified")
```

## Axis 2 and 3

```{r}
a <- fviz_mfa_var(res.mfa, 
             axes = c(2,3), 
             repel = TRUE, 
             col.var = custom_groups,  
             ggtheme = theme_minimal(),
             title = "") +
  labs(color='')  +
  theme(legend.position = "bottom")

b <- fviz_mfa_ind(res.mfa,
            axes = c(2,3),
            repel = TRUE,
            habillage = factor(month(data_small$Corrected_Dates)),
            ggtheme = theme_minimal(),
            title = "") +
  labs(color='Months')

ggarrange(a, b, labels = c("MFA Variables Plot", "MFA Individuals Plot"))
```

```{r}
ggplotly(b) %>% 
  layout(hovermode = "x unified")
```

# From 2011 to 2019

```{r}
res.mfa <- MFA(data_diff[-1], group=c(6,5), graph=F)
barplot(res.mfa$eig[,2])
```
The first component has more than 50% of the variability, we'll probably need to study only the first two dimensions.

```{r}
res.mfa$quanti.var$contrib
```
pH, Conductivité, Arsenic_Total, Plomb, Zinc and Ion_Ferreux contribute the most to the first component.
Rainfall and Rainfall_lag_1 contribute the most to the second one.
Rainfall and Rainfall_lag_1 contribute the most to the third one, we can read the second and the third components as one and the same.
Temperature contributes the most to the forth one.
Rainfall and O2 contribute the most to the fifth one.

```{r}
gr_names <- rownames(res.mfa$partial.axes$coord)
gr_names <- ifelse(gr_names %in% gr_names[c(1:5)],
                        str_replace(gr_names, "Gr1", "Élém. ph-chimiques + temps"),
                        str_replace(gr_names, "Gr2", "Élém. traces"))
rownames(res.mfa$partial.axes$coord) <- gr_names

fviz_mfa_axes(res.mfa,
            repel = TRUE, 
            ggtheme = theme_minimal(),
            title = "MFA Partial Axis Plot")
```
The dimension 1 in both group almost fits the natural dimension 1.
The dimension 2 also but one contributes positively while the other contributes negatively.
The other dimensions are negligible.

Let's analyse the axis 1 and 2.

```{r}
a <- fviz_mfa_var(res.mfa, 
             axes = c(1,2), 
             repel = TRUE, 
             col.var = custom_groups,  
             ggtheme = theme_minimal(),
             title = "") +
  labs(color='')  +
  theme(legend.position = "bottom")

b <- fviz_mfa_ind(res.mfa,
            axes = c(1,2),
            repel = TRUE,
            habillage = factor(month(data_diff$Corrected_Dates)),
            ggtheme = theme_minimal(),
            title = "") +
  labs(color='Months')

ggarrange(a, b, labels = c("MFA Variables Plot", "MFA Individuals Plot"))
```

```{r}
ggplotly(b) %>% 
  layout(hovermode = "x unified")
```


