---
title: "ARIMA and GAMMs predictions at station S1 using COWG's data"
author: "CLETZ Laura"
date: "08/07/2025"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Packages needed:
```{r}
library(dplyr)
library(mgcv)
library(gam)
library(forecast)
here::i_am("Correlations_and_Predictions/S1_by_COWG_predictions.Rmd")
library(here)
```

In this file, we will predict S1 using COWG where we couldn't in the file named "S1_predictions.Rmd" as long as we predicted correctly the corresponding variables in COWG.

```{r}
#data <- read.table(here('processed', 'COWG_gamm_2004_2022.csv'), dec = '.', sep = ';', header = T) %>%
#  mutate_at(1:2, function(x) as.Date(x, format="%Y-%m-%d"))

data <- read.table(here('processed', 'COWG_gamm_2004_2022_twice.csv'), dec = '.', sep = ';', header = T) %>%
  mutate_at(1:2, function(x) as.Date(x, format="%Y-%m-%d"))
```

```{r}
#S1 <- read.table(here('processed', 'S1_gamm_2004_2019.csv'), dec = '.', sep = ';', header = T) %>%
#  mutate_at(1:2, function(x) as.Date(x, format="%Y-%m-%d"))

S1 <- read.table(here('processed', 'S1_gamm_2004_2019_twice.csv'), dec = '.', sep = ';', header = T) %>%
  mutate_at(1:2, function(x) as.Date(x, format="%Y-%m-%d"))
```

```{r}
new_data <- data.frame(
  Dates = as.Date("2025-04-24", format = "%Y-%m-%d"),
  Temperature = 16.8,
  Rainfall = 107.4,
  Rainfall_lag_1 = 243.1,
  Conductivité = 4560,
  pH = 2.97,
  O2 = 2.86,
  Ion_Ferreux = 1069.64,
  Sulfate = (1240903.62 * 96) / 32000,
  Zinc = 19337.51,
  Arsenic_Total = 115181.67,
  Plomb = 1614.53
  ) %>%
  mutate(Numeric_Dates = as.numeric(Dates)) 
```

```{r}
test <- read.table(here('processed','Source_S1.csv'), dec = '.', sep = ";", header = TRUE)[77:174,] %>%
  mutate_at(1:2, function(x) as.Date(x, format="%Y-%m-%d")) %>%
  mutate(Numeric_Dates = as.numeric(Corrected_Dates)) %>%
  select(names(new_data))
```

# Temperature

We recall that : S1 = 7.81202 + 0.51873 * COWG

```{r}
temp <- ts(data$Temperature, start = c(2004,11), end = c(2022,4), frequency = 12)

fit_temp <- arima(temp, order = c(1, 1, 1),
      seasonal = list(order = c(1, 1, 1), period = 12), method = "CSS-ML")

pred_temp <- predict(fit_temp, n.ahead = 51) 
```

```{r}
temp <- ts(S1$Temperature, start = c(2004,11), end = c(2019,1), frequency = 12)

plot.ts(temp, type='l', xlab='Temps', ylab="Température (C°)",
        xlim = c(2004,2026), xaxt = 'n')

cowg <- ts(7.81202 + 0.51873 * data$Temperature[174:213], start = c(2019,1), end = c(2022,4), frequency = 12)
points(cowg, type='l', col='blue')

ts <- ts(7.81202 + 0.51873 * pred_temp$pred, start = c(2022,4), end = c(2026,6), frequency = 12)
ts_se <- ts(pred_temp$se, start = c(2022,4), end = c(2026,6), frequency = 12)

x_coords <- c(time(ts), rev(time(ts)))
y_coords <- c(ts + ts_se, rev(ts - ts_se))
polygon(x_coords, y_coords, col = 'cyan3', border = "cyan")

points(ts, type='l', col='blue')
points(ts(test$Temperature, start = c(2011,2), end = c(2019,1), frequency = 12),
       pch = 19, col = "darkred")

points(2025.33, new_data$Temperature, col='red', pch=19)

abline(h=13, col='red')
abline(h=18, col='red')

legend(x = "bottomright",
       legend=c("S1", "7.81202 + 0.51873 * COWG", "Intervalle de prédiction", "Points non-imputés", "Point test"),
       col=c("black", "blue", "cyan3", "darkred", "red"),
       pch=c(NA, NA, NA, 19, 19),
       lty=c(1, 1, 1, NA, NA),
       lwd=c(1, 1, 5, 1, 1),
       bty = "n")

axis(1, at = seq(2004, 2026, by = 1), las = 2)
```

The prediction curve is accurate!

# O2

We recall that : S1 = -0.3946 + 0.2911 * COWG (intercept NS)

```{r}
o2 <- ts(data$O2, start = c(2004,11), end = c(2022,4), frequency = 12)

fit_o2 <- arima(o2, order = c(1, 1, 1),
      seasonal = list(order = c(1, 1, 1), period = 12), method = "CSS-ML")

pred_o2 <- predict(fit_o2, n.ahead = 51) 
```

```{r}
o2 <- ts(S1$O2, start = c(2004,11), end = c(2019,1), frequency = 12)

plot.ts(o2, type='l', xlab='Temps', ylab="O2",
        xlim = c(2004,2026), xaxt = 'n')

cowg <- ts(-0.3946 + 0.2911 * data$O2[174:213], start = c(2019,1), end = c(2022,4), frequency = 12)
points(cowg, type='l', col='blue')

ts <- ts(-0.3946 + 0.2911 * pred_o2$pred, start = c(2022,4), end = c(2026,6), frequency = 12)
ts_se <- ts(pred_o2$se, start = c(2022,4), end = c(2026,6), frequency = 12)
x_coords <- c(time(ts), rev(time(ts)))

y_coords <- c(ts + ts_se, rev(ts - ts_se))
polygon(x_coords, y_coords, col = 'cyan3', border = "cyan")

points(ts, type='l', col='blue')
points(ts(test$O2, start = c(2011,2), end = c(2019,1), frequency = 12),
       pch = 19, col = "darkred")

points(2025.33, new_data$O2, col='red', pch=19)

abline(h=0.5, col='red')
abline(h=3.5, col='red')

legend(x = "topright",
       legend=c("S1", "0.7349 + 0.6365 * COWG", "Intervalle de prédiction", "Points non-imputés", "Point test"),
       col=c("black", "blue", "cyan3", "darkred", "red"),
       pch=c(NA, NA, NA, 19, 19),
       lty=c(1, 1, 1, NA, NA),
       lwd=c(1, 1, 5, 1, 1),
       cex=c(0.6, 0.6, 0.6, 0.6, 0.6),
       bty = "n")

axis(1, at = seq(2004, 2026, by = 1), las = 2)
```

# pH

We recall that : S1 = 0.7349 + 0.6365 * COWG (intercept not much significant)

```{r}
ph <- ts(data$pH, start = c(2004,11), end = c(2022,4), frequency = 12)

fit_ph <- arima(ph, order = c(1, 1, 1),
      seasonal = list(order = c(1, 1, 1), period = 12), method = "CSS-ML")

pred_ph <- predict(fit_ph, n.ahead = 51) 
```

```{r}
ph <- ts(S1$pH, start = c(2004,11), end = c(2019,1), frequency = 12)

plot.ts(ph, type='l', xlab='Temps', ylab="pH",
        xlim = c(2004,2026), xaxt = 'n')

cowg <- ts(0.7349 + 0.6365 * data$pH[174:213], start = c(2019,1), end = c(2022,4), frequency = 12)
points(cowg, type='l', col='blue')

ts <- ts(0.7349 + 0.6365 * pred_ph$pred, start = c(2022,4), end = c(2026,6), frequency = 12)
ts_se <- ts(pred_ph$se, start = c(2022,4), end = c(2026,6), frequency = 12)
x_coords <- c(time(ts), rev(time(ts)))

y_coords <- c(ts + ts_se, rev(ts - ts_se))
polygon(x_coords, y_coords, col = 'cyan3', border = "cyan")

points(ts, type='l', col='blue')
points(ts(test$pH, start = c(2011,2), end = c(2019,1), frequency = 12),
       pch = 19, col = "darkred")

points(2025.33, new_data$pH, col='red', pch=19)

abline(h=2.5, col='red')
abline(h=3.5, col='red')

legend(x = "topright",
       legend=c("S1", "0.7349 + 0.6365 * COWG", "Intervalle de prédiction", "Points non-imputés", "Point test"),
       col=c("black", "blue", "cyan3", "darkred", "red"),
       pch=c(NA, NA, NA, 19, 19),
       lty=c(1, 1, 1, NA, NA),
       lwd=c(1, 1, 5, 1, 1),
       cex=c(0.6, 0.6, 0.6, 0.6, 0.6),
       bty = "n")

axis(1, at = seq(2004, 2026, by = 1), las = 2)
```
Since the prediction didn't fit on S1's data, that oddly decreasing curve the imputed data made doesn't join the predicted curve but the predicted curve predicts accurately the point in 2025!

# Arsenic_Total

We recall that : S1 = 1.021e+05 + 4.862e-01 * COWG

```{r}
ars <- ts(data$Arsenic_Total, start = c(2004,11), end = c(2022,4), frequency = 12)

fit_ars <- arima(ars, order = c(1, 1, 1),
      seasonal = list(order = c(1, 1, 1), period = 12), method = "CSS-ML")

pred_ars <- predict(fit_ars, n.ahead = 51) 
```

```{r}
ars <- ts(S1$Arsenic_Total, start = c(2004,11), end = c(2019,1), frequency = 12)

plot.ts(ars, type='l', xlab='Temps', ylab="Concentration en arsenic (ppb)",
        xlim = c(2004,2026), xaxt = 'n')

cowg <- ts(1.021e+05 + 4.862e-01 * data$Arsenic_Total[174:213], start = c(2019,1), end = c(2022,4), frequency = 12)
points(cowg, type='l', col='blue')

ts <- ts(1.021e+05 + 4.862e-01 * pred_ars$pred, start = c(2022,4), end = c(2026,6), frequency = 12)
ts_se <- ts(pred_ars$se, start = c(2022,4), end = c(2026,6), frequency = 12)
x_coords <- c(time(ts), rev(time(ts)))
y_coords <- c(ts + ts_se, rev(ts - ts_se))
polygon(x_coords, y_coords, col = 'cyan3', border = "cyan")

points(ts, type='l', col='blue')
points(ts(test$Arsenic_Total, start = c(2011,2), end = c(2019,1), frequency = 12),
       pch = 19, col = "darkred")

points(2025.33, new_data$Arsenic_Total, col='red', pch=19)

abline(h=200000, col='red')
abline(h=100000, col='red')

legend(x = "bottomright",
       legend=c("S1", "1.021e+05 + 4.862e-01 * COWG", "Intervalle de prédiction", "Points non-imputés", "Point test"),
       col=c("black", "blue", "cyan3", "darkred", "red"),
       pch=c(NA, NA, NA, 19, 19),
       lty=c(1, 1, 1, NA, NA),
       lwd=c(1, 1, 5, 1, 1),
       cex=c(0.6, 0.6, 0.6, 0.6, 0.6),
       bty = "n")

axis(1, at = seq(2004, 2026, by = 1), las = 2)
```
The imputed data whether that be from S1 or COWG might be too high, we can see that the non-missing data seem to decrease but the imputation had an increasing effect. We can't really predict S1 by COWG when it comes to Arsenic...

# Zinc

We recall that : S1 = 1.069e+04 + 6.795e-01 * COWG

```{r}
zinc <- ts(data$Zinc, start = c(2004,11), end = c(2022,4), frequency = 12)

fit_zinc <- arima(zinc, order = c(1, 1, 1),
      seasonal = list(order = c(1, 1, 1), period = 12), method = "CSS-ML")

pred_zinc <- predict(fit_zinc, n.ahead = 51) 
```

```{r}
zinc <- ts(S1$Zinc, start = c(2004,11), end = c(2019,1), frequency = 12)

plot.ts(zinc, type='l', xlab='Temps', ylab="Concentration en zinc (ppb)",
        xlim = c(2004,2026), xaxt = 'n')

cowg <- ts(1.069e+04 + 6.795e-01 * data$Zinc[174:213], start = c(2019,1), end = c(2022,4), frequency = 12)
points(cowg, type='l', col='blue')

ts <- ts(1.069e+04 + 6.795e-01 * pred_zinc$pred, start = c(2022,4), end = c(2026,6), frequency = 12)
ts_se <- ts(pred_zinc$se, start = c(2022,4), end = c(2026,6), frequency = 12)
x_coords <- c(time(ts), rev(time(ts)))
y_coords <- c(ts + ts_se, rev(ts - ts_se))
polygon(x_coords, y_coords, col = 'cyan3', border = "cyan")

points(ts, type='l', col='blue')
points(ts(test$Zinc, start = c(2011,2), end = c(2019,1), frequency = 12),
       pch = 19, col = "darkred")

points(2025.33, new_data$Zinc, col='red', pch=19)

abline(h=15000, col='red')
abline(h=30000, col='red')

legend(x = "topright",
       legend=c("S1", "1.069e+04 + 6.795e-01 * COWG", "Intervalle de prédiction", "Points non-imputés", "Point test"),
       col=c("black", "blue", "cyan3", "darkred", "red"),
       pch=c(NA, NA, NA, 19, 19),
       lty=c(1, 1, 1, NA, NA),
       lwd=c(1, 1, 5, 1, 1),
       cex=c(0.6, 0.6, 0.6, 0.6, 0.6),
       bty = "n")

axis(1, at = seq(2004, 2026, by = 1), las = 2)
```

We can't predict Zinc in S1 from COWG data.

# Plomb

We recall that : S1 = 1.163e+03 + 4.234e-01 * COWG

```{r}
lead <- ts(data$Plomb, start = c(2004,11), end = c(2022,4), frequency = 12)

fit_lead <- arima(lead, order = c(1, 1, 1),
      seasonal = list(order = c(1, 1, 1), period = 12), method = "CSS-ML")

pred_lead <- predict(fit_lead, n.ahead = 51) 
```

```{r}
lead <- ts(S1$Plomb, start = c(2004,11), end = c(2019,1), frequency = 12)

plot.ts(lead, type='l', xlab='Temps', ylab="Concentration en Plomb (ppb)",
        xlim = c(2004,2026), xaxt = 'n')

cowg <- ts(1.163e+03 + 4.234e-01 * data$Plomb[174:213], start = c(2019,1), end = c(2022,4), frequency = 12)
points(cowg, type='l', col='blue')

ts <- ts(1.163e+03 + 4.234e-01 * pred_lead$pred, start = c(2022,4), end = c(2026,6), frequency = 12)
ts_se <- ts(pred_lead$se, start = c(2022,4), end = c(2026,6), frequency = 12)
x_coords <- c(time(ts), rev(time(ts)))
y_coords <- c(ts + ts_se, rev(ts - ts_se))
polygon(x_coords, y_coords, col = 'cyan3', border = "cyan")

points(ts, type='l', col='blue')
points(ts(test$Plomb, start = c(2011,2), end = c(2019,1), frequency = 12),
       pch = 19, col = "darkred")

points(2025.33, new_data$Plomb, col='red', pch=19)

abline(h=15000, col='red')
abline(h=25000, col='red')

legend(x = "topright",
       legend=c("S1", "1.163e+03 + 4.234e-01  * COWG", "Intervalle de prédiction", "Points non-imputés", "Point test"),
       col=c("black", "blue", "cyan3", "darkred", "red"),
       pch=c(NA, NA, NA, 19, 19),
       lty=c(1, 1, 1, NA, NA),
       lwd=c(1, 1, 5, 1, 1),
       cex=c(0.6, 0.6, 0.6, 0.6, 0.6),
       bty = "n")

axis(1, at = seq(2004, 2026, by = 1), las = 2)
```

It's good!

# Ion_Ferreux

We recall that : S1 = 735.90420 + 0.58520 * COWG

```{r}
fer <- ts(data$Ion_Ferreux, start = c(2004,11), end = c(2022,4), frequency = 12)

fit_fer <- arima(fer, order = c(1, 1, 1),
      seasonal = list(order = c(1, 1, 1), period = 12), method = "CSS-ML")

pred_fer <- predict(fit_fer, n.ahead = 51) 
```

```{r}
fer <- ts(S1$Ion_Ferreux, start = c(2004,11), end = c(2019,1), frequency = 12)

plot.ts(fer, type='l', xlab='Temps', ylab="Concentration en FeII (ppb)",
        xlim = c(2004,2026), xaxt = 'n')

cowg <- ts(735.90420 + 0.58520 * data$Ion_Ferreux[174:213], start = c(2019,1), end = c(2022,4), frequency = 12)
points(cowg, type='l', col='blue')

ts <- ts(735.90420 + 0.58520 * pred_fer$pred, start = c(2022,4), end = c(2026,6), frequency = 12)
ts_se <- ts(pred_fer$se, start = c(2022,4), end = c(2026,6), frequency = 12)
x_coords <- c(time(ts), rev(time(ts)))
y_coords <- c(ts + ts_se, rev(ts - ts_se))
polygon(x_coords, y_coords, col = 'cyan3', border = "cyan")

points(ts, type='l', col='blue')
points(ts(test$Ion_Ferreux, start = c(2011,2), end = c(2019,1), frequency = 12),
       pch = 19, col = "darkred")

points(2025.33, new_data$Ion_Ferreux, col='red', pch=19)

abline(h=15000, col='red')
abline(h=25000, col='red')

legend(x = "topright",
       legend=c("S1", "735.90420 + 0.58520 * COWG", "Intervalle de prédiction", "Points non-imputés", "Point test"),
       col=c("black", "blue", "cyan3", "darkred", "red"),
       pch=c(NA, NA, NA, 19, 19),
       lty=c(1, 1, 1, NA, NA),
       lwd=c(1, 1, 5, 1, 1),
       cex=c(0.6, 0.6, 0.6, 0.6, 0.6),
       bty = "n")

axis(1, at = seq(2004, 2026, by = 1), las = 2)
```

Almost...

