---
title: "ARIMA and GAMMs predictions at station S1"
author: "CLETZ Laura"
date: "20/06/2025"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Packages needed:
```{r}
library(dplyr)
library(mgcv)
library(gam)
library(forecast)
here::i_am("Correlations_and_Predictions/S1_predictions.Rmd")
library(here)
```

```{r}
#data <- read.table(here('processed', 'S1_gamm_2004_2019.csv'), dec = '.', sep = ';', header = T) %>%
#  mutate_at(1:2, function(x) as.Date(x, format="%Y-%m-%d"))

data <- read.table(here('processed', 'S1_gamm_2004_2019_twice.csv'), dec = '.', sep = ';', header = T) %>%
  mutate_at(1:2, function(x) as.Date(x, format="%Y-%m-%d"))
```

```{r}
data_small <- read.table(here('processed', 'S1_gamm_2004_2010.csv'), dec = '.', sep = ';', header = T) %>%
  mutate_at(1:2, function(x) as.Date(x, format="%Y-%m-%d"))
```

```{r}
new_data <- data.frame(
  dates = as.Date("2025-04-24", format = "%Y-%m-%d"),
  temperature = 16.8,
  rainfall = 107.4,
  rainfall_lag_1 = 243.1,
  conductivity = 4560,
  ph = 2.97,
  o2 = 2.86,
  feII = 1069.64,
  sulfate = (1240903.62 * 96) / 32000,
  zinc = 19337.51,
  total_arsenic = 115181.67,
  lead = 1614.53
  ) %>%
  mutate(numeric_dates = as.numeric(dates)) 
```

```{r}
test <- read.table(here('processed','Source_S1.csv'), dec = '.', sep = ";", header = TRUE)[77:174,] %>%
  mutate_at(1:2, function(x) as.Date(x, format="%Y-%m-%d")) %>%
  mutate(numeric_dates = as.numeric(corrected_dates)) %>%
  select(names(new_data))
```

```{r}
newest_data <- data.frame(
  dates = as.Date("2025-07-17", format = "%Y-%m-%d"),
  temperature = 16.6,
  ph = 3.30,
  eh = 232,
  conductivity = 4690,
  o2 = 2.28
)
```

# temperature
## Whole data

```{r}
temp <- ts(data$temperature, start = c(2004,11), end = c(2019,1), frequency = 12)

fit_temp <- arima(temp, order = c(1, 1, 1),
      seasonal = list(order = c(1, 1, 1), period = 12), method = "CSS-ML")

pred_temp <- predict(fit_temp, n.ahead = 90) # because difftime("2026-06-01","2019-01-01")/30 = 90
```

```{r}
plot.ts(temp, type='l', xlab='Temps', ylab="Température (C°)",
        xlim = c(2004,2026), xaxt = 'n')

ts <- ts(pred_temp$pred, start = c(2019,1), end = c(2026,6), frequency = 12)
ts_se <- ts(pred_temp$se, start = c(2019,1), end = c(2026,6), frequency = 12)
x_coords <- c(time(ts), rev(time(ts)))
y_coords <- c(ts + ts_se, rev(ts - ts_se))
polygon(x_coords, y_coords, col = 'cyan3', border = "cyan")

points(ts, type='l', col='blue')
points(ts(test$temperature, start = c(2011,2), end = c(2019,1), frequency = 12),
       pch = 19, col = "darkred")

points(2025.33, new_data$temperature, col='red', pch=19)
points(2025.583, newest_data$temperature, col='red', pch=19)

abline(h=13, col='red')
abline(h=18, col='red')

legend(x = "bottomright",
       legend=c("S1", "Intervalle de prédiction", "Points non-imputés", "Point test"),
       col=c("black", "cyan3", "darkred", "red"),
       pch=c(NA, NA, 19, 19),
       lty=c(1, 1, NA, NA),
       lwd=c(1, 5, 1, 1),
       cex=c(0.6, 0.6, 0.6, 0.6),
       bty = "n")

axis(1, at = seq(2004, 2026, by = 1), las = 2)
```

## Small data

```{r}
temp <- ts(data_small$temperature, start = c(2004,11), end = c(2011,1), frequency = 12)

fit_temp <- arima(temp, order = c(1, 1, 1),
      seasonal = list(order = c(1, 1, 1), period = 12), method = "CSS-ML")

pred_temp <- predict(fit_temp, n.ahead = 187) # because difftime("2026-06-01","2011-01-01")/30 = 187
```

```{r}
plot.ts(temp, type='l', xlab='Temps', ylab="Température (C°)",
        xlim = c(2004,2026), xaxt = 'n')

ts <- ts(pred_temp$pred, start = c(2011,1), end = c(2026,6), frequency = 12)
ts_se <- ts(pred_temp$se, start = c(2011,1), end = c(2026,6), frequency = 12)
x_coords <- c(time(ts), rev(time(ts)))
y_coords <- c(ts + ts_se, rev(ts - ts_se))
polygon(x_coords, y_coords, col = 'cyan3', border = "cyan")

points(ts, type='l', col='blue')
points(ts(test$temperature, start = c(2011,2), end = c(2019,1), frequency = 12),
       pch = 19, col = "red")

points(2025.33, new_data$temperature, col='red', pch=19)
points(2025.583, newest_data$temperature, col='red', pch=19)

abline(h=13, col='red')
abline(h=18, col='red')

legend(x = "bottomright",
       legend=c("S1", "Intervalle de prédiction", "Points tests"),
       col=c("black", "cyan3",  "red"),
       pch=c(NA, NA, 19),
       lty=c(1, 1, NA),
       lwd=c(1, 5, 1),
       cex=c(0.6, 0.6, 0.6),
       bty = "n")

axis(1, at = seq(2004, 2026, by = 1), las = 2)
```

Both prediction are valid, but the one with the whole is more precise.

# o2
## Whole data

```{r}
o2 <- ts(data$o2, start = c(2004,11), end = c(2019,1), frequency = 12)

fit_o2 <- arima(data$o2, order = c(1, 1, 2),
      seasonal = list(order = c(1, 1, 2), period = 12), method = "CSS-ML")

pred_o2 <- predict(fit_o2, n.ahead = 90)
```

```{r}
plot.ts(o2, type='l', xlab='Temps', ylab="o2",
        xlim = c(2004,2026), xaxt = 'n')

ts <- ts(pred_o2$pred, start = c(2019,1), end = c(2026,6), frequency = 12)
ts_se <- ts(pred_o2$se, start = c(2019,1), end = c(2026,6), frequency = 12)
x_coords <- c(time(ts), rev(time(ts)))
y_coords <- c(ts + ts_se, rev(ts - ts_se))
polygon(x_coords, y_coords, col = 'cyan3', border = "cyan")

points(ts, type='l', col='blue')
points(2025.33, new_data$o2, col="red", pch=19)
points(2025.583, newest_data$o2, col='red', pch=19)

points(ts(test$o2, start = c(2011,2), end = c(2019,1), frequency = 12),
       pch = 19, col = "darkred")

axis(1, at = seq(2004, 2026, by = 1), las = 2)
```

## Small data

```{r}
o2 <- ts(data_small$o2, start = c(2004,11), end = c(2011,1), frequency = 12)

fit_o2 <- arima(o2, order = c(1, 1, 1),
      seasonal = list(order = c(1, 1, 1), period = 12), method = "CSS-ML")

pred_o2 <- predict(fit_o2, n.ahead = 187)
```

```{r}
plot.ts(o2, type='l', xlab='Temps', ylab="o2",
        xlim = c(2004,2026), xaxt = 'n')

ts <- ts(pred_o2$pred, start = c(2011,1), end = c(2026,6), frequency = 12)
ts_se <- ts(pred_o2$se, start = c(2011,1), end = c(2026,6), frequency = 12)
x_coords <- c(time(ts), rev(time(ts)))
y_coords <- c(ts + ts_se, rev(ts - ts_se))
polygon(x_coords, y_coords, col = 'cyan3', border = "cyan")

points(ts, type='l', col='blue')
points(ts(test$o2, start = c(2011,2), end = c(2019,1), frequency = 12),
       pch = 19, col = "darkred")

points(2025.33, new_data$o2, col='red', pch=19)
points(2025.583, newest_data$o2, col='red', pch=19)

abline(h=13, col='red')
abline(h=18, col='red')

axis(1, at = seq(2004, 2026, by = 1), las = 2)
```

The second prediction is not precise enough.

# ph
## Whole data

```{r}
ph <- ts(data$ph, start = c(2004,11), end = c(2019,1), frequency = 12)

fit_ph <- arima(data$ph, order = c(1, 1, 1),
      seasonal = list(order = c(1, 1, 1), period = 12), method = "CSS-ML")

pred_ph <- predict(fit_ph, n.ahead = 90)
```

```{r}
plot.ts(ph, type='l', xlab='Temps', ylab="ph",
        xlim = c(2004,2026), ylim = c(0,7), xaxt = 'n')

ts <- ts(pred_ph$pred, start = c(2019,1), end = c(2026,6), frequency = 12)
ts_se <- ts(pred_ph$se, start = c(2019,1), end = c(2026,6), frequency = 12)
x_coords <- c(time(ts), rev(time(ts)))
y_coords <- c(ts + ts_se, rev(ts - ts_se))
polygon(x_coords, y_coords, col = 'cyan3', border = "cyan")

points(ts, type='l', col='blue')
points(2025.33, new_data$ph, col="red", pch=19)
points(2025.583, newest_data$ph, col='red', pch=19)

points(ts(test$ph, start = c(2011,2), end = c(2019,1), frequency = 12),
       pch = 19, col = "darkred")

axis(1, at = seq(2004, 2026, by = 1), las = 2)
```

## Small data

```{r}
ph <- ts(data_small$ph, start = c(2004,11), end = c(2011,1), frequency = 12)

fit_ph <- arima(ph, order = c(1, 1, 1),
      seasonal = list(order = c(1, 1, 1), period = 12), method = "CSS-ML")

pred_ph <- predict(fit_ph, n.ahead = 187)
```

```{r}
plot.ts(ph, type='l', xlab='Temps', ylab="ph",
        xlim = c(2004,2026), xaxt = 'n')

ts <- ts(pred_ph$pred, start = c(2011,1), end = c(2026,6), frequency = 12)
ts_se <- ts(pred_ph$se, start = c(2011,1), end = c(2026,6), frequency = 12)
x_coords <- c(time(ts), rev(time(ts)))
y_coords <- c(ts + ts_se, rev(ts - ts_se))
polygon(x_coords, y_coords, col = 'cyan3', border = "cyan")

points(ts, type='l', col='blue')
points(ts(test$ph, start = c(2011,2), end = c(2019,1), frequency = 12),
       pch = 19, col = "darkred")

points(2025.33, new_data$ph, col='red', pch=19)
points(2025.583, newest_data$ph, col='red', pch=19)

abline(h=13, col='red')
abline(h=18, col='red')

axis(1, at = seq(2004, 2026, by = 1), las = 2)
```

# total_arsenic
## Whole data

```{r}
ars <- ts(data$total_arsenic, start = c(2004,11), end = c(2019,1), frequency = 12)

fit_ars <- arima(data$total_arsenic, order = c(1, 1, 1),
      seasonal = list(order = c(1, 1, 1), period = 12), method = "CSS-ML")

pred_ars <- predict(fit_ars, n.ahead = 90)
```

```{r}
plot.ts(ars, type='l', xlab='Temps', ylab="Concentration en Arsenic (ppb)",
        xlim = c(2004,2026), xaxt = 'n')

ts <- ts(pred_ars$pred, start = c(2019,1), end = c(2026,6), frequency = 12)
ts_se <- ts(pred_ars$se, start = c(2019,1), end = c(2026,6), frequency = 12)
x_coords <- c(time(ts), rev(time(ts)))
y_coords <- c(ts + ts_se, rev(ts - ts_se))
polygon(x_coords, y_coords, col = 'cyan3', border = "cyan")

points(ts, type='l', col='blue')
points(2025.33, new_data$total_arsenic, col="red", pch=19)
points(ts(test$total_arsenic, start = c(2011,2), end = c(2019,1), frequency = 12),
       pch = 19, col = "darkred")

legend(x = "bottomright",
       legend=c("S1", "Intervalle de prédiction", "Points non-imputés", "Points tests"),
       col=c("black", "cyan3", "darkred", "red"),
       pch=c(NA, NA, 19, 19),
       lty=c(1, 1, NA, NA),
       lwd=c(1, 5, 1, 1),
       cex=c(0.6, 0.6, 0.6, 0.6),
       bty = "n")

abline(h=200000, col='red')
abline(h=75000, col='red')

axis(1, at = seq(2004, 2026, by = 1), las = 2)
```

Absolutely not a good prediction.

## Small data

```{r}
ars <- ts(data_small$total_arsenic, start = c(2004,11), end = c(2011,1), frequency = 12)

fit_ars <- arima(ars, order = c(1, 1, 1),
      seasonal = list(order = c(1, 1, 1), period = 12), method = "CSS-ML")

pred_ars <- predict(fit_ars, n.ahead = 187)
```

```{r}
plot.ts(ars, type='l', xlab='Temps', ylab="Concentration en arsenic (ppb)",
        xlim = c(2004,2026), xaxt = 'n')

ts <- ts(pred_ars$pred, start = c(2011,1), end = c(2026,6), frequency = 12)
ts_se <- ts(pred_ars$se, start = c(2011,1), end = c(2026,6), frequency = 12)
x_coords <- c(time(ts), rev(time(ts)))
y_coords <- c(ts + ts_se, rev(ts - ts_se))
polygon(x_coords, y_coords, col = 'cyan3', border = "cyan")

points(ts, type='l', col='blue')
points(ts(test$total_arsenic, start = c(2011,2), end = c(2019,1), frequency = 12),
       pch = 19, col = "darkred")

points(2025.33, new_data$total_arsenic, col='red', pch=19)

abline(h=13, col='red')
abline(h=18, col='red')

axis(1, at = seq(2004, 2026, by = 1), las = 2)
```

Better prediction but prediction interval too wide.

## GAMM and whole data

```{r}
ars <- ts(data_small$total_arsenic, start = c(2004,11), end = c(2019,1), frequency = 12)

plot.ts(ars, type='l', xlab='Temps', ylab="Concentration en Arsenic",
        xlim = c(2004,2026), xaxt = 'n')

gam_ars <- gamm(total_arsenic ~ s(rainfall_lag_1, bs='re') + s(o2, bs='re') + s(sulfate, bs='re') + feII + 
                 numeric_dates + lead + rainfall_lag_1:o2 + conductivity^2 + zinc + 
                 feII:lead + o2:feII + zinc:ph + zinc:sulfate + sulfate:feII, 
                data = data, correlation = corARMA(form = ~1, p=1))

pred_ars <- predict.gam(gam_ars$gam, newdata = test, se.fit = TRUE)
points(ts(pred_ars$fit, start = c(2011,2), end = c(2019,1), frequency = 12),
       pch = 15, col = "cyan3")
points(spline(ts(pred_ars$fit, start = c(2011,2), end = c(2019,1), frequency = 12)),
       type='l', col = "cyan3")
points(ts(test$total_arsenic, start = c(2011,2), end = c(2019,1), frequency = 12),
       pch = 19, col = "darkred")

upper <- spline(ts(pred_ars$fit + pred_ars$se.fit, start = c(2011,2), end = c(2019,1), frequency = 12))
lower <- spline(ts(pred_ars$fit - pred_ars$se.fit, start = c(2011,2), end = c(2019,1), frequency = 12))
points(upper, type='l', col = "cyan4")
points(lower, type='l', col = "cyan4")

pred_ars <- predict.gam(gam_ars$gam, newdata = new_data, se.fit = TRUE)
points(2025.33, pred_ars$fit, pch=19, col='cyan3')
points(2025.33, pred_ars$fit + pred_ars$se.fit, pch=15, col='cyan4')
points(2025.33, pred_ars$fit - pred_ars$se.fit, pch=15, col='cyan4')

points(2025.33, new_data$total_arsenic, col="red", pch=19)

axis(1, at = seq(2004, 2026, by = 1), las = 2)
```

Still not conclusive...

## GAMM and small data

```{r}
ars <- ts(data_small$total_arsenic, start = c(2004,11), end = c(2011,1), frequency = 12)

plot.ts(ars, type='l', xlab='Temps', ylab="Concentration en Arsenic",
        xlim = c(2004,2026), xaxt = 'n')

gam_ars <- gamm(total_arsenic ~ s(rainfall_lag_1, bs='re') + s(o2, bs='re') + s(sulfate, bs='re') + feII + 
                 numeric_dates + lead + rainfall_lag_1:o2 + conductivity^2 + zinc + 
                 feII:lead + o2:feII + zinc:ph + zinc:sulfate + sulfate:feII, 
                data = data_small, correlation = corARMA(form = ~1, p=1))

pred_ars <- predict.gam(gam_ars$gam, newdata = test, se.fit = TRUE)
points(ts(pred_ars$fit, start = c(2011,2), end = c(2019,1), frequency = 12),
       pch = 15, col = "cyan3")
points(spline(ts(pred_ars$fit, start = c(2011,2), end = c(2019,1), frequency = 12)),
       type='l', col = "cyan3")
points(ts(test$total_arsenic, start = c(2011,2), end = c(2019,1), frequency = 12),
       pch = 19, col = "darkred")

upper <- spline(ts(pred_ars$fit + pred_ars$se.fit, start = c(2011,2), end = c(2019,1), frequency = 12))
lower <- spline(ts(pred_ars$fit - pred_ars$se.fit, start = c(2011,2), end = c(2019,1), frequency = 12))
points(upper, type='l', col = "cyan4")
points(lower, type='l', col = "cyan4")

pred_ars <- predict.gam(gam_ars$gam, newdata = new_data, se.fit = TRUE)
points(2025.33, pred_ars$fit, pch=19, col='cyan3')
points(2025.33, pred_ars$fit + pred_ars$se.fit, pch=15, col='cyan4')
points(2025.33, pred_ars$fit - pred_ars$se.fit, pch=15, col='cyan4')

points(2025.33, new_data$total_arsenic, col="red", pch=19)

axis(1, at = seq(2004, 2026, by = 1), las = 2)
```

Hard to say.

# zinc
## Whole data

```{r}
zinc <- ts(data$zinc, start = c(2004,11), end = c(2019,1), frequency = 12)

fit_zinc <- arima(data$zinc, order = c(1, 1, 0),
      seasonal = list(order = c(1, 1, 0), period = 12), method = "CSS-ML")

pred_zinc <- predict(fit_zinc, n.ahead = 90)
```

```{r}
plot.ts(zinc, type='l', xlab='Temps', ylab="Concentration en zinc (ppb)",
        xlim = c(2004,2026), xaxt = 'n')

ts <- ts(pred_zinc$pred, start = c(2019,1), end = c(2026,6), frequency = 12)
ts_se <- ts(pred_zinc$se, start = c(2019,1), end = c(2026,6), frequency = 12)
x_coords <- c(time(ts), rev(time(ts)))
y_coords <- c(ts + ts_se, rev(ts - ts_se))
polygon(x_coords, y_coords, col = 'cyan3', border = "cyan")

points(ts, type='l', col='blue')
points(2025.33, new_data$zinc, col="red", pch=19)
points(ts(test$zinc, start = c(2011,2), end = c(2019,1), frequency = 12),
       pch = 19, col = "darkred")

axis(1, at = seq(2004, 2026, by = 1), las = 2)
```

Absolutely not a good prediction.

## Small data

```{r}
zinc <- ts(data_small$zinc, start = c(2004,11), end = c(2011,1), frequency = 12)

fit_zinc <- arima(zinc, order = c(1, 1, 1),
      seasonal = list(order = c(1, 1, 1), period = 12), method = "CSS-ML")

pred_zinc <- predict(fit_zinc, n.ahead = 187)
```

```{r}
plot.ts(zinc, type='l', xlab='Temps', ylab="Concentration en zinc (ppb)",
        xlim = c(2004,2026), xaxt = 'n')

ts <- ts(pred_zinc$pred, start = c(2011,1), end = c(2026,6), frequency = 12)
ts_se <- ts(pred_zinc$se, start = c(2011,1), end = c(2026,6), frequency = 12)
x_coords <- c(time(ts), rev(time(ts)))
y_coords <- c(ts + ts_se, rev(ts - ts_se))
polygon(x_coords, y_coords, col = 'cyan3', border = "cyan")

points(ts, type='l', col='blue')
points(ts(test$zinc, start = c(2011,2), end = c(2019,1), frequency = 12),
       pch = 19, col = "darkred")

points(2025.33, new_data$zinc, col='red', pch=19)

abline(h=13, col='red')
abline(h=18, col='red')

axis(1, at = seq(2004, 2026, by = 1), las = 2)
```

Better prediction but prediction interval too wide.

## GAMM and whole data

```{r}
zinc <- ts(data_small$zinc, start = c(2004,11), end = c(2019,1), frequency = 12)

plot.ts(zinc, type='l', xlab='Temps', ylab="Concentration en zinc (ppb)",
        xlim = c(2004,2026), xaxt = 'n')

lm_zinc <- lm(zinc ~ numeric_dates + lead + total_arsenic, data = data)

pred_zinc <- predict(lm_zinc, newdata = test, se.fit = TRUE)
points(ts(pred_zinc$fit, start = c(2011,2), end = c(2019,1), frequency = 12),
       pch = 15, col = "cyan3")
points(spline(ts(pred_zinc$fit, start = c(2011,2), end = c(2019,1), frequency = 12)),
       type='l', col = "cyan3")
points(ts(test$zinc, start = c(2011,2), end = c(2019,1), frequency = 12),
       pch = 19, col = "darkred")

upper <- spline(ts(pred_zinc$fit + pred_zinc$se.fit, start = c(2011,2), end = c(2019,1), frequency = 12))
lower <- spline(ts(pred_zinc$fit - pred_zinc$se.fit, start = c(2011,2), end = c(2019,1), frequency = 12))
points(upper, type='l', col = "cyan4")
points(lower, type='l', col = "cyan4")

pred_zinc <- predict(lm_zinc, newdata = new_data, se.fit = TRUE)
points(2025.33, pred_zinc$fit, pch=19, col='cyan3')
points(2025.33, pred_zinc$fit + pred_zinc$se.fit, pch=15, col='cyan4')
points(2025.33, pred_zinc$fit - pred_zinc$se.fit, pch=15, col='cyan4')

points(2025.33, new_data$zinc, col="red", pch=19)

axis(1, at = seq(2004, 2026, by = 1), las = 2)
```

Not good.

## GAMM and small data

```{r}
zinc <- ts(data_small$zinc, start = c(2004,11), end = c(2011,1), frequency = 12)

plot.ts(zinc, type='l', xlab='Temps', ylab="Concentration en zinc (ppb)",
        xlim = c(2004,2026), xaxt = 'n')

lm_zinc <- lm(zinc ~ numeric_dates + lead + total_arsenic, data = data_small)

pred_zinc <- predict(lm_zinc, newdata = test, se.fit = TRUE)
points(ts(pred_zinc$fit, start = c(2011,2), end = c(2019,1), frequency = 12),
       pch = 15, col = "cyan3")
points(spline(ts(pred_zinc$fit, start = c(2011,2), end = c(2019,1), frequency = 12)),
       type='l', col = "cyan3")
points(ts(test$zinc, start = c(2011,2), end = c(2019,1), frequency = 12),
       pch = 19, col = "darkred")

upper <- spline(ts(pred_zinc$fit + pred_zinc$se.fit, start = c(2011,2), end = c(2019,1), frequency = 12))
lower <- spline(ts(pred_zinc$fit - pred_zinc$se.fit, start = c(2011,2), end = c(2019,1), frequency = 12))
points(upper, type='l', col = "cyan4")
points(lower, type='l', col = "cyan4")

pred_zinc <- predict(lm_zinc, newdata = new_data, se.fit = TRUE)
points(2025.33, pred_zinc$fit, pch=19, col='cyan3')
points(2025.33, pred_zinc$fit + pred_zinc$se.fit, pch=15, col='cyan4')
points(2025.33, pred_zinc$fit - pred_zinc$se.fit, pch=15, col='cyan4')

points(2025.33, new_data$zinc, col="red", pch=19)

axis(1, at = seq(2004, 2026, by = 1), las = 2)
```

Nope but not too far.

# lead
## Whole data

```{r}
fer <- ts(data$lead, start = c(2004,11), end = c(2019,1), frequency = 12)

fit_fer <- arima(data$lead, order = c(1, 1, 1),
      seasonal = list(order = c(1, 1, 1), period = 12), method = "CSS-ML")

pred_fer <- predict(fit_fer, n.ahead = 90)
```

```{r}
plot.ts(fer, type='l', xlab='Temps', ylab="Concentration en lead (ppb)",
        xlim = c(2004,2026), xaxt = 'n')

ts <- ts(pred_fer$pred, start = c(2019,1), end = c(2026,6), frequency = 12)
ts_se <- ts(pred_fer$se, start = c(2019,1), end = c(2026,6), frequency = 12)
x_coords <- c(time(ts), rev(time(ts)))
y_coords <- c(ts + ts_se, rev(ts - ts_se))
polygon(x_coords, y_coords, col = 'cyan3', border = "cyan")

points(ts, type='l', col='blue')
points(2025.33, new_data$lead, col="red", pch=19)
points(ts(test$lead, start = c(2011,2), end = c(2019,1), frequency = 12),
       pch = 19, col = "darkred")

axis(1, at = seq(2004, 2026, by = 1), las = 2)
```

Absolutely not a good prediction.

## Small data

```{r}
fer <- ts(data_small$lead, start = c(2004,11), end = c(2011,1), frequency = 12)

fit_fer <- arima(fer, order = c(1, 1, 1),
      seasonal = list(order = c(1, 1, 1), period = 12), method = "CSS-ML")

pred_fer <- predict(fit_fer, n.ahead = 187)
```

```{r}
plot.ts(fer, type='l', xlab='Temps', ylab="Concentration en lead (ppb)",
        xlim = c(2004,2026), xaxt = 'n')

ts <- ts(pred_fer$pred, start = c(2011,1), end = c(2026,6), frequency = 12)
ts_se <- ts(pred_fer$se, start = c(2011,1), end = c(2026,6), frequency = 12)
x_coords <- c(time(ts), rev(time(ts)))
y_coords <- c(ts + ts_se, rev(ts - ts_se))
polygon(x_coords, y_coords, col = 'cyan3', border = "cyan")

points(ts, type='l', col='blue')
points(ts(test$lead, start = c(2011,2), end = c(2019,1), frequency = 12),
       pch = 19, col = "darkred")

points(2025.33, new_data$lead, col='red', pch=19)

abline(h=13, col='red')
abline(h=18, col='red')

axis(1, at = seq(2004, 2026, by = 1), las = 2)
```

Way better prediction.

## GAMM and whole data

```{r}
fer <- ts(data_small$lead, start = c(2004,11), end = c(2019,1), frequency = 12)

plot.ts(fer, type='l', xlab='Temps', ylab="Concentration en lead",
        xlim = c(2004,2026), xaxt = 'n')

gam_fer <- gamm(lead ~ s(rainfall_lag_1, bs='re') + s(o2, bs='re') + s(ph, bs='re') + 
               zinc + conductivity + ph:rainfall_lag_1 + conductivity:rainfall_lag_1, 
                data = data, correlation = corARMA(form = ~1, p=1))

pred_fer <- predict.gam(gam_fer$gam, newdata = test, se.fit = TRUE)
points(ts(pred_fer$fit, start = c(2011,2), end = c(2019,1), frequency = 12),
       pch = 15, col = "cyan3")
points(spline(ts(pred_fer$fit, start = c(2011,2), end = c(2019,1), frequency = 12)),
       type='l', col = "cyan3")
points(ts(test$lead, start = c(2011,2), end = c(2019,1), frequency = 12),
       pch = 19, col = "darkred")

upper <- spline(ts(pred_fer$fit + pred_fer$se.fit, start = c(2011,2), end = c(2019,1), frequency = 12))
lower <- spline(ts(pred_fer$fit - pred_fer$se.fit, start = c(2011,2), end = c(2019,1), frequency = 12))
points(upper, type='l', col = "cyan4")
points(lower, type='l', col = "cyan4")

pred_fer <- predict.gam(gam_fer$gam, newdata = new_data, se.fit = TRUE)
points(2025.33, pred_fer$fit, pch=19, col='cyan3')
points(2025.33, pred_fer$fit + pred_fer$se.fit, pch=15, col='cyan4')
points(2025.33, pred_fer$fit - pred_fer$se.fit, pch=15, col='cyan4')

points(2025.33, new_data$lead, col="red", pch=19)

axis(1, at = seq(2004, 2026, by = 1), las = 2)
```

Still not conclusive...

## GAMM and small data

```{r}
lead <- ts(data_small$lead, start = c(2004,11), end = c(2011,1), frequency = 12)

plot.ts(lead, type='l', xlab='Temps', ylab="Concentration en lead (ppb)",
        xlim = c(2004,2026), xaxt = 'n')

gam_lead <- gamm(lead ~ s(rainfall_lag_1, bs='re') + s(o2, bs='re') + s(ph, bs='re') + 
               zinc + conductivity + ph:rainfall_lag_1 + conductivity:rainfall_lag_1, 
                data = data_small, correlation = corARMA(form = ~1, p=1))

pred_lead <- predict.gam(gam_lead$gam, newdata = test, se.fit = TRUE)
points(ts(pred_lead$fit, start = c(2011,2), end = c(2019,1), frequency = 12),
       pch = 15, col = "cyan3")
points(spline(ts(pred_lead$fit, start = c(2011,2), end = c(2019,1), frequency = 12)),
       type='l', col = "cyan3")
points(ts(test$lead, start = c(2011,2), end = c(2019,1), frequency = 12),
       pch = 19, col = "darkred")

upper <- spline(ts(pred_lead$fit + pred_lead$se.fit, start = c(2011,2), end = c(2019,1), frequency = 12))
lower <- spline(ts(pred_lead$fit - pred_lead$se.fit, start = c(2011,2), end = c(2019,1), frequency = 12))
points(upper, type='l', col = "cyan4")
points(lower, type='l', col = "cyan4")

pred_lead <- predict.gam(gam_lead$gam, newdata = new_data, se.fit = TRUE)
points(2025.33, pred_lead$fit, pch=19, col='cyan3')
points(2025.33, pred_lead$fit + pred_lead$se.fit, pch=15, col='cyan4')
points(2025.33, pred_lead$fit - pred_lead$se.fit, pch=15, col='cyan4')

points(2025.33, new_data$lead, col="red", pch=19)

axis(1, at = seq(2004, 2026, by = 1), las = 2)
```

Almost!

# feII
## Whole data

```{r}
fer <- ts(data$feII, start = c(2004,11), end = c(2019,1), frequency = 12)

fit_fer <- arima(data$feII, order = c(1, 1, 0),
      seasonal = list(order = c(1, 1, 0), period = 12), method = "CSS-ML")

pred_fer <- predict(fit_fer, n.ahead = 90)
```

```{r}
plot.ts(fer, type='l', xlab='Temps', ylab="Concentration en feII (ppb)",
        xlim = c(2004,2026), xaxt = 'n')

ts <- ts(pred_fer$pred, start = c(2019,1), end = c(2026,6), frequency = 12)
ts_se <- ts(pred_fer$se, start = c(2019,1), end = c(2026,6), frequency = 12)
x_coords <- c(time(ts), rev(time(ts)))
y_coords <- c(ts + ts_se, rev(ts - ts_se))
polygon(x_coords, y_coords, col = 'cyan3', border = "cyan")

points(ts, type='l', col='blue')
points(2025.33, new_data$feII, col="red", pch=19)
points(ts(test$feII, start = c(2011,2), end = c(2019,1), frequency = 12),
       pch = 19, col = "darkred")

axis(1, at = seq(2004, 2026, by = 1), las = 2)
```

Good prediction but not precise enough.

## Small data

```{r}
fer <- ts(data_small$feII, start = c(2004,11), end = c(2011,1), frequency = 12)

fit_fer <- arima(fer, order = c(1, 1, 1),
      seasonal = list(order = c(1, 1, 1), period = 12), method = "CSS-ML")

pred_fer <- predict(fit_fer, n.ahead = 187)
```

```{r}
plot.ts(fer, type='l', xlab='Temps', ylab="Concentration en feII (ppb)",
        xlim = c(2004,2026), xaxt = 'n')

ts <- ts(pred_fer$pred, start = c(2011,1), end = c(2026,6), frequency = 12)
ts_se <- ts(pred_fer$se, start = c(2011,1), end = c(2026,6), frequency = 12)
x_coords <- c(time(ts), rev(time(ts)))
y_coords <- c(ts + ts_se, rev(ts - ts_se))
polygon(x_coords, y_coords, col = 'cyan3', border = "cyan")

points(ts, type='l', col='blue')
points(ts(test$feII, start = c(2011,2), end = c(2019,1), frequency = 12),
       pch = 19, col = "darkred")

points(2025.33, new_data$feII, col='red', pch=19)

abline(h=13, col='red')
abline(h=18, col='red')

axis(1, at = seq(2004, 2026, by = 1), las = 2)
```

Way better prediction!

## GAMM and whole data

```{r}
fer <- ts(data_small$feII, start = c(2004,11), end = c(2019,1), frequency = 12)

plot.ts(fer, type='l', xlab='Temps', ylab="Concentration en feII",
        xlim = c(2004,2026), xaxt = 'n')

gam_fer <- gamm(feII ~ s(rainfall, bs='re') + s(o2, bs='re') + s(sulfate, bs='re') +
               numeric_dates + ph + conductivity + lead + sqrt(total_arsenic) +
               lead:conductivity + ph:conductivity, 
                data = data, correlation = corARMA(form = ~1, p=1))

pred_fer <- predict.gam(gam_fer$gam, newdata = test, se.fit = TRUE)
points(ts(pred_fer$fit, start = c(2011,2), end = c(2019,1), frequency = 12),
       pch = 15, col = "cyan3")
points(spline(ts(pred_fer$fit, start = c(2011,2), end = c(2019,1), frequency = 12)),
       type='l', col = "cyan3")
points(ts(test$feII, start = c(2011,2), end = c(2019,1), frequency = 12),
       pch = 19, col = "darkred")

upper <- spline(ts(pred_fer$fit + pred_fer$se.fit, start = c(2011,2), end = c(2019,1), frequency = 12))
lower <- spline(ts(pred_fer$fit - pred_fer$se.fit, start = c(2011,2), end = c(2019,1), frequency = 12))
points(upper, type='l', col = "cyan4")
points(lower, type='l', col = "cyan4")

pred_fer <- predict.gam(gam_fer$gam, newdata = new_data, se.fit = TRUE)
points(2025.33, pred_fer$fit, pch=19, col='cyan3')
points(2025.33, pred_fer$fit + pred_fer$se.fit, pch=15, col='cyan4')
points(2025.33, pred_fer$fit - pred_fer$se.fit, pch=15, col='cyan4')

points(2025.33, new_data$feII, col="red", pch=19)

axis(1, at = seq(2004, 2026, by = 1), las = 2)
```

Nope.

## GAMM and small data

```{r}
fer <- ts(data_small$feII, start = c(2004,11), end = c(2011,1), frequency = 12)

plot.ts(fer, type='l', xlab='Temps', ylab="Concentration en feII (ppb)",
        xlim = c(2004,2026), xaxt = 'n')

gam_fer <- gamm(feII ~ s(rainfall, bs='re') + s(o2, bs='re') + s(sulfate, bs='re') +
               numeric_dates + ph + conductivity + lead + sqrt(total_arsenic) +
               lead:conductivity + ph:conductivity, 
                data = data_small, correlation = corARMA(form = ~1, p=1))

pred_fer <- predict.gam(gam_fer$gam, newdata = test, se.fit = TRUE)
points(ts(pred_fer$fit, start = c(2011,2), end = c(2019,1), frequency = 12),
       pch = 15, col = "cyan3")
points(spline(ts(pred_fer$fit, start = c(2011,2), end = c(2019,1), frequency = 12)),
       type='l', col = "cyan3")
points(ts(test$feII, start = c(2011,2), end = c(2019,1), frequency = 12),
       pch = 19, col = "darkred")

upper <- spline(ts(pred_fer$fit + pred_fer$se.fit, start = c(2011,2), end = c(2019,1), frequency = 12))
lower <- spline(ts(pred_fer$fit - pred_fer$se.fit, start = c(2011,2), end = c(2019,1), frequency = 12))
points(upper, type='l', col = "cyan4")
points(lower, type='l', col = "cyan4")

pred_fer <- predict.gam(gam_fer$gam, newdata = new_data, se.fit = TRUE)
points(2025.33, pred_fer$fit, pch=19, col='cyan3')
points(2025.33, pred_fer$fit + pred_fer$se.fit, pch=15, col='cyan4')
points(2025.33, pred_fer$fit - pred_fer$se.fit, pch=15, col='cyan4')

points(2025.33, new_data$feII, col="red", pch=19)

axis(1, at = seq(2004, 2026, by = 1), las = 2)
```

Nope.
