---
title: "Sensitivity Analysis with COWG's data"
author: "CLETZ Laura"
date: "03/07/2025"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Packages needed:
```{r}
library(readr)
library(mgcv)
library(ggplot2)
library(dplyr)
library(scales)
library(sensitivity)
here::i_am("Correlations_and_Predictions/sensitivity_analysis_COWG.Rmd")
library(here)
```

Step 1:
```{r}
scaled_data <- data_diff %>%
  mutate_at(3:17, rescale)

model <- gamm(Arsenic_Total ~ s(Rainfall_lag_1, bs='re') + s(O2, bs='re') + s(Temperature, bs='re') 
              + s(Cuivre, bs='re') + s(Conductivité, bs='re') + s(Zinc, bs='re') + s(Ion_Ferreux, bs='re'),
              correlation = corARMA(form = ~ 1, p = 1, q = 1), 
              data = scaled_data, na.action = na.exclude,
              control = lmeControl(msMaxIter = 1000, msMaxEval = 1000))
```

```{r}
predict_model <- function(x) {
  new_data <- data.frame(
    Rainfall_lag_1 = x[, 1],
    O2 = x[, 2],
    Temperature = x[, 3], 
    Cuivre = x[, 4],
    Conductivité = x[, 5],
    Zinc = x[, 6],
    Ion_Ferreux = x[, 7]
  )
  predict(model$gam, newdata = new_data, type = "response")
}
```

```{r, warning=FALSE}
n <- 1000
X1 <- data.frame(matrix(runif(7 * n), nrow = n))
X2 <- data.frame(matrix(runif(7 * n), nrow = n))

sa_result <- sobol2002(
  model = predict_model,
  X1 = X1,
  X2 = X2,
  nboot = 100
)

var_labels <- c("X1" = "Pluie à t-1", "X2" = "O2", "X3" = "Température", "X4" = "Cuivre", 
  "X5" = "Conductivité", "X6" = "Zinc", "X7" = "Ion ferreux")

df_plot <- bind_rows(
  sa_result$S %>%
    mutate(Variable = rownames(.), Index = "First-order") %>%
    rename(y = original, ymin = `min. c.i.`, ymax = `max. c.i.`),
  sa_result$T %>%
    mutate(Variable = rownames(.), Index = "Total-order") %>%
    rename(y = original, ymin = `min. c.i.`, ymax = `max. c.i.`)
) %>%
  mutate(
    Variable = factor(Variable, levels = names(var_labels)),
    Label = var_labels[as.character(Variable)]
  )

ggplot(df_plot, aes(x = Label, y = y, ymin = ymin, ymax = ymax, shape = Index)) +
  geom_pointrange(position = position_dodge(width = 0.4), size = 0.6) +
  scale_shape_manual(values = c("First-order" = 16, "Total-order" = 1),
  labels = c("First-order" = "Indice d'ordre 1", "Total-order" = "Indice total")
  ) +
  labs(
    x = "",
    y = "Indice de sensibilité",
    shape = ""
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1),
    legend.position = "right"
  ) +
  coord_cartesian(ylim = c(0, 1))
```

Step 2:
```{r}
model <- gamm(Arsenic_Total ~ s(Rainfall_lag_1, bs='re') + s(O2, bs='re') + s(Temperature, bs='re') 
              + s(Cuivre, bs='re') + s(Conductivité, bs='re') + s(Zinc, bs='re'),
              correlation = corARMA(form = ~ 1, p = 1, q = 1), 
              data = scaled_data, na.action = na.exclude,
              control = lmeControl(msMaxIter = 1000, msMaxEval = 1000))
```

```{r}
predict_model <- function(x) {
  new_data <- data.frame(
    Rainfall_lag_1 = x[, 1],
    O2 = x[, 2],
    Temperature = x[, 3], 
    Cuivre = x[, 4],
    Conductivité = x[, 5],
    Zinc = x[, 6]
  )
  predict(model$gam, newdata = new_data, type = "response")
}
```

```{r, warning=FALSE}
n <- 1000
X1 <- data.frame(matrix(runif(6 * n), nrow = n))
X2 <- data.frame(matrix(runif(6 * n), nrow = n))

sa_result <- sobol2002(
  model = predict_model,
  X1 = X1,
  X2 = X2,
  nboot = 100
)

var_labels <- c("X1" = "Pluie à t-1", "X2" = "O2", "X3" = "Température", "X4" = "Cuivre", 
  "X5" = "Conductivité", "X6" = "Zinc")

df_plot <- bind_rows(
  sa_result$S %>%
    mutate(Variable = rownames(.), Index = "First-order") %>%
    rename(y = original, ymin = `min. c.i.`, ymax = `max. c.i.`),
  sa_result$T %>%
    mutate(Variable = rownames(.), Index = "Total-order") %>%
    rename(y = original, ymin = `min. c.i.`, ymax = `max. c.i.`)
) %>%
  mutate(
    Variable = factor(Variable, levels = names(var_labels)),
    Label = var_labels[as.character(Variable)]
  )

ggplot(df_plot, aes(x = Label, y = y, ymin = ymin, ymax = ymax, shape = Index)) +
  geom_pointrange(position = position_dodge(width = 0.4), size = 0.6) +
  scale_shape_manual(values = c("First-order" = 16, "Total-order" = 1),
  labels = c("First-order" = "Indice d'ordre 1", "Total-order" = "Indice total")
  ) +
  labs(
    x = "",
    y = "Indice de sensibilité",
    shape = ""
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1),
    legend.position = "right"
  ) +
  coord_cartesian(ylim = c(0, 1))
```

```{r}
model <- gamm(Arsenic_Total ~ s(Rainfall_lag_1, bs='re') + s(O2, bs='re') + s(Temperature, bs='re') 
              + s(Cuivre, bs='re') + s(Conductivité, bs='re'),
              correlation = corARMA(form = ~ 1, p = 1, q = 1), 
              data = scaled_data, na.action = na.exclude)
```

```{r}
predict_model <- function(x) {
  new_data <- data.frame(
    Rainfall_lag_1 = x[, 1],
    O2 = x[, 2],
    Temperature = x[, 3], 
    Cuivre = x[, 4],
    Conductivité = x[, 5]
  )
  predict(model$gam, newdata = new_data, type = "response")
}
```

```{r, warning=FALSE}
n <- 1000
X1 <- data.frame(matrix(runif(5 * n), nrow = n))
X2 <- data.frame(matrix(runif(5 * n), nrow = n))

sa_result <- sobol2002(
  model = predict_model,
  X1 = X1,
  X2 = X2,
  nboot = 100
)

var_labels <- c("X1" = "Pluie à t-1", "X2" = "O2", "X3" = "Température", "X4" = "Cuivre", 
  "X5" = "Conductivité")

df_plot <- bind_rows(
  sa_result$S %>%
    mutate(Variable = rownames(.), Index = "First-order") %>%
    rename(y = original, ymin = `min. c.i.`, ymax = `max. c.i.`),
  sa_result$T %>%
    mutate(Variable = rownames(.), Index = "Total-order") %>%
    rename(y = original, ymin = `min. c.i.`, ymax = `max. c.i.`)
) %>%
  mutate(
    Variable = factor(Variable, levels = names(var_labels)),
    Label = var_labels[as.character(Variable)]
  )

ggplot(df_plot, aes(x = Label, y = y, ymin = ymin, ymax = ymax, shape = Index)) +
  geom_pointrange(position = position_dodge(width = 0.4), size = 0.6) +
  scale_shape_manual(values = c("First-order" = 16, "Total-order" = 1),
  labels = c("First-order" = "Indice d'ordre 1", "Total-order" = "Indice total")
  ) +
  labs(
    x = "",
    y = "Indice de sensibilité",
    shape = ""
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1),
    legend.position = "right"
  ) +
  coord_cartesian(ylim = c(0, 1))
```
A higher value of S (First-Order Indices) indicates that the input variable has a significant individual effect on the output variance.

A higher value of T (Total-Order Indices) indicates that the input variable, including its interactions with other variables, has a significant effect on the output variance.

Too much variances caught, must be removed: Zinc, Ion_Ferreux.
Equal high variances caught: Cuivre, O2, Rainfall_lag_1.

Let's compare the three consecutive results with the smaller set of data.
Step 1:
```{r}
scaled_data_small <- data_small %>%
  mutate_at(3:17, rescale)

model_small <- gamm(Arsenic_Total ~ s(Rainfall_lag_1, bs='re') + s(O2, bs='re') + s(Temperature, bs='re') 
              + s(Cuivre, bs='re') + s(Conductivité, bs='re') + s(Zinc, bs='re') + s(Ion_Ferreux, bs='re'),
              correlation = corARMA(form = ~ 1, p = 1, q = 1), 
              data = scaled_data_small, na.action = na.exclude,
              control = lmeControl(msMaxIter = 1000, msMaxEval = 1000))
```

```{r}
predict_model_small <- function(x) {
  new_data <- data.frame(
    Rainfall_lag_1 = x[, 1],
    O2 = x[, 2],
    Temperature = x[, 3], 
    Cuivre = x[, 4],
    Conductivité = x[, 5],
    Zinc = x[, 6],
    Ion_Ferreux = x[, 7]
  )
  predict(model_small$gam, newdata = new_data, type = "response")
}
```

```{r, warning=FALSE}
n <- 1000
X1 <- data.frame(matrix(runif(7 * n), nrow = n))
X2 <- data.frame(matrix(runif(7 * n), nrow = n))

sa_result_small <- sobol2002(
  model = predict_model_small,
  X1 = X1,
  X2 = X2,
  nboot = 100
)

var_labels <- c("X1" = "Pluie à t-1", "X2" = "O2", "X3" = "Température", "X4" = "Cuivre", 
  "X5" = "Conductivité", "X6" = "Zinc", "X7" = "Ion ferreux")

df_plot <- bind_rows(
  sa_result_small$S %>%
    mutate(Variable = rownames(.), Index = "First-order") %>%
    rename(y = original, ymin = `min. c.i.`, ymax = `max. c.i.`),
  sa_result_small$T %>%
    mutate(Variable = rownames(.), Index = "Total-order") %>%
    rename(y = original, ymin = `min. c.i.`, ymax = `max. c.i.`)
) %>%
  mutate(
    Variable = factor(Variable, levels = names(var_labels)),
    Label = var_labels[as.character(Variable)]
  )

ggplot(df_plot, aes(x = Label, y = y, ymin = ymin, ymax = ymax, shape = Index)) +
  geom_pointrange(position = position_dodge(width = 0.4), size = 0.6) +
  scale_shape_manual(values = c("First-order" = 16, "Total-order" = 1),
  labels = c("First-order" = "Indice d'ordre 1", "Total-order" = "Indice total")
  ) +
  labs(
    x = "",
    y = "Indice de sensibilité",
    shape = ""
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1),
    legend.position = "right"
  ) +
  coord_cartesian(ylim = c(0, 1))
```

Step 2:
```{r}
model_small <- gamm(Arsenic_Total ~ s(Rainfall_lag_1, bs='re') + s(O2, bs='re') + s(Temperature, bs='re') 
              + s(Cuivre, bs='re') + s(Conductivité, bs='re') + s(Ion_Ferreux, bs='re'),
              correlation = corARMA(form = ~ 1, p = 1, q = 1), 
              data = scaled_data_small, na.action = na.exclude,
              control = lmeControl(msMaxIter = 1000, msMaxEval = 1000))
```

```{r}
predict_model_small <- function(x) {
  new_data <- data.frame(
    Rainfall_lag_1 = x[, 1],
    O2 = x[, 2],
    Temperature = x[, 3], 
    Cuivre = x[, 4],
    Conductivité = x[, 5],
    Ion_Ferreux = x[, 6]
  )
  predict(model_small$gam, newdata = new_data, type = "response")
}
```

```{r, warning=FALSE}
n <- 1000
X1 <- data.frame(matrix(runif(6 * n), nrow = n))
X2 <- data.frame(matrix(runif(6 * n), nrow = n))

sa_result_small <- sobol2002(
  model = predict_model_small,
  X1 = X1,
  X2 = X2,
  nboot = 100
)

var_labels <- c("X1" = "Pluie à t-1", "X2" = "O2", "X3" = "Température", "X4" = "Cuivre", 
  "X5" = "Conductivité", "X6" = "Ion ferreux")

df_plot <- bind_rows(
  sa_result_small$S %>%
    mutate(Variable = rownames(.), Index = "First-order") %>%
    rename(y = original, ymin = `min. c.i.`, ymax = `max. c.i.`),
  sa_result_small$T %>%
    mutate(Variable = rownames(.), Index = "Total-order") %>%
    rename(y = original, ymin = `min. c.i.`, ymax = `max. c.i.`)
) %>%
  mutate(
    Variable = factor(Variable, levels = names(var_labels)),
    Label = var_labels[as.character(Variable)]
  )

ggplot(df_plot, aes(x = Label, y = y, ymin = ymin, ymax = ymax, shape = Index)) +
  geom_pointrange(position = position_dodge(width = 0.4), size = 0.6) +
  scale_shape_manual(values = c("First-order" = 16, "Total-order" = 1),
  labels = c("First-order" = "Indice d'ordre 1", "Total-order" = "Indice total")
  ) +
  labs(
    x = "",
    y = "Indice de sensibilité",
    shape = ""
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1),
    legend.position = "right"
  ) +
  coord_cartesian(ylim = c(0, 1))
```

Step 3:
```{r}
model_small <- gamm(Arsenic_Total ~ s(Rainfall_lag_1, bs='re') + s(O2, bs='re') + s(Temperature, bs='re') 
              + s(Cuivre, bs='re') + s(Conductivité, bs='re'),
              correlation = corARMA(form = ~ 1, p = 1, q = 1), 
              data = scaled_data_small, na.action = na.exclude,
              control = lmeControl(msMaxIter = 1000, msMaxEval = 1000))
```

```{r}
predict_model_small <- function(x) {
  new_data <- data.frame(
    Rainfall_lag_1 = x[, 1],
    O2 = x[, 2],
    Temperature = x[, 3], 
    Cuivre = x[, 4],
    Conductivité = x[, 5]
  )
  predict(model_small$gam, newdata = new_data, type = "response")
}
```

```{r, warning=FALSE}
n <- 1000
X1 <- data.frame(matrix(runif(5 * n), nrow = n))
X2 <- data.frame(matrix(runif(5 * n), nrow = n))

sa_result_small <- sobol2002(
  model = predict_model_small,
  X1 = X1,
  X2 = X2,
  nboot = 100
)

var_labels <- c("X1" = "Pluie à t-1", "X2" = "O2", "X3" = "Température", "X4" = "Cuivre", 
  "X5" = "Conductivité")

df_plot <- bind_rows(
  sa_result_small$S %>%
    mutate(Variable = rownames(.), Index = "First-order") %>%
    rename(y = original, ymin = `min. c.i.`, ymax = `max. c.i.`),
  sa_result_small$T %>%
    mutate(Variable = rownames(.), Index = "Total-order") %>%
    rename(y = original, ymin = `min. c.i.`, ymax = `max. c.i.`)
) %>%
  mutate(
    Variable = factor(Variable, levels = names(var_labels)),
    Label = var_labels[as.character(Variable)]
  )

ggplot(df_plot, aes(x = Label, y = y, ymin = ymin, ymax = ymax, shape = Index)) +
  geom_pointrange(position = position_dodge(width = 0.4), size = 0.6) +
  scale_shape_manual(values = c("First-order" = 16, "Total-order" = 1),
  labels = c("First-order" = "Indice d'ordre 1", "Total-order" = "Indice total")
  ) +
  labs(
    x = "",
    y = "Indice de sensibilité",
    shape = ""
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1),
    legend.position = "right"
  ) +
  coord_cartesian(ylim = c(0, 1))
```

Equal high variances caught: Copper, Conductivity.
Too much variances caught, must be removed: Zinc.
A little too much variance caught, must be removed: FeII.
