---
title: "Sensitivity Analysis"
author: "CLETZ Laura"
date: "30/06/2025"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Packages needed:
```{r}
library(readr)
library(mgcv)
library(rlang)
library(ggplot2)
library(Hmisc)
library(GGally)
library(dplyr)
library(FactoMineR)
library(factoextra)
here::i_am("Correlations_and_Predictions/sensitivity_analysis.Rmd")
library(here)
```

```{r}
data_small <- read.table(here("processed", "COWG_gamm_2004_2011.csv"), sep=";", dec=".", header=T) %>% 
  mutate_at(18:23, as.factor)
data_small$Dates <- as.Date(data_small$Dates, format = "%Y-%m-%d")
data_small$Corrected_Dates <- as.Date(data_small$Corrected_Dates, format = "%Y-%m-%d")
```

```{r}
data_big <- read.table(here("processed", "COWG_gamm_2004_2018.csv"), sep=";", dec=".", header=T) %>% 
  mutate_at(18:23, as.factor)
data_big$Dates <- as.Date(data_big$Dates, format = "%Y-%m-%d")
data_big$Corrected_Dates <- as.Date(data_big$Corrected_Dates, format = "%Y-%m-%d")
```

# Correlations table

```{r}
custom_gam_plot <- function(data, mapping, ...) {
  
  x_var <- rlang::as_label(mapping$x)
  y_var <- rlang::as_label(mapping$y)

  formula <- substitute( y ~ s(x, bs='cr'), list(y = as.name(y_var), x = as.name(x_var)) )
  
  gam_model <- gamm(formula, correlation = corARMA(form = ~1, p=1), data = data, na.action = na.exclude)

  pred_data <- data.frame(x = seq(min(data[[x_var]], na.rm = TRUE),
                                    max(data[[x_var]], na.rm = TRUE),
                                    length.out = 100))
  names(pred_data) <- x_var
  
  pred_data$y <- predict(gam_model$gam, newdata = pred_data)

  ggplot(data, mapping = mapping) +
    geom_point() +
    geom_line(data = pred_data, 
              mapping = aes_string(x = x_var, y = "y"), 
              color = "blue") +
    labs(title = paste0("GAM fit:", y_var, "~ s(", x_var, ")"), 
         x = x_var, y = y_var) +
    theme_minimal()
}
```

```{r, warning=FALSE, echo=FALSE}
g <- ggpairs(
  data = data_small[, c(3:23)],
  lower = list(continuous = custom_gam_plot, combo = 'facetdensity'),
  upper = list(continuous = wrap('cor', size = 3, hjust = 0.8)),
  progress = FALSE
)
```

```{r, warning=FALSE, echo=FALSE}
g
```

Rainfall and Dates:
```{r}
ggplot(data_small, aes(x = Numeric_Dates)) +
  geom_point(aes(y = Rainfall), col = "darkred") +
  geom_smooth(aes(y = Rainfall), method = 'gam', formula = y ~ s(x, bs='re'), col = "red") +
  geom_point(aes(y = Rainfall_lag_1), col = "darkgreen") +
  geom_smooth(aes(y = Rainfall_lag_1), method = 'gam', formula = y ~ s(x, bs='re'), col = "green") +
  theme_minimal()
```

Rainfall at one lag and O2:
```{r}
ggplot(data_small, aes(x = Rainfall_lag_1, y = O2, colour = O2_method)) +
  geom_point() +
  geom_smooth(method = 'gam', formula = y ~ s(x, bs='re')) +
  theme_minimal()
```

Dates and O2:
```{r}
ggplot(data_small, aes(x = Numeric_Dates, y = O2, colour = O2_method)) +
  geom_point() +
  geom_smooth(method = 'gam', formula = y ~ s(x, bs='re')) +
  theme_minimal()
```

Rainfall at one lag and Copper:
```{r}
ggplot(data_small, aes(x = Rainfall_lag_1, y = Cuivre, colour = Cuivre_method)) +
  geom_point() +
  geom_smooth(method = 'gam', formula = y ~ s(x, bs='re')) +
  theme_minimal()
```

FeII and Zinc:
```{r}
ggplot(data_small, aes(x = Ion_Ferreux, y = Zinc, colour = Zinc_method)) +
  geom_point() +
  geom_smooth(method = 'gam', formula = y ~ s(x, bs='re')) +
  theme_minimal()
```

Arsenic and O2:
```{r}
ggplot(data_small, aes(x = Arsenic_Total, y = O2, colour = O2_method)) +
  geom_point() +
  geom_smooth(method = 'gam', formula = y ~ s(x, bs='re')) +
  theme_minimal()
```

Lead and Copper:
```{r}
ggplot(data_small, aes(x = Cuivre, y = Plomb, colour = Plomb_method)) +
  geom_point() +
  geom_smooth(method = 'gam', formula = y ~ s(x, bs='re')) +
  theme_minimal()
```

Measurement methods seem to add some precision at predicting confidence intervals in some case only. There are methods like Colorimétrie for O2 that don't help at all, giving straight lines instead of smooth curves. We did not remove NAs because they are too many to avoid, we need to count them in.

# Variances comparison

```{r}
library(car)
leveneTest(data_small$Rainfall, group = data_small$O2_method, data = data_small)
leveneTest(data_small$Conductivité, group = data_small$Plomb_method, data = data_small)
leveneTest(data_big$Rainfall, group = data_big$O2_method, data = data_big)
leveneTest(data_big$Conductivité, group = data_big$Plomb_method, data = data_big)
```
Variances not homogenous.

```{r}
data_diff <- data_big %>%
  anti_join(data_small, by = "Corrected_Dates")

sd(data_big$Arsenic_Total)
sd(data_diff$Arsenic_Total)
sd(data_small$Arsenic_Total)
```

# Par période (nombre d'imputation)

```{r}
Period <- c( rep("2004-2009", 47), rep("2009-2012", 87-47), rep("2012-2019", 174-87), rep("2019-2022", 213-174))
data_big <- data_big %>% bind_cols(Period)
colnames(data_big)[24] <- "Period"
data_big$Period <- as.factor(data_big$Period)
```

```{r}
var.test(data_small$Temperature, data_diff$Temperature)
var.test(data_small$Antimoine, data_diff$Antimoine)
var.test(data_small$O2, data_diff$O2)
var.test(data_small$Arsenic_Total, data_diff$Arsenic_Total)
```
Variances between 2004-2009 and 2009-2022 are different.

```{r, warning=FALSE}
ggplot(data_big, aes(x = Period, y = Arsenic_Total, colour = Period)) +
  geom_violin() + 
  geom_jitter(shape=16, position=position_jitter(0.2)) +
  theme_minimal() +
  stat_summary(fun.data=mean_sdl, mult=1, 
                 geom="pointrange", color="red")
```

```{r, warning=FALSE}
ggplot(data_big, aes(x = Period, y = Rainfall_lag_1, colour = Period)) +
  geom_violin() + 
  geom_jitter(shape=16, position=position_jitter(0.2)) +
  theme_minimal() +
  stat_summary(fun.data=mean_sdl, mult=1, 
                 geom="pointrange", color="red")
```

```{r, warning=FALSE}
ggplot(data_big, aes(x = Period, y = O2, colour = O2_method)) +
  geom_violin() + 
  geom_jitter(shape=16, position=position_jitter(0.2)) +
  theme_minimal() +
  stat_summary(fun.data=mean_sdl, mult=1, 
                 geom="pointrange", color="red")
```

```{r, warning=FALSE}
ggplot(data_big, aes(x = Period, y = Cuivre, colour = Cuivre_method)) +
  geom_violin() + 
  geom_jitter(shape=16, position=position_jitter(0.2)) +
  theme_minimal() +
  stat_summary(fun.data=mean_sdl, mult=1, 
                 geom="pointrange", color="red")
```

# Factor Analysis of Mixed Data (FAMD)

```{r}
levels(data_diff$O2_method) <- c(levels(data_diff$O2_method), "Undetermined_O2_method")
data_diff$O2_method[ is.na( data_diff$O2_method ) ] <- "Undetermined_O2_method"

levels(data_diff$Cadmium_method) <- c(levels(data_diff$Cadmium_method), "Undetermined_Cadmium_method")
data_diff$Cadmium_method[ is.na( data_diff$Cadmium_method ) ] <- "Undetermined_Cadmium_method"

levels(data_diff$Antimoine_method) <- c(levels(data_diff$Antimoine_method), "Undetermined_Antimoine_method")
data_diff$Antimoine_method[ is.na( data_diff$Antimoine_method ) ] <- "Undetermined_Antimoine_method"

levels(data_diff$Cuivre_method) <- c(levels(data_diff$Cuivre_method), "Undetermined_Cuivre_method")
data_diff$Cuivre_method[ is.na( data_diff$Cuivre_method ) ] <- "Undetermined_Cuivre_method"

levels(data_diff$Plomb_method) <- c(levels(data_diff$Plomb_method), "Undetermined_Plomb_method")
data_diff$Plomb_method[ is.na( data_diff$Plomb_method ) ] <- "Undetermined_Plomb_method"

levels(data_diff$Zinc_method) <- c(levels(data_diff$Zinc_method), "Undetermined_Zinc_method")
data_diff$Zinc_method[ is.na( data_diff$Zinc_method ) ] <- "Undetermined_Zinc_method"
```

```{r}
Period <- c( rep("2009-2012", 8), rep("2012-2015", 43-8), rep("2015-2019", 88-43), rep("2019-2022", 127-88) )
data_diff <- data_diff %>% bind_cols(Period)
colnames(data_diff)[24] <- "Period"
data_diff$Period <- as.factor(data_diff$Period)
```

```{r}
res.famd <- FAMD(data_diff[,3:24], graph = FALSE, ncp = Inf)
res.famd$eig
```
```{r}
var_names <- rownames(res.famd$var$coord)
custom_groups <- ifelse(var_names %in% var_names[c(1:7, 22)],
                        "Éléments physico-chimiques et temps",
                        ifelse(var_names %in% var_names[8:15],
                        "Éléments traces (métalloïdes)",
                        "Méthodes de mesure"))

fviz_famd_var(res.famd, 
             axes = c(1,2), 
             repel = TRUE, 
             col.var = custom_groups,  
             palette = c("blue", "red", "darkgreen"),
             ggtheme = theme_minimal(),
             title = "FAMD Variables Plot")

fviz_famd_ind(res.famd,
             axes = c(1, 2),
             repel = TRUE,
             ggtheme = theme_minimal(),
             col.ind = "cos2",
             label = "quali.var",
             title = "FAMD Individuals Plot")
```
```{r}
res.famd <- FAMD(data_diff[,3:24], graph = FALSE, sup.var = c(17:23))
res.famd$eig
```

```{r}
var_names <- rownames(res.famd$var$coord)
custom_groups <- ifelse(var_names %in% var_names[c(1:7, 22)],
                        "Éléments physico-chimiques et temps",
                        "Éléments traces (métalloïdes)")

fviz_famd_var(res.famd, 
             axes = c(1, 2), 
             repel = TRUE, 
             col.var = custom_groups,  
             palette = c("blue", "red"),
             col.var.sup = "darkgreen",
             ggtheme = theme_minimal(),
             title = "FAMD Variables Plot")

fviz_famd_ind(res.famd,
             axes = c(1, 2),
             repel = TRUE,
             ggtheme = theme_minimal(),
             col.ind = "cos2",
             label = "none",
             title = "FAMD Individuals Plot")
```

```{r}
fviz_famd_var(res.famd, 
             axes = c(1,3), 
             repel = TRUE, 
             col.var = custom_groups,  
             palette = c("blue", "red"),
             col.var.sup = "darkgreen",
             ggtheme = theme_minimal(),
             title = "FAMD Variables Plot")

fviz_famd_ind(res.famd,
             axes = c(1,3),
             repel = TRUE,
             ggtheme = theme_minimal(),
             col.ind = "cos2",
             label = "none",
             title = "FAMD Individuals Plot")
```

```{r}
dimdesc(res.famd)
```

```{r}
test <- read.table(here("processed","COWG.csv"), sep=";", dec=".", header=T)

test <- test[which(colSums(is.na(test[(1:86),]))*100/nrow(test[(1:86),])<=35)] %>%
  mutate_at(1:2, function(x) as.Date(x, format="%Y-%m-%d")) %>%
  factor_processing() %>%
  select(names(data_diff[,-c(5,24)]))
```

```{r}
na_index <- unique(unlist(
  lapply(test[, 5:16], function(col) which(is.na(col)))
))
```

```{r}
#res.famd <- FAMD(data_diff[,3:24], graph = FALSE, sup.var = c(17:23), ind.sup = na_index)
#res.famd$eig
#It seems that we can't run a FAMD with 152 supplementary rows...
```

```{r}
var_names <- rownames(res.famd$var$coord)
custom_groups <- ifelse(var_names %in% var_names[c(1:7, 22)],
                        "Éléments physico-chimiques et temps",
                        "Éléments traces (métalloïdes)")

fviz_famd_var(res.famd, 
             axes = c(1,2), 
             repel = TRUE, 
             col.var = custom_groups,  
             palette = c("blue", "red"),
             col.var.sup = "darkgreen",
             ggtheme = theme_minimal(),
             title = "FAMD Variables Plot")

fviz_famd_ind(res.famd,
             axes = c(1, 2),
             repel = TRUE,
             ggtheme = theme_minimal(),
             col.ind = "cos2",
             label = "none",
             title = "FAMD Individuals Plot")
```

```{r}
cd <- catdes(data_diff[,3:24], 18)
plot.catdes(cd)
```

# Sensitivity Analysis

Step 1:
```{r}
library(scales)
scaled_data <- data_diff %>%
  mutate_at(3:17, rescale)

model <- gamm(Arsenic_Total ~ s(Rainfall_lag_1, bs='re') + s(O2, bs='re') + s(Temperature, bs='re') 
              + s(Cuivre, bs='re') + s(Conductivité, bs='re') + s(Zinc, bs='re') + s(Ion_Ferreux, bs='re'),
              correlation = corARMA(form = ~ 1, p = 1, q = 1), 
              data = scaled_data, na.action = na.exclude,
              control = lmeControl(msMaxIter = 1000, msMaxEval = 1000))
```

```{r}
predict_model <- function(x) {
  new_data <- data.frame(
    Rainfall_lag_1 = x[, 1],
    O2 = x[, 2],
    Temperature = x[, 3], 
    Cuivre = x[, 4],
    Conductivité = x[, 5],
    Zinc = x[, 6],
    Ion_Ferreux = x[, 7]
  )
  predict(model$gam, newdata = new_data, type = "response")
}
```

```{r, warning=FALSE}
n <- 1000
X1 <- data.frame(matrix(runif(7 * n), nrow = n))
X2 <- data.frame(matrix(runif(7 * n), nrow = n))

library(sensitivity)
sa_result <- sobol2002(
  model = predict_model,
  X1 = X1,
  X2 = X2,
  nboot = 100
)

var_labels <- c("X1" = "Pluie à t-1", "X2" = "O2", "X3" = "Température", "X4" = "Cuivre", 
  "X5" = "Conductivité", "X6" = "Zinc", "X7" = "Ion ferreux")

df_plot <- bind_rows(
  sa_result$S %>%
    mutate(Variable = rownames(.), Index = "First-order") %>%
    rename(y = original, ymin = `min. c.i.`, ymax = `max. c.i.`),
  sa_result$T %>%
    mutate(Variable = rownames(.), Index = "Total-order") %>%
    rename(y = original, ymin = `min. c.i.`, ymax = `max. c.i.`)
) %>%
  mutate(
    Variable = factor(Variable, levels = names(var_labels)),
    Label = var_labels[as.character(Variable)]
  )

ggplot(df_plot, aes(x = Label, y = y, ymin = ymin, ymax = ymax, shape = Index)) +
  geom_pointrange(position = position_dodge(width = 0.4), size = 0.6) +
  scale_shape_manual(values = c("First-order" = 16, "Total-order" = 1),
  labels = c("First-order" = "Indice d'ordre 1", "Total-order" = "Indice total")
  ) +
  labs(
    x = "",
    y = "Indice de sensibilité",
    shape = ""
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1),
    legend.position = "right"
  ) +
  coord_cartesian(ylim = c(0, 1))
```

Step 2:
```{r}
model <- gamm(Arsenic_Total ~ s(Rainfall_lag_1, bs='re') + s(O2, bs='re') + s(Temperature, bs='re') 
              + s(Cuivre, bs='re') + s(Conductivité, bs='re') + s(Zinc, bs='re'),
              correlation = corARMA(form = ~ 1, p = 1, q = 1), 
              data = scaled_data, na.action = na.exclude,
              control = lmeControl(msMaxIter = 1000, msMaxEval = 1000))
```

```{r}
predict_model <- function(x) {
  new_data <- data.frame(
    Rainfall_lag_1 = x[, 1],
    O2 = x[, 2],
    Temperature = x[, 3], 
    Cuivre = x[, 4],
    Conductivité = x[, 5],
    Zinc = x[, 6]
  )
  predict(model$gam, newdata = new_data, type = "response")
}
```

```{r, warning=FALSE}
n <- 1000
X1 <- data.frame(matrix(runif(6 * n), nrow = n))
X2 <- data.frame(matrix(runif(6 * n), nrow = n))

sa_result <- sobol2002(
  model = predict_model,
  X1 = X1,
  X2 = X2,
  nboot = 100
)

var_labels <- c("X1" = "Pluie à t-1", "X2" = "O2", "X3" = "Température", "X4" = "Cuivre", 
  "X5" = "Conductivité", "X6" = "Zinc")

df_plot <- bind_rows(
  sa_result$S %>%
    mutate(Variable = rownames(.), Index = "First-order") %>%
    rename(y = original, ymin = `min. c.i.`, ymax = `max. c.i.`),
  sa_result$T %>%
    mutate(Variable = rownames(.), Index = "Total-order") %>%
    rename(y = original, ymin = `min. c.i.`, ymax = `max. c.i.`)
) %>%
  mutate(
    Variable = factor(Variable, levels = names(var_labels)),
    Label = var_labels[as.character(Variable)]
  )

ggplot(df_plot, aes(x = Label, y = y, ymin = ymin, ymax = ymax, shape = Index)) +
  geom_pointrange(position = position_dodge(width = 0.4), size = 0.6) +
  scale_shape_manual(values = c("First-order" = 16, "Total-order" = 1),
  labels = c("First-order" = "Indice d'ordre 1", "Total-order" = "Indice total")
  ) +
  labs(
    x = "",
    y = "Indice de sensibilité",
    shape = ""
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1),
    legend.position = "right"
  ) +
  coord_cartesian(ylim = c(0, 1))
```

```{r}
model <- gamm(Arsenic_Total ~ s(Rainfall_lag_1, bs='re') + s(O2, bs='re') + s(Temperature, bs='re') 
              + s(Cuivre, bs='re') + s(Conductivité, bs='re'),
              correlation = corARMA(form = ~ 1, p = 1, q = 1), 
              data = scaled_data, na.action = na.exclude)
```

```{r}
predict_model <- function(x) {
  new_data <- data.frame(
    Rainfall_lag_1 = x[, 1],
    O2 = x[, 2],
    Temperature = x[, 3], 
    Cuivre = x[, 4],
    Conductivité = x[, 5]
  )
  predict(model$gam, newdata = new_data, type = "response")
}
```

```{r, warning=FALSE}
n <- 1000
X1 <- data.frame(matrix(runif(5 * n), nrow = n))
X2 <- data.frame(matrix(runif(5 * n), nrow = n))

sa_result <- sobol2002(
  model = predict_model,
  X1 = X1,
  X2 = X2,
  nboot = 100
)

var_labels <- c("X1" = "Pluie à t-1", "X2" = "O2", "X3" = "Température", "X4" = "Cuivre", 
  "X5" = "Conductivité")

df_plot <- bind_rows(
  sa_result$S %>%
    mutate(Variable = rownames(.), Index = "First-order") %>%
    rename(y = original, ymin = `min. c.i.`, ymax = `max. c.i.`),
  sa_result$T %>%
    mutate(Variable = rownames(.), Index = "Total-order") %>%
    rename(y = original, ymin = `min. c.i.`, ymax = `max. c.i.`)
) %>%
  mutate(
    Variable = factor(Variable, levels = names(var_labels)),
    Label = var_labels[as.character(Variable)]
  )

ggplot(df_plot, aes(x = Label, y = y, ymin = ymin, ymax = ymax, shape = Index)) +
  geom_pointrange(position = position_dodge(width = 0.4), size = 0.6) +
  scale_shape_manual(values = c("First-order" = 16, "Total-order" = 1),
  labels = c("First-order" = "Indice d'ordre 1", "Total-order" = "Indice total")
  ) +
  labs(
    x = "",
    y = "Indice de sensibilité",
    shape = ""
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1),
    legend.position = "right"
  ) +
  coord_cartesian(ylim = c(0, 1))
```
A higher value of S (First-Order Indices) indicates that the input variable has a significant individual effect on the output variance.

A higher value of T (Total-Order Indices) indicates that the input variable, including its interactions with other variables, has a significant effect on the output variance.

Too much variances caught, must be removed: Zinc, Ion_Ferreux.
Equal high variances caught: Cuivre, O2, Rainfall_lag_1.

Let's compare the three consecutive results with the smaller set of data.
Step 1:
```{r}
scaled_data_small <- data_small %>%
  mutate_at(3:17, rescale)

model_small <- gamm(Arsenic_Total ~ s(Rainfall_lag_1, bs='re') + s(O2, bs='re') + s(Temperature, bs='re') 
              + s(Cuivre, bs='re') + s(Conductivité, bs='re') + s(Zinc, bs='re') + s(Ion_Ferreux, bs='re'),
              correlation = corARMA(form = ~ 1, p = 1, q = 1), 
              data = scaled_data_small, na.action = na.exclude,
              control = lmeControl(msMaxIter = 1000, msMaxEval = 1000))
```

```{r}
predict_model_small <- function(x) {
  new_data <- data.frame(
    Rainfall_lag_1 = x[, 1],
    O2 = x[, 2],
    Temperature = x[, 3], 
    Cuivre = x[, 4],
    Conductivité = x[, 5],
    Zinc = x[, 6],
    Ion_Ferreux = x[, 7]
  )
  predict(model_small$gam, newdata = new_data, type = "response")
}
```

```{r, warning=FALSE}
n <- 1000
X1 <- data.frame(matrix(runif(7 * n), nrow = n))
X2 <- data.frame(matrix(runif(7 * n), nrow = n))

sa_result_small <- sobol2002(
  model = predict_model_small,
  X1 = X1,
  X2 = X2,
  nboot = 100
)

var_labels <- c("X1" = "Pluie à t-1", "X2" = "O2", "X3" = "Température", "X4" = "Cuivre", 
  "X5" = "Conductivité", "X6" = "Zinc", "X7" = "Ion ferreux")

df_plot <- bind_rows(
  sa_result_small$S %>%
    mutate(Variable = rownames(.), Index = "First-order") %>%
    rename(y = original, ymin = `min. c.i.`, ymax = `max. c.i.`),
  sa_result_small$T %>%
    mutate(Variable = rownames(.), Index = "Total-order") %>%
    rename(y = original, ymin = `min. c.i.`, ymax = `max. c.i.`)
) %>%
  mutate(
    Variable = factor(Variable, levels = names(var_labels)),
    Label = var_labels[as.character(Variable)]
  )

ggplot(df_plot, aes(x = Label, y = y, ymin = ymin, ymax = ymax, shape = Index)) +
  geom_pointrange(position = position_dodge(width = 0.4), size = 0.6) +
  scale_shape_manual(values = c("First-order" = 16, "Total-order" = 1),
  labels = c("First-order" = "Indice d'ordre 1", "Total-order" = "Indice total")
  ) +
  labs(
    x = "",
    y = "Indice de sensibilité",
    shape = ""
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1),
    legend.position = "right"
  ) +
  coord_cartesian(ylim = c(0, 1))
```

Step 2:
```{r}
model_small <- gamm(Arsenic_Total ~ s(Rainfall_lag_1, bs='re') + s(O2, bs='re') + s(Temperature, bs='re') 
              + s(Cuivre, bs='re') + s(Conductivité, bs='re') + s(Ion_Ferreux, bs='re'),
              correlation = corARMA(form = ~ 1, p = 1, q = 1), 
              data = scaled_data_small, na.action = na.exclude,
              control = lmeControl(msMaxIter = 1000, msMaxEval = 1000))
```

```{r}
predict_model_small <- function(x) {
  new_data <- data.frame(
    Rainfall_lag_1 = x[, 1],
    O2 = x[, 2],
    Temperature = x[, 3], 
    Cuivre = x[, 4],
    Conductivité = x[, 5],
    Ion_Ferreux = x[, 6]
  )
  predict(model_small$gam, newdata = new_data, type = "response")
}
```

```{r, warning=FALSE}
n <- 1000
X1 <- data.frame(matrix(runif(6 * n), nrow = n))
X2 <- data.frame(matrix(runif(6 * n), nrow = n))

sa_result_small <- sobol2002(
  model = predict_model_small,
  X1 = X1,
  X2 = X2,
  nboot = 100
)

var_labels <- c("X1" = "Pluie à t-1", "X2" = "O2", "X3" = "Température", "X4" = "Cuivre", 
  "X5" = "Conductivité", "X6" = "Ion ferreux")

df_plot <- bind_rows(
  sa_result_small$S %>%
    mutate(Variable = rownames(.), Index = "First-order") %>%
    rename(y = original, ymin = `min. c.i.`, ymax = `max. c.i.`),
  sa_result_small$T %>%
    mutate(Variable = rownames(.), Index = "Total-order") %>%
    rename(y = original, ymin = `min. c.i.`, ymax = `max. c.i.`)
) %>%
  mutate(
    Variable = factor(Variable, levels = names(var_labels)),
    Label = var_labels[as.character(Variable)]
  )

ggplot(df_plot, aes(x = Label, y = y, ymin = ymin, ymax = ymax, shape = Index)) +
  geom_pointrange(position = position_dodge(width = 0.4), size = 0.6) +
  scale_shape_manual(values = c("First-order" = 16, "Total-order" = 1),
  labels = c("First-order" = "Indice d'ordre 1", "Total-order" = "Indice total")
  ) +
  labs(
    x = "",
    y = "Indice de sensibilité",
    shape = ""
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1),
    legend.position = "right"
  ) +
  coord_cartesian(ylim = c(0, 1))
```

Step 3:
```{r}
model_small <- gamm(Arsenic_Total ~ s(Rainfall_lag_1, bs='re') + s(O2, bs='re') + s(Temperature, bs='re') 
              + s(Cuivre, bs='re') + s(Conductivité, bs='re'),
              correlation = corARMA(form = ~ 1, p = 1, q = 1), 
              data = scaled_data_small, na.action = na.exclude,
              control = lmeControl(msMaxIter = 1000, msMaxEval = 1000))
```

```{r}
predict_model_small <- function(x) {
  new_data <- data.frame(
    Rainfall_lag_1 = x[, 1],
    O2 = x[, 2],
    Temperature = x[, 3], 
    Cuivre = x[, 4],
    Conductivité = x[, 5]
  )
  predict(model_small$gam, newdata = new_data, type = "response")
}
```

```{r, warning=FALSE}
n <- 1000
X1 <- data.frame(matrix(runif(5 * n), nrow = n))
X2 <- data.frame(matrix(runif(5 * n), nrow = n))

sa_result_small <- sobol2002(
  model = predict_model_small,
  X1 = X1,
  X2 = X2,
  nboot = 100
)

var_labels <- c("X1" = "Pluie à t-1", "X2" = "O2", "X3" = "Température", "X4" = "Cuivre", 
  "X5" = "Conductivité")

df_plot <- bind_rows(
  sa_result_small$S %>%
    mutate(Variable = rownames(.), Index = "First-order") %>%
    rename(y = original, ymin = `min. c.i.`, ymax = `max. c.i.`),
  sa_result_small$T %>%
    mutate(Variable = rownames(.), Index = "Total-order") %>%
    rename(y = original, ymin = `min. c.i.`, ymax = `max. c.i.`)
) %>%
  mutate(
    Variable = factor(Variable, levels = names(var_labels)),
    Label = var_labels[as.character(Variable)]
  )

ggplot(df_plot, aes(x = Label, y = y, ymin = ymin, ymax = ymax, shape = Index)) +
  geom_pointrange(position = position_dodge(width = 0.4), size = 0.6) +
  scale_shape_manual(values = c("First-order" = 16, "Total-order" = 1),
  labels = c("First-order" = "Indice d'ordre 1", "Total-order" = "Indice total")
  ) +
  labs(
    x = "",
    y = "Indice de sensibilité",
    shape = ""
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1),
    legend.position = "right"
  ) +
  coord_cartesian(ylim = c(0, 1))
```

Equal high variances caught: Copper, Conductivity.
Too much variances caught, must be removed: Zinc.
A little too much variance caught, must be removed: FeII.

# Methods grouping

```{r}
O2_method <- data_big %>%
  group_by(O2_method) %>%
  summarise(Period) %>%
  filter(O2_method != "Undetermined_O2_method")

Plomb_method <- data_big %>%
  group_by(Plomb_method) %>%
  summarise(Period) %>%
  filter(Plomb_method != "Undetermined_Plomb_method")

Antimoine_method <- data_big %>%
  group_by(Antimoine_method) %>%
  summarise(Period) %>%
  filter(Antimoine_method != "Undetermined_Antimoine_method")
```

```{r}
summary(O2_method)
summary(Plomb_method)
summary(Antimoine_method)
```

Before 2012, O2 was most of the time measured by Colorimétrie. After then, it was measured by either methods or just HACH's method but we decided to only keep one method (HACH) if more were used.
208Pb is the usual method to measure Lead.
Antimony is sometimes measured by 121Sb and sometimes by 123Sb, there is no notable timeframe for using only one preferable method.





