---
title: "Optimisation of prediction formula in station S1"
author: "CLETZ Laura"
date: "07/07/2025"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Packages needed:
```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(lubridate)
library(mgcv)
here::i_am("Formulae_Optimisation/COWG_Formula_Opt.Rmd")
library(here)
```

```{r}
data <- read.table(here("processed","COWG_gamm_2004_2022.csv"), sep=";", dec=".", header=T)
```

In this RMarkdown file, there has been various attempt to find the best formula for Arsenic_Total, Zinc, Plomb and Ion_Ferreux variables following four steps:
- thinking of a plausible and logical correlation between variables ;
- looking at plots between the outcome and each plausible variable to find a function that might fit better ;
- using ANOVA on small LM' fits to find which interactions are needed ;
- repeating GAMM' and LM' fits in a backward step-wise logic until each remaining predictor is significant and the BIC is minimized.

# Arsenic_Total

```{r}
old_bic <- bic
control <- lmeControl(msMaxIter = 200, msMaxEval = 200, tolerance = 1e-4,
                        niterEM = 50, opt = "optim")

test <- gamm(Arsenic_Total ~ s(O2, bs='re') + s(Sulfate, bs='re') + Ion_Ferreux + Conductivité^2 + Zinc + 
               Zinc:Conductivité + O2:Conductivité + 
               Zinc:Ion_Ferreux + O2:Ion_Ferreux + Zinc:Sulfate + Sulfate:Ion_Ferreux, 
             data=data, correlation=corARMA(form = ~1, p=1),
             control=control)

bic <- BIC(test$lme)
summary(test$gam)
```

The best formulae is:
Arsenic_Total ~ s(O2, bs='re') + s(Sulfate, bs='re') + Ion_Ferreux + Conductivité^2 + Zinc + 
               Zinc:Conductivité + O2:Conductivité + 
               Zinc:Ion_Ferreux + O2:Ion_Ferreux + Zinc:Sulfate + Sulfate:Ion_Ferreux

with correlation AR(1).

# Zinc

```{r}
old_bic <- bic
control <- lmeControl(msMaxIter = 200, msMaxEval = 200, tolerance = 1e-4,
                        niterEM = 50, opt = "optim")

test <- gamm(Zinc ~ s(Sulfate, bs='re') + Ion_Ferreux + Plomb + Conductivité^2 + Arsenic_Total +
               Sulfate:Ion_Ferreux +Sulfate:Plomb + Sulfate:Conductivité, data=data, 
             correlation=corARMA(form = ~1, p=1), control=control)

bic <- BIC(test$lme)
summary(test$gam)
```

The best formulae is:
Zinc ~ s(Sulfate, bs='re') + Ion_Ferreux + Plomb + Conductivité^2 + Arsenic_Total +
               Sulfate:Ion_Ferreux +Sulfate:Plomb + Sulfate:Conductivité

with correlation AR(1).

# Plomb

```{r}
old_bic <- bic
control <- lmeControl(msMaxIter = 200, msMaxEval = 200, tolerance = 1e-4,
                        niterEM = 50, opt = "optim")

test <- gamm(Plomb ~ s(Rainfall_lag_1, bs='re') + s(O2, bs='re') + 
               Zinc + Ion_Ferreux + pH + Conductivité + Arsenic_Total +
               Conductivité:Arsenic_Total + Rainfall_lag_1:Conductivité + Arsenic_Total:Ion_Ferreux, 
             data=data, correlation=corARMA(form = ~1, p=1),
             control=control)

bic <- BIC(test$lme)
summary(test$gam)
```

The best formulae is:
Plomb ~ s(Rainfall_lag_1, bs='re') + s(O2, bs='re') + Zinc + Ion_Ferreux + pH + Conductivité + Arsenic_Total +
               Conductivité:Arsenic_Total + Rainfall_lag_1:Conductivité + Arsenic_Total:Ion_Ferreux

with correlation AR(1). (It's not that different from the formulae in S1.)

# Ion_Ferreux

```{r}
old_bic <- bic
control <- lmeControl(msMaxIter = 200, msMaxEval = 200, tolerance = 1e-4,
                        niterEM = 50, opt = "optim")

test <- gamm(Ion_Ferreux ~ s(Rainfall, bs='re') + s(O2, bs='re') + Plomb + Conductivité + 
               sqrt(Arsenic_Total) + Plomb:Conductivité + Rainfall:O2, 
             data=data, correlation=corARMA(form = ~1, p=1),
             control=control)

bic <- BIC(test$lme)
summary(test$gam)
```

The best formulae is:
Ion_Ferreux ~ s(Rainfall, bs='re') + s(O2, bs='re') + Plomb + Conductivité + 
               sqrt(Arsenic_Total) + Plomb:Conductivité + Rainfall:O2

with correlation AR(1).
