---
title: "Optimisation of prediction formula in station S1"
author: "CLETZ Laura"
date: "07/07/2025"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Packages needed:
```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(lubridate)
library(mgcv)
here::i_am("Formulae_Optimisation/S1_Formula_Opt.Rmd")
library(here)
```

```{r}
data <- read.table(here("processed","S1_gamm_2004_2019_twice.csv"), sep=";", dec=".", header=T)
```

In this RMarkdown file, there has been various attempt to find the best formula for Arsenic_Total, Zinc, Plomb and Ion_Ferreux variables following four steps:
- thinking of a plausible and logical correlation between variables ;
- looking at plots between the outcome and each plausible variable to find a function that might fit better ;
- using ANOVA on small LM' fits to find which interactions are needed ;
- repeating GAMM' and LM' fits in a backward step-wise logic until each remaining predictor is significant and the BIC is minimized.

# Arsenic_Total

```{r}
old_bic <- bic
control <- lmeControl(msMaxIter = 200, msMaxEval = 200, tolerance = 1e-4,
                        niterEM = 50, opt = "optim")

test <- gamm(log(Arsenic_Total + 1) ~ s(Rainfall_lag_1, bs='re') + s(O2, bs='re') +
               s(Sulfate, bs='re') + Ion_Ferreux +  Numeric_Dates + Plomb + Conductivité^2 + Zinc + 
               Zinc:Sulfate + Sulfate:Ion_Ferreux + Plomb:O2 + Zinc:Ion_Ferreux, 
             data=data, correlation=corARMA(form = ~1, p=1),
             control=control)

bic <- BIC(test$lme)
summary(test$gam)
```

The best formulae is:
log(Arsenic_Total + 1) ~ s(Rainfall_lag_1, bs='re') + s(O2, bs='re') +
               s(Sulfate, bs='re') + Ion_Ferreux +  Numeric_Dates + Plomb + Conductivité^2 + Zinc + 
               Zinc:Sulfate + Sulfate:Ion_Ferreux + Plomb:O2 + Zinc:Ion_Ferreux

with correlation AR(1).

# Zinc

```{r}
old_bic <- bic

test <- lm(Zinc ~ Numeric_Dates + Ion_Ferreux +  Arsenic_Total, data=data)

bic <- BIC(test)
summary(test)
```

The best formulae is:
Zinc ~ Numeric_Dates + Ion_Ferreux +  Arsenic_Total

for basic linear regression.

# Plomb

```{r}
old_bic <- bic

test <- gamm(Plomb ~ s(Rainfall_lag_1, bs='re') + s(pH, bs='re') + 
               log(Arsenic_Total + 1) + Zinc + Conductivité + pH:Rainfall_lag_1 + 
               Conductivité:Rainfall_lag_1 + Conductivité:Arsenic_Total, 
             data=data, correlation=corARMA(form = ~1, p=1),
             control=control)

bic <- BIC(test$lme)
summary(test$gam)
```

The best formulae is:
Plomb ~ s(Rainfall_lag_1, bs='re') + s(pH, bs='re') + 
               log(Arsenic_Total + 1) + Zinc + Conductivité + pH:Rainfall_lag_1 + 
               Conductivité:Rainfall_lag_1 + Conductivité:Arsenic_Total

with correlation AR(1).

# Ion_Ferreux

```{r}
old_bic <- bic

test <- gamm(Ion_Ferreux ~ s(Rainfall, bs='re') + s(O2, bs='re') + s(Sulfate, bs='re') +
               Conductivité + sqrt(Arsenic_Total) + Rainfall:O2, 
             data=data, correlation=corARMA(form = ~1, p=1),
             control=control)

bic <- BIC(test$lme)
summary(test$gam)
```

The best formulae is:
Ion_Ferreux ~ s(Rainfall, bs='re') + s(O2, bs='re') + s(Sulfate, bs='re') +
               Conductivité + sqrt(Arsenic_Total) + Rainfall:O2

with correlation AR(1).

