---
title: "Optimisation of prediction formula in station S1"
author: "CLETZ Laura"
date: "07/07/2025"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Packages needed:
```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(lubridate)
library(mgcv)
library(scales)
here::i_am("Formulae_Optimisation/S1_Formula_Opt.Rmd")
library(here)
```

```{r}
data <- read.table(here("processed","S1_scaled.csv"), sep=";", dec=".", header=T)
```

```{r}
data_big <- read.table(here("processed","S1_gamm_2004_2019_twice.csv"), sep=";", dec=".", header=T)%>%
  mutate_at(c(3:18), rescale)
```

In this RMarkdown file, there has been various attempt to find the best formula for total_arsenic, Zinc, Plomb and Ion_Ferreux variables following four steps:
- thinking of a plausible and logical correlation between variables ;
- looking at plots between the outcome and each plausible variable to find a function that might fit better ;
- using ANOVA on small LM' fits to find which interactions are needed ;
- repeating GAMM' and LM' fits in a backward step-wise logic until each remaining predictor is significant and the BIC is minimized.

```{r}
bic <- NA
aic <- NA
```

The MFA results showed that rainfall at any time, O2, pH, conductivity, to-explain variables (arsenic, iron, lead and zinc), Sulfate, and the dates/periods are somehow correlated. We created 4 new variables out of it.

# total_arsenic
## From 2004 to 2010

```{r}
reg1 <- lm(total_arsenic ~ rainfall + o2, data=data)
reg2 <- lm(total_arsenic ~ redox, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(total_arsenic ~ redox + conductivity, data=data)
reg2 <- lm(total_arsenic ~ redox * conductivity, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(total_arsenic ~ redox + sulfate, data=data)
reg2 <- lm(total_arsenic ~ redox * sulfate, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(total_arsenic ~ zinc + redox, data=data)
reg2 <- lm(total_arsenic ~ zinc * redox, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(total_arsenic ~ redox + feII, data=data)
reg2 <- lm(total_arsenic ~ redox * feII, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(total_arsenic ~ lead + redox, data=data)
reg2 <- lm(total_arsenic ~ lead * redox, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(total_arsenic ~ metaux_temp + redox, data=data)
reg2 <- lm(total_arsenic ~ metaux_temp * redox, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(total_arsenic ~ lead + zinc, data=data)
reg2 <- lm(total_arsenic ~ lead * zinc, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(total_arsenic ~ metaux_temp + metaux, data=data)
reg2 <- lm(total_arsenic ~ metaux_temp * metaux, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(total_arsenic ~ lead + feII, data=data)
reg2 <- lm(total_arsenic ~ lead * feII, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(total_arsenic ~ conductivity + redox, data=data)
reg2 <- lm(total_arsenic ~ conductivity * redox, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(total_arsenic ~ zinc + feII, data=data)
reg2 <- lm(total_arsenic ~ zinc * feII, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(total_arsenic ~ zinc + sulfate, data=data)
reg2 <- lm(total_arsenic ~ zinc * sulfate, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(total_arsenic ~ sulfate + conductivity, data=data)
reg2 <- lm(total_arsenic ~ sulfate * conductivity, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(total_arsenic ~ sulfate + lead, data=data)
reg2 <- lm(total_arsenic ~ sulfate * lead, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(total_arsenic ~ rainfall_lag_1 + conductivity, data=data)
reg2 <- lm(total_arsenic ~ rainfall_lag_1 * conductivity, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(total_arsenic ~ rainfall_lag_1 + feII, data=data)
reg2 <- lm(total_arsenic ~ rainfall_lag_1 * feII, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(total_arsenic ~ rainfall_lag_1 + zinc, data=data)
reg2 <- lm(total_arsenic ~ rainfall_lag_1 * zinc, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(total_arsenic ~ redox + numeric_dates, data=data)
reg2 <- lm(total_arsenic ~ redox * numeric_dates, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(total_arsenic ~ feII + numeric_dates, data=data)
reg2 <- lm(total_arsenic ~ feII * numeric_dates, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(total_arsenic ~ lead + numeric_dates, data=data)
reg2 <- lm(total_arsenic ~ lead * numeric_dates, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(total_arsenic ~ zinc + numeric_dates, data=data)
reg2 <- lm(total_arsenic ~ zinc * numeric_dates, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(total_arsenic ~ rainfall_lag_1 + numeric_dates, data=data)
reg2 <- lm(total_arsenic ~ rainfall_lag_1 * numeric_dates, data=data)
anova(reg1, reg2)
```

Thus the initial formula:
total_arsenic ~ redox + rainfall_lag_1 + zinc + s(lead, bs='bs') + feII + s(sulfate, bs='bs') + numeric_dates +
                    redox:numeric_dates + redox:zinc + redox:feII + 
                    zinc:numeric_dates + zinc:feII + zinc:lead +
                    lead:feII + rainfall_lag_1:feII
               
```{r}
old_bic <- bic
old_aic <- aic

test <- gamm(total_arsenic ~ redox + rainfall_lag_1 + zinc + s(lead, bs='bs') + feII + s(sulfate, bs='bs') + numeric_dates +
                    redox:numeric_dates + redox:zinc + redox:feII + 
                    zinc:numeric_dates + zinc:feII + zinc:lead +
                    lead:feII + rainfall_lag_1:feII, 
             data=data, correlation=corARMA(form = ~1, p=1))

bic <- BIC(test$lme)
aic <- AIC(test$lme)

summary(test$gam)
```

```{r}
old_bic <- bic
old_aic <- aic

test <- gamm(total_arsenic ~ redox + rainfall_lag_1 + zinc + s(lead, bs='bs') + feII + s(sulfate, bs='bs') + numeric_dates +
                    redox:numeric_dates + redox:zinc + redox:feII + 
                    zinc:numeric_dates + zinc:feII + zinc:lead +
                    lead:feII, 
             data=data, correlation=corARMA(form = ~1, p=1))

bic <- BIC(test$lme)
aic <- AIC(test$lme)

summary(test$gam)
```

```{r}
old_bic <- bic
old_aic <- aic

test <- gamm(total_arsenic ~ redox + rainfall_lag_1 + zinc + s(lead, bs='bs') + feII + s(sulfate, bs='bs') + numeric_dates +
                    redox:numeric_dates + redox:zinc + 
                    zinc:numeric_dates + zinc:feII + zinc:lead +
                    lead:feII, 
             data=data, correlation=corARMA(form = ~1, p=1))

bic <- BIC(test$lme)
aic <- AIC(test$lme)

summary(test$gam)
```

```{r}
old_bic <- bic
old_aic <- aic

test <- gamm(total_arsenic ~ redox + rainfall_lag_1 + zinc + s(lead, bs='bs') + feII + s(sulfate, bs='bs') + numeric_dates +
                    redox:numeric_dates + redox:zinc + 
                    zinc:numeric_dates + zinc:feII + 
                    lead:feII, 
             data=data, correlation=corARMA(form = ~1, p=1))

bic <- BIC(test$lme)
aic <- AIC(test$lme)

summary(test$gam)
```

```{r}
old_bic <- bic
old_aic <- aic

test <- gamm(total_arsenic ~ redox + rainfall_lag_1 + zinc + s(lead, bs='bs') + feII + s(sulfate, bs='bs') + numeric_dates +
                    redox:numeric_dates + redox:zinc + 
                    zinc:numeric_dates + zinc:feII, 
             data=data, correlation=corARMA(form = ~1, p=1))

bic <- BIC(test$lme)
aic <- AIC(test$lme)

summary(test$gam)
```

```{r}
old_bic <- bic
old_aic <- aic

test <- gamm(total_arsenic ~ redox + zinc + s(lead, bs='bs') + feII + s(sulfate, bs='bs') + numeric_dates +
                    redox:numeric_dates + redox:zinc + 
                    zinc:numeric_dates + zinc:feII, 
             data=data, correlation=corARMA(form = ~1, p=1))

bic <- BIC(test$lme)
aic <- AIC(test$lme)

summary(test$gam)
```

```{r}
old_bic <- bic
old_aic <- aic

test <- gamm(total_arsenic ~ redox + zinc + feII + s(sulfate, bs='bs') + numeric_dates +
                    redox:numeric_dates + redox:zinc + 
                    zinc:numeric_dates + zinc:feII, 
             data=data, correlation=corARMA(form = ~1, p=1))

bic <- BIC(test$lme)
aic <- AIC(test$lme)

summary(test$gam)
```

The best formula is:
total_arsenic ~ redox + zinc + feII + s(sulfate, bs='bs') + numeric_dates +
                    redox:numeric_dates + redox:zinc + 
                    zinc:numeric_dates + zinc:feII
with correlation AR(1).

```{r}
qqnorm(test$gam$residuals/sqrt(test$gam$sig2))
abline(a=0,b=1,col='red')
```

```{r}
qqnorm(test$lme$residuals/sqrt(test$lme$sigma^2))
abline(a=0,b=1,col='red')
```

```{r, echo=FALSE}
shapiro.test(test$lme$residuals)
shapiro.test(test$gam$residuals)
```

```{r, echo=FALSE}
shapiro.test(test$lme$residuals)
shapiro.test(test$gam$residuals)
```

```{r}
acf(test$gam$residuals,
    main = "Auto-corrélation des résidus",
    ylab = expression(gamma),
    type = "correlation")
```

## From 2004 to 2019

```{r}
old_bic <- bic
old_aic <- aic

test <- gamm(total_arsenic ~ rainfall + rainfall_lag_1 + 
               o2 + conductivity + s(sulfate, bs = "bs") + feII + zinc +
               rainfall:conductivity + rainfall:o2 +
               rainfall_lag_1:feII + numeric_dates, 
             data=data_big, correlation=corARMA(form = ~1, p=1))


#AIC(test1, mod_alt)

bic <- BIC(test$lme)
aic <- AIC(test$lme)

summary(test$gam)
```

The best formulae is:
total_arsenic ~ rainfall + rainfall_lag_1 + 
               o2 + conductivity + s(sulfate, bs = "bs") + feII + zinc +
               rainfall:conductivity + rainfall:o2 +
               rainfall_lag_1:feII + numeric_dates

with correlation AR(1).

# Zinc
## From 2004 to 2010

```{r}
reg1 <- lm(zinc ~ rainfall + o2, data=data)
reg2 <- lm(zinc ~ redox, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(zinc ~ rainfall + conductivity, data=data)
reg2 <- lm(zinc ~ rainfall * conductivity, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(zinc ~ rainfall + sulfate, data=data)
reg2 <- lm(zinc ~ rainfall * sulfate, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(zinc ~ total_arsenic + rainfall, data=data)
reg2 <- lm(zinc ~ total_arsenic * rainfall, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(zinc ~ rainfall + feII, data=data)
reg2 <- lm(zinc ~ rainfall * feII, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(zinc ~ lead + rainfall, data=data)
reg2 <- lm(zinc ~ lead * rainfall, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(zinc ~ lead + o2, data=data)
reg2 <- lm(zinc ~ lead * o2, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(zinc ~ feII + o2, data=data)
reg2 <- lm(zinc ~ feII * o2, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(zinc ~ total_arsenic + o2, data=data)
reg2 <- lm(zinc ~ total_arsenic * o2, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(zinc ~ metaux_temp + o2, data=data)
reg2 <- lm(zinc ~ metaux_temp * o2, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(zinc ~ metaux_temp + rainfall, data=data)
reg2 <- lm(zinc ~ metaux_temp * rainfall, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(zinc ~ lead + total_arsenic, data=data)
reg2 <- lm(zinc ~ lead * total_arsenic, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(zinc ~ lead + feII, data=data)
reg2 <- lm(zinc ~ lead * feII, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(zinc ~ conductivity + o2, data=data)
reg2 <- lm(zinc ~ conductivity * o2, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(zinc ~ total_arsenic + feII, data=data)
reg2 <- lm(zinc ~ total_arsenic * feII, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(zinc ~ total_arsenic + sulfate, data=data)
reg2 <- lm(zinc ~ total_arsenic * sulfate, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(zinc ~ sulfate + conductivity, data=data)
reg2 <- lm(zinc ~ sulfate * conductivity, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(zinc ~ sulfate + lead, data=data)
reg2 <- lm(zinc ~ sulfate * lead, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(zinc ~ rainfall_lag_1 + conductivity, data=data)
reg2 <- lm(zinc ~ rainfall_lag_1 * conductivity, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(zinc ~ rainfall_lag_1 + feII, data=data)
reg2 <- lm(zinc ~ rainfall_lag_1 * feII, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(zinc ~ rainfall_lag_1 + total_arsenic, data=data)
reg2 <- lm(zinc ~ rainfall_lag_1 * total_arsenic, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(zinc ~ rainfall + numeric_dates, data=data)
reg2 <- lm(zinc ~ rainfall * numeric_dates, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(zinc ~ feII + numeric_dates, data=data)
reg2 <- lm(zinc ~ feII * numeric_dates, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(zinc ~ lead + numeric_dates, data=data)
reg2 <- lm(zinc ~ lead * numeric_dates, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(zinc ~ total_arsenic + numeric_dates, data=data)
reg2 <- lm(zinc ~ total_arsenic * numeric_dates, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(zinc ~ rainfall_lag_1 + numeric_dates, data=data)
reg2 <- lm(zinc ~ rainfall_lag_1 * numeric_dates, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(zinc ~ ph + feII, data=data)
reg2 <- lm(zinc ~ ph * feII, data=data)
anova(reg1, reg2)
```

Thus the initial formula:
zinc ~ rainfall + o2 + conductivity + s(lead, bs='bs') + feII + total_arsenic + numeric_dates + s(sulfate, bs='bs') +
          rainfall:conductivity +
          o2:lead + o2:feII +
          numeric_dates:lead + numeric_dates:total_arsenic
               
```{r}
old_bic <- bic
old_aic <- aic

test <- gamm(zinc ~ rainfall + o2 + conductivity + s(lead, bs='bs') + feII + total_arsenic + numeric_dates + s(sulfate, bs='bs') +
                rainfall:conductivity +
                o2:lead + o2:feII +
                numeric_dates:lead + numeric_dates:total_arsenic, 
             data=data, correlation=corARMA(form = ~1, p=1))

bic <- BIC(test$lme)
aic <- AIC(test$lme)

summary(test$gam)
```

```{r}
old_bic <- bic
old_aic <- aic

test <- gamm(zinc ~ rainfall + o2 + conductivity + s(lead, bs='bs') + feII + total_arsenic + numeric_dates + s(sulfate, bs='bs') +
                o2:lead + o2:feII +
                numeric_dates:lead + numeric_dates:total_arsenic, 
             data=data, correlation=corARMA(form = ~1, p=1))

bic <- BIC(test$lme)
aic <- AIC(test$lme)

summary(test$gam)
```

```{r}
old_bic <- bic
old_aic <- aic

test <- gamm(zinc ~ rainfall + o2 + conductivity + s(lead, bs='bs') + feII + total_arsenic + numeric_dates + s(sulfate, bs='bs') +
                o2:lead + 
                numeric_dates:lead + numeric_dates:total_arsenic, 
             data=data, correlation=corARMA(form = ~1, p=1))

bic <- BIC(test$lme)
aic <- AIC(test$lme)

summary(test$gam)
```

```{r}
old_bic <- bic
old_aic <- aic

test <- gamm(zinc ~ rainfall + o2 + conductivity + s(lead, bs='bs') + feII + total_arsenic + numeric_dates + s(sulfate, bs='bs') +
                o2:lead + 
                numeric_dates:lead, 
             data=data, correlation=corARMA(form = ~1, p=1))

bic <- BIC(test$lme)
aic <- AIC(test$lme)

summary(test$gam)
```

```{r}
old_bic <- bic
old_aic <- aic

test <- gamm(zinc ~ rainfall + o2 + conductivity + s(lead, bs='bs') + feII + total_arsenic + numeric_dates + s(sulfate, bs='bs') +
                numeric_dates:lead, 
             data=data, correlation=corARMA(form = ~1, p=1))

bic <- BIC(test$lme)
aic <- AIC(test$lme)

summary(test$gam)
```

```{r}
old_bic <- bic
old_aic <- aic

test <- gamm(zinc ~ rainfall + o2 + s(lead, bs='bs') + feII + total_arsenic + numeric_dates + s(sulfate, bs='bs') +
                numeric_dates:lead, 
             data=data, correlation=corARMA(form = ~1, p=1))

bic <- BIC(test$lme)
aic <- AIC(test$lme)

summary(test$gam)
```

The best formula is:
zinc ~ rainfall + o2 + s(lead, bs='bs') + feII + total_arsenic + numeric_dates + s(sulfate, bs='bs') +
                numeric_dates:lead
with correlation AR(1).

## From 2004 to 2019
               
```{r}
old_bic <- bic
old_aic <- aic

test <- gamm(zinc ~ rainfall_lag_1 + 
               s(sulfate, bs = "bs") + feII + lead + lead^3 + total_arsenic +
               rainfall_lag_1:total_arsenic + numeric_dates, 
             data=data_big, correlation=corARMA(form = ~1, p=1))


#AIC(test1, mod_alt)

bic <- BIC(test$lme)
aic <- AIC(test$lme)

summary(test$gam)
```

The best formulae is:
zinc ~ rainfall_lag_1 + 
               s(sulfate, bs = "bs") + feII + lead + lead^3 + total_arsenic +
               rainfall_lag_1:total_arsenic + numeric_dates

with correlation AR(1).

# lead
## From 2004 to 2010

```{r}
reg1 <- lm(lead ~ rainfall + o2, data=data)
reg2 <- lm(lead ~ redox, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(lead ~ rainfall + zinc, data=data)
reg2 <- lm(lead ~ rainfall * zinc, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(lead ~ rainfall + total_arsenic, data=data)
reg2 <- lm(lead ~ rainfall * total_arsenic, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(lead ~ rainfall + sulfate, data=data)
reg2 <- lm(lead ~ rainfall * sulfate, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(lead ~ o2 + zinc, data=data)
reg2 <- lm(lead ~ o2 * zinc, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(lead ~ o2 + total_arsenic, data=data)
reg2 <- lm(lead ~ o2 * total_arsenic, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(lead ~ o2 + sulfate, data=data)
reg2 <- lm(lead ~ o2 * sulfate, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(lead ~ o2 + feII, data=data)
reg2 <- lm(lead ~ o2 * feII, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(lead ~ ph + zinc, data=data)
reg2 <- lm(lead ~ ph * zinc, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(lead ~ ph + feII, data=data)
reg2 <- lm(lead ~ ph * feII, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(lead ~ rainfall_lag_1 + zinc, data=data)
reg2 <- lm(lead ~ rainfall_lag_1 * zinc, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(lead ~ rainfall_lag_1 + total_arsenic, data=data)
reg2 <- lm(lead ~ rainfall_lag_1 * total_arsenic, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(lead ~ rainfall_lag_1 + feII, data=data)
reg2 <- lm(lead ~ rainfall_lag_1 * feII, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(lead ~ total_arsenic + zinc, data=data)
reg2 <- lm(lead ~ total_arsenic * zinc, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(lead ~ sulfate + zinc, data=data)
reg2 <- lm(lead ~ sulfate * zinc, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(lead ~ feII + zinc, data=data)
reg2 <- lm(lead ~ feII * zinc, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(lead ~ feII + sulfate, data=data)
reg2 <- lm(lead ~ feII * sulfate, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(lead ~ zinc + conductivity, data=data)
reg2 <- lm(lead ~ metaux_cond, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(lead ~ feII + conductivity, data=data)
reg2 <- lm(lead ~ feII * conductivity, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(lead ~ feII + numeric_dates, data=data)
reg2 <- lm(lead ~ feII * numeric_dates, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(lead ~ total_arsenic + numeric_dates, data=data)
reg2 <- lm(lead ~ total_arsenic * numeric_dates, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(lead ~ zinc + numeric_dates, data=data)
reg2 <- lm(lead ~ zinc * numeric_dates, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(lead ~ sulfate + numeric_dates, data=data)
reg2 <- lm(lead ~ sulfate * numeric_dates, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(lead ~ temperature + numeric_dates, data=data)
reg2 <- lm(lead ~ temperature * numeric_dates, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(lead ~ temperature + zinc, data=data)
reg2 <- lm(lead ~ temperature * zinc, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(lead ~ temperature + sulfate, data=data)
reg2 <- lm(lead ~ temperature * sulfate, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(lead ~ rainfall, data=data)
reg2 <- lm(lead ~ rainfall_lag_1, data=data)
anova(reg1, reg2)
```

Thus the nitial formula:
lead ~ rainfall_lag_1 + o2 + temperature + zinc + feII + total_arsenic + numeric_dates +
          o2:feII + o2:zinc + s(sulfate, bs='bs') +
          total_arsenic:numeric_dates

```{r}
old_bic <- bic
old_aic <- aic

test <- gamm(lead ~ rainfall_lag_1 + o2 + temperature + zinc + feII + total_arsenic + numeric_dates +
                 o2:feII + o2:zinc + s(sulfate, bs='bs') +
                 total_arsenic:numeric_dates,
             data=data, correlation=corARMA(form = ~1, p=1))

bic <- BIC(test$lme)
aic <- AIC(test$lme)
summary(test$gam)
```

```{r}
old_bic <- bic
old_aic <- aic

test <- gamm(lead ~ rainfall_lag_1 + o2 + temperature + zinc + feII + total_arsenic + numeric_dates +
                 o2:zinc + s(sulfate, bs='bs') +
                 total_arsenic:numeric_dates,
             data=data, correlation=corARMA(form = ~1, p=1))

bic <- BIC(test$lme)
aic <- AIC(test$lme)
summary(test$gam)
```

```{r}
old_bic <- bic
old_aic <- aic

test <- gamm(lead ~ rainfall_lag_1 + o2 + temperature + zinc + feII + total_arsenic + numeric_dates +
                 o2:zinc + s(sulfate, bs='bs'),
             data=data, correlation=corARMA(form = ~1, p=1))

bic <- BIC(test$lme)
aic <- AIC(test$lme)
summary(test$gam)
```

```{r}
old_bic <- bic
old_aic <- aic

test <- gamm(lead ~ rainfall_lag_1 + o2 + temperature + zinc + feII + numeric_dates +
                 o2:zinc + s(sulfate, bs='bs'),
             data=data, correlation=corARMA(form = ~1, p=1))

bic <- BIC(test$lme)
aic <- AIC(test$lme)
summary(test$gam)
```

```{r}
old_bic <- bic
old_aic <- aic

test <- gamm(lead ~ rainfall_lag_1 + o2 + temperature + zinc + numeric_dates +
                 o2:zinc + s(sulfate, bs='bs'),
             data=data, correlation=corARMA(form = ~1, p=1))

bic <- BIC(test$lme)
aic <- AIC(test$lme)
summary(test$gam)
```

```{r}
old_bic <- bic
old_aic <- aic

test <- gamm(lead ~ rainfall_lag_1 + o2 + temperature + zinc + 
                 o2:zinc + s(sulfate, bs='bs'),
             data=data, correlation=corARMA(form = ~1, p=1))

bic <- BIC(test$lme)
aic <- AIC(test$lme)
summary(test$gam)
```

The best formula is:
lead ~ rainfall_lag_1 + o2 + temperature + zinc + 
                 o2:zinc + s(sulfate, bs='bs')

with correlation AR(1).

## From 2004 to 2019

```{r}
old_bic <- bic
old_aic <- aic

test <- gamm(lead ~ rainfall_lag_1 + 
               o2 + s(conductivity, bs='bs') + s(sulfate, bs = "bs") + zinc + zinc^3 + 
               rainfall_lag_1:zinc + rainfall_lag_1:o2, 
             data=data_big, correlation=corARMA(form = ~1, p=1))

bic <- BIC(test$lme)
aic <- AIC(test$lme)
summary(test$gam)
```

The best formulae is:
lead ~ rainfall_lag_1 + 
               o2 + s(conductivity, bs='bs') + s(sulfate, bs = "bs") + zinc + zinc^3 + 
               rainfall_lag_1:zinc + rainfall_lag_1:o2

with correlation AR(1).

# feII
## From 2004 to 2010

```{r}
reg1 <- lm(feII ~ rainfall, data=data)
reg2 <- lm(feII ~ rainfall_lag_1, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(feII ~ rainfall + ph, data=data)
reg2 <- lm(feII ~ redox, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(feII ~ redox + total_arsenic, data=data)
reg2 <- lm(feII ~ redox * total_arsenic, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(feII ~ redox + zinc, data=data)
reg2 <- lm(feII ~ redox * zinc, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(feII ~ redox + lead, data=data)
reg2 <- lm(feII ~ redox * lead, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(feII ~ redox + sulfate, data=data)
reg2 <- lm(feII ~ redox * sulfate, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(feII ~ conductivity + sulfate, data=data)
reg2 <- lm(feII ~ conductivity * sulfate, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(feII ~ conductivity + total_arsenic, data=data)
reg2 <- lm(feII ~ conductivity * total_arsenic, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(feII ~ conductivity + zinc, data=data)
reg2 <- lm(feII ~ conductivity * zinc, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(feII ~ zinc + sulfate, data=data)
reg2 <- lm(feII ~ zinc * sulfate, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(feII ~ total_arsenic + sulfate, data=data)
reg2 <- lm(feII ~ total_arsenic * sulfate, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(feII ~ zinc + total_arsenic, data=data)
reg2 <- lm(feII ~ zinc * total_arsenic, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(feII ~ lead + temperature, data=data)
reg2 <- lm(feII ~ metaux_temp, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(feII ~ zinc + numeric_dates, data=data)
reg2 <- lm(feII ~ zinc * numeric_dates, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(feII ~ sulfate + numeric_dates, data=data)
reg2 <- lm(feII ~ sulfate * numeric_dates, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(feII ~ total_arsenic + numeric_dates, data=data)
reg2 <- lm(feII ~ total_arsenic * numeric_dates, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(feII ~ lead + numeric_dates, data=data)
reg2 <- lm(feII ~ lead * numeric_dates, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(feII ~ metaux_temp + numeric_dates, data=data)
reg2 <- lm(feII ~ metaux_temp * numeric_dates, data=data)
anova(reg1, reg2)
```

```{r}
reg1 <- lm(feII ~ conductivity + numeric_dates, data=data)
reg2 <- lm(feII ~ conductivity * numeric_dates, data=data)
anova(reg1, reg2)
```
Initial formulae:
feII ~ redox + conductivity + zinc + total_arsenic + lead + s(sulfate, bs="bs") + numeric_dates +
          sulfate:redox + sulfate:conductivity + sulfate:zinc + sulfate:numeric_dates +
          zinc:total_arsenic
               
```{r}
old_bic <- bic
old_aic <- aic

test <- gamm(feII ~ redox + conductivity + zinc + total_arsenic + lead + s(sulfate, bs="bs") + numeric_dates +
                     sulfate:redox + sulfate:conductivity + sulfate:zinc + sulfate:numeric_dates +
                     zinc:total_arsenic, 
             data=data, correlation=corARMA(form = ~1, p=1))

bic <- BIC(test$lme)
aic <- AIC(test$lme)
summary(test$gam)
```

```{r}
old_bic <- bic
old_aic <- aic

test <- gamm(feII ~ redox + conductivity + zinc + total_arsenic + lead + s(sulfate, bs="bs") + numeric_dates +
                     sulfate:redox + sulfate:conductivity + sulfate:numeric_dates +
                     zinc:total_arsenic, 
             data=data, correlation=corARMA(form = ~1, p=1))

bic <- BIC(test$lme)
aic <- AIC(test$lme)
summary(test$gam)
```

```{r}
old_bic <- bic
old_aic <- aic

test <- gamm(feII ~ redox + conductivity + zinc + total_arsenic + s(sulfate, bs="bs") + numeric_dates +
                     sulfate:redox + sulfate:conductivity + sulfate:numeric_dates +
                     zinc:total_arsenic, 
             data=data, correlation=corARMA(form = ~1, p=1))

bic <- BIC(test$lme)
aic <- AIC(test$lme)
summary(test$gam)
```

```{r}
old_bic <- bic
old_aic <- aic

test <- gamm(feII ~ redox + conductivity + zinc + total_arsenic + s(sulfate, bs="bs") + numeric_dates +
                     sulfate:conductivity + sulfate:numeric_dates +
                     zinc:total_arsenic, 
             data=data, correlation=corARMA(form = ~1, p=1))

bic <- BIC(test$lme)
aic <- AIC(test$lme)
summary(test$gam)
```

```{r}
old_bic <- bic
old_aic <- aic

test <- gamm(feII ~ conductivity + zinc + total_arsenic + s(sulfate, bs="bs") + numeric_dates +
                     sulfate:conductivity + sulfate:numeric_dates +
                     zinc:total_arsenic, 
             data=data, correlation=corARMA(form = ~1, p=1))

bic <- BIC(test$lme)
aic <- AIC(test$lme)
summary(test$gam)
```

The best formula is:
feII ~ rainfall + rainfall_lag_1 + 
               conductivity + s(sulfate, bs = "bs") + arsenic_zinc +
               rainfall:sulfate + 
               rainfall_lag_1:arsenic_zinc + rainfall_lag_1:sulfate +
               numeric_dates

with correlation AR(1).

## From 2004 to 2019

```{r}
old_bic <- bic
old_aic <- aic

test <- gamm(feII ~ rainfall + rainfall_lag_1 + 
               conductivity + s(sulfate, bs = "bs") + zinc + total_arsenic +
               rainfall:sulfate + 
               rainfall_lag_1:total_arsenic + rainfall_lag_1:zinc + rainfall_lag_1:sulfate, 
             data=data_big, correlation=corARMA(form = ~1, p=1))

bic <- BIC(test$lme)
aic <- AIC(test$lme)
summary(test$gam)
```

The best formulae is:
feII ~ rainfall + rainfall_lag_1 + 
               conductivity + s(sulfate, bs = "bs") + zinc + total_arsenic +
               rainfall:sulfate + 
               rainfall_lag_1:total_arsenic + rainfall_lag_1:zinc + rainfall_lag_1:sulfate

with correlation AR(1).

## From 2004 to 2010, new concept

Initial formulae:
feII ~ ph_rain + rainfall_lag_1 + 
               s(o2, bs='bs') + s(conductivity, bs='bs') + s(sulfate, bs = "bs") + arsenic_zinc + I(lead + lead^3) +
               lead:sulfate + arsenic_zinc:sulfate +
               ph_rain:arsenic_zinc + ph_rain:lead + ph_rain:sulfate + ph_rain:o2 +
               rainfall_lag_1:arsenic_zinc + rainfall_lag_1:lead + rainfall_lag_1:sulfate + 
               rainfall_lag_1:o2 + s(numeric_dates, bs='bs')
               
```{r}
old_bic <- bic
old_aic <- aic

test <- gamm(feII ~ rainfall_lag_1 + 
               s(sulfate, bs = "bs") + arsenic_zinc + I(lead + lead^3) +
               arsenic_zinc:sulfate + rainfall_lag_1:sulfate + 
               s(numeric_dates, bs='bs'), 
             data=data, correlation=corARMA(form = ~1, p=1))

bic <- BIC(test$lme)
aic <- AIC(test$lme)

summary(test$gam)
```

The best formulae is:
feII ~ rainfall_lag_1 + 
               s(sulfate, bs = "bs") + arsenic_zinc + I(lead + lead^3) +
               arsenic_zinc:sulfate + rainfall_lag_1:sulfate + 
               s(numeric_dates, bs='bs')
               
with correlation AR(1).
