---
title: "Optimisation of prediction formula in station S1"
author: "CLETZ Laura"
date: "07/07/2025"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Packages needed:
```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(lubridate)
library(mgcv)
library(scales)
here::i_am("Formulae_Optimisation/S1_Formula_Opt.Rmd")
library(here)
```

```{r}
data <- read.table(here("processed","S1_scaled.csv"), sep=";", dec=".", header=T)
```

```{r}
data_big <- read.table(here("processed","S1_gamm_2004_2019_twice.csv"), sep=";", dec=".", header=T)%>%
  mutate_at(c(3:18), rescale)
```

In this RMarkdown file, there has been various attempt to find the best formula for total_arsenic, Zinc, Plomb and Ion_Ferreux variables following four steps:
- thinking of a plausible and logical correlation between variables ;
- looking at plots between the outcome and each plausible variable to find a function that might fit better ;
- using ANOVA on small LM' fits to find which interactions are needed ;
- repeating GAMM' and LM' fits in a backward step-wise logic until each remaining predictor is significant and the BIC is minimized.

```{r}
bic <- NA
aic <- NA
```

# total_arsenic
## From 2004 to 2010

Initial formulae:
total_arsenic ~ rainfall + rainfall_lag_1 + 
               s(o2, bs='bs') + s(conductivity, bs='bs') + s(sulfate, bs = "bs") + feII + lead + lead^3 + zinc +
               rainfall:zinc + rainfall:lead + rainfall:feII + rainfall:sulfate + rainfall:conductivity + rainfall:o2 +
               rainfall_lag_1:zinc + rainfall_lag_1:lead + rainfall_lag_1:feII + rainfall_lag_1:sulfate +
               rainfall_lag_1:conductivity + rainfall_lag_1:o2 + numeric_dates
               
```{r}
old_bic <- bic
old_aic <- aic

test <- gamm(total_arsenic ~ rainfall +  
               o2 + s(conductivity, bs='bs') + s(sulfate, bs = "bs") + feII + lead + lead^3 + zinc +
               rainfall:zinc + rainfall:lead + rainfall:feII + rainfall:o2, 
             data=data, correlation=corARMA(form = ~1, p=1))

bic <- BIC(test$lme)
aic <- AIC(test$lme)

summary(test$gam)
```

The best formulae is:
total_arsenic ~ rainfall +  
               o2 + s(conductivity, bs='bs') + s(sulfate, bs = "bs") + feII + lead + lead^3 + zinc +
               rainfall:zinc + rainfall:lead + rainfall:feII + rainfall:o2
               
with correlation AR(1).

## From 2004 to 2019

```{r}
old_bic <- bic
old_aic <- aic

test <- gamm(total_arsenic ~ rainfall + rainfall_lag_1 + 
               o2 + conductivity + s(sulfate, bs = "bs") + feII + zinc +
               rainfall:conductivity + rainfall:o2 +
               rainfall_lag_1:feII + numeric_dates, 
             data=data_big, correlation=corARMA(form = ~1, p=1))


#AIC(test1, mod_alt)

bic <- BIC(test$lme)
aic <- AIC(test$lme)

summary(test$gam)
```

The best formulae is:
total_arsenic ~ rainfall + rainfall_lag_1 + 
               o2 + conductivity + s(sulfate, bs = "bs") + feII + zinc +
               rainfall:conductivity + rainfall:o2 +
               rainfall_lag_1:feII + numeric_dates

with correlation AR(1).

# Zinc
## From 2004 to 2010

Initial formulae:
zinc ~ rainfall + rainfall_lag_1 + 
               s(o2, bs='bs') + s(conductivity, bs='bs') + s(sulfate, bs = "bs") + feII + lead + lead^3 + total_arsenic +
               rainfall:total_arsenic + rainfall:lead + rainfall:feII + rainfall:sulfate + rainfall:conductivity + rainfall:o2 +
               rainfall_lag_1:total_arsenic + rainfall_lag_1:lead + rainfall_lag_1:feII + rainfall_lag_1:sulfate +
               rainfall_lag_1:conductivity + rainfall_lag_1:o2 + numeric_dates
            
```{r}
old_bic <- bic
old_aic <- aic

test <- gls(zinc ~ rainfall_lag_1 + 
               feII + lead + lead^3 + total_arsenic +
               rainfall_lag_1:total_arsenic, 
             data=data, correlation=corARMA(form = ~1, p=1))

bic <- BIC(test)
aic <- AIC(test)

summary(test)$tTable
```

The best formulae is:
zinc ~ rainfall_lag_1 + 
               feII + lead + lead^3 + total_arsenic +
               rainfall_lag_1:total_arsenic
               
with correlation AR(1).

## From 2004 to 2019
               
```{r}
old_bic <- bic
old_aic <- aic

test <- gamm(zinc ~ rainfall_lag_1 + 
               s(sulfate, bs = "bs") + feII + lead + lead^3 + total_arsenic +
               rainfall_lag_1:total_arsenic + numeric_dates, 
             data=data_big, correlation=corARMA(form = ~1, p=1))


#AIC(test1, mod_alt)

bic <- BIC(test$lme)
aic <- AIC(test$lme)

summary(test$gam)
```

The best formulae is:
zinc ~ rainfall_lag_1 + 
               s(sulfate, bs = "bs") + feII + lead + lead^3 + total_arsenic +
               rainfall_lag_1:total_arsenic + numeric_dates

with correlation AR(1).

# lead
## From 2004 to 2010

Initial formulae:
lead ~ rainfall + rainfall_lag_1 + 
               s(o2, bs='bs') + s(conductivity, bs='bs') + s(sulfate, bs = "bs") + feII + zinc + zinc^3 + total_arsenic +
               rainfall:total_arsenic + rainfall:zinc + rainfall:feII + rainfall:sulfate + rainfall:conductivity + rainfall:o2 +
               rainfall_lag_1:total_arsenic + rainfall_lag_1:zinc + rainfall_lag_1:feII + rainfall_lag_1:sulfate +
               rainfall_lag_1:conductivity + rainfall_lag_1:o2 + numeric_dates

```{r}
old_bic <- bic
old_aic <- aic

test <- gamm(lead ~ rainfall_lag_1 + 
               s(conductivity, bs='bs') + feII + zinc + zinc^3 + total_arsenic +
               rainfall_lag_1:total_arsenic,
             data=data, correlation=corARMA(form = ~1, p=1))

bic <- BIC(test$lme)
aic <- AIC(test$lme)
summary(test$gam)
```

```{r}
reg1 <- gamm(lead ~ rainfall_lag_1 + 
               s(conductivity, bs='bs') + feII + zinc + zinc^3 + total_arsenic +
               rainfall_lag_1:total_arsenic, 
             data=data, correlation=corARMA(form = ~1, p=1))
reg2 <- gamm(lead ~ rainfall_lag_1 + 
               s(conductivity, bs='bs') + feII + arsenic_zinc
               rainfall_lag_1:arsenic_zinc, 
             data=data, correlation=corARMA(form = ~1, p=1))

AIC(reg1$lme, reg2$lme)
BIC(reg1$lme, reg2$lme)
```

The best formulae is:
lead ~ rainfall_lag_1 + 
               s(conductivity, bs='bs') + feII + zinc + zinc^3 + total_arsenic +
               rainfall_lag_1:total_arsenic

with correlation AR(1).

## From 2004 to 2019

```{r}
old_bic <- bic
old_aic <- aic

test <- gamm(lead ~ rainfall_lag_1 + 
               o2 + s(conductivity, bs='bs') + s(sulfate, bs = "bs") + zinc + zinc^3 + 
               rainfall_lag_1:zinc + rainfall_lag_1:o2, 
             data=data_big, correlation=corARMA(form = ~1, p=1))

bic <- BIC(test$lme)
aic <- AIC(test$lme)
summary(test$gam)
```

The best formulae is:
lead ~ rainfall_lag_1 + 
               o2 + s(conductivity, bs='bs') + s(sulfate, bs = "bs") + zinc + zinc^3 + 
               rainfall_lag_1:zinc + rainfall_lag_1:o2

with correlation AR(1).

# feII
## From 2004 to 2010

Initial formulae:
feII ~ rainfall + rainfall_lag_1 + 
               s(o2, bs='bs') + s(conductivity, bs='bs') + s(sulfate, bs = "bs") + lead + lead^3 + zinc + total_arsenic +
               rainfall:total_arsenic + rainfall:zinc + rainfall:lead + rainfall:sulfate + rainfall:conductivity + rainfall:o2 +
               rainfall_lag_1:total_arsenic + rainfall_lag_1:zinc + rainfall_lag_1:lead + rainfall_lag_1:sulfate +
               rainfall_lag_1:conductivity + rainfall_lag_1:o2 + numeric_dates
               
```{r}
old_bic <- bic
old_aic <- aic

test <- gamm(feII ~ rainfall + rainfall_lag_1 + 
               s(sulfate, bs = "bs") + zinc + total_arsenic +
               rainfall:sulfate +
               rainfall_lag_1:total_arsenic + rainfall_lag_1:zinc + rainfall_lag_1:sulfate +
               numeric_dates, 
             data=data, correlation=corARMA(form = ~1, p=1))

bic <- BIC(test$lme)
aic <- AIC(test$lme)
summary(test$gam)
```

```{r}
reg1 <- gamm(feII ~ rainfall + rainfall_lag_1 + 
               s(sulfate, bs = "bs") + zinc + total_arsenic +
               rainfall:sulfate +
               rainfall_lag_1:total_arsenic + rainfall_lag_1:zinc + rainfall_lag_1:sulfate +
               numeric_dates, 
             data=data, correlation=corARMA(form = ~1, p=1))
reg2 <- gamm(feII ~ rainfall + rainfall_lag_1 + 
               conductivity + s(sulfate, bs = "bs") + arsenic_zinc +
               rainfall:sulfate + 
               rainfall_lag_1:arsenic_zinc + rainfall_lag_1:sulfate +
               numeric_dates, 
             data=data, correlation=corARMA(form = ~1, p=1))

AIC(reg1$lme, reg2$lme)
BIC(reg1$lme, reg2$lme)
```

```{r}
reg1 <- gamm(feII ~ rainfall + rainfall_lag_1 + 
               conductivity + s(sulfate, bs = "bs") + arsenic_zinc +
               rainfall:sulfate + 
               rainfall_lag_1:arsenic_zinc + rainfall_lag_1:sulfate +
               numeric_dates, 
             data=data, correlation=corARMA(form = ~1, p=1))
reg2 <- gamm(feII ~ rainfall + cond_rain_lag_1 + s(sulfate, bs = "bs") + arsenic_zinc +
               rainfall:sulfate + 
               cond_rain_lag_1:arsenic_zinc + cond_rain_lag_1:sulfate +
               numeric_dates, 
             data=data, correlation=corARMA(form = ~1, p=1))

AIC(reg1$lme, reg2$lme)
BIC(reg1$lme, reg2$lme)
```

The best formulae is:
feII ~ rainfall + rainfall_lag_1 + 
               conductivity + s(sulfate, bs = "bs") + arsenic_zinc +
               rainfall:sulfate + 
               rainfall_lag_1:arsenic_zinc + rainfall_lag_1:sulfate +
               numeric_dates

with correlation AR(1).

## From 2004 to 2019

```{r}
old_bic <- bic
old_aic <- aic

test <- gamm(feII ~ rainfall + rainfall_lag_1 + 
               conductivity + s(sulfate, bs = "bs") + zinc + total_arsenic +
               rainfall:sulfate + 
               rainfall_lag_1:total_arsenic + rainfall_lag_1:zinc + rainfall_lag_1:sulfate, 
             data=data_big, correlation=corARMA(form = ~1, p=1))

bic <- BIC(test$lme)
aic <- AIC(test$lme)
summary(test$gam)
```

The best formulae is:
feII ~ rainfall + rainfall_lag_1 + 
               conductivity + s(sulfate, bs = "bs") + zinc + total_arsenic +
               rainfall:sulfate + 
               rainfall_lag_1:total_arsenic + rainfall_lag_1:zinc + rainfall_lag_1:sulfate

with correlation AR(1).

