---
title: "Descriptive Statistics on Confluence with Generargues weather"
author: "CLETZ Laura"
date: "2025-06-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Packages needed:
```{r}
library(data.table)
library(dplyr)
library(tidyverse)
library(lubridate)
library(imputeTS)
here::i_am("R/Confluence_desc_stat.Rmd")
library(here)
```

# Confluence MD imputation

```{r, warning=FALSE}
conf <- read.table(here('processed','Confluence.csv'), dec = c(',','.') , sep = ";", header = TRUE) %>% 
  mutate_at(2:45, function(x) as.numeric(as.character(gsub(',','.', x))))
conf$Dates <- as.Date(conf$Dates, format="%Y-%m-%d")
```

```{r}
conf$Dates <- as.Date(substr(conf$Dates, 1, 10))
all_months <- seq(
  from = as.Date(format(min(conf$Dates), "%Y-%m-01")),
  to   = as.Date(format(max(conf$Dates), "%Y-%m-01")),
  by = "month"
)
all_months_df <- data.frame(Dates = all_months)

conf$MonthDate <- as.Date(format(conf$Dates, "%Y-%m-01"))
conf <- all_months_df %>%
  left_join(conf, by = c("Dates" = "MonthDate")) %>%
  select(-Dates.y)

conf <- conf %>% select(which(colSums(is.na(conf))<120))
```

```{r}
gen <- read.table(here('processed','Generargues_complete.csv'), dec = ',', sep = ";", header = TRUE) 
gen$Dates <- as.Date(gen$Dates, format="%Y-%m-%d")
gen$monthly_rainfall <- as.numeric(gen$monthly_rainfall)

conf <- gen %>%
  inner_join(conf, by="Dates")
```

```{r}
dis <- read.table(here('processed','S1_discharge.csv'), dec = c(',','.'), sep = ";", header = TRUE)
dis$Dates <- as.Date(dis$Dates, format="%Y-%m-%d")
dis$monthly_discharge <- as.numeric(dis$monthly_discharge)

dis2 <- read.table(here('processed','discharge_ESU.csv'), dec = c(',','.'), sep = ";", header = TRUE)
dis2$Dates <- as.Date(dis2$Dates, format="%Y-%m-%d")

dis3 <- dis %>%
  bind_rows(dis2)

conf <- conf %>%
  left_join(dis3) %>% 
  distinct(Dates, .keep_all = TRUE)
```

```{r}
conf_ts <- conf %>%
  mutate_at(2:18, function(x) ts(x, start=c(2004,11), end=c(2019,1), frequency = 12))
complete_conf <- conf_ts %>%
  na_kalman(model = "auto.arima")
```

```{r}
ggplot_na_imputations(conf_ts$Temperature, complete_conf$Temperature)
```

```{r}
write_delim(complete_conf, here("processed", "conf_ts.csv"), delim = ";")
```

# Monthly rainfall time series 

```{r}
data <- complete_conf
rain <- ts(data$monthly_rainfall, start = c(2004,11), end = c(2019,1), frequency = 12)
plot.ts(rain, xlab='Time')
```

```{r, echo=FALSE}
fit_rain <- decompose(rain)
plot(fit_rain)
```

```{r}
monthplot(rain, ylab = "Precipitations (in mm)")
```

```{r, echo=FALSE}
par(mfrow=c(1,2))
acf(rain, type="correlation", main='')
acf(rain, type="partial", main='')
```

# Temperature time series 

```{r}
temp <- ts(data$Temperature, start = c(2004,11), end = c(2019,1), frequency = 12)
plot.ts(temp, xlab='Time')
```

```{r, echo=FALSE}
fit_temp <- decompose(temp)
plot(fit_temp)
```

```{r}
monthplot(temp)
```

```{r, echo=FALSE}
par(mfrow=c(1,2))
acf(temp, type="correlation", main='')
acf(temp, type="partial", main='')
```

# pH time series 

```{r}
pH <- ts(data$pH, start = c(2004,11), end = c(2019,1), frequency = 12)
plot.ts(pH, xlab='Time')
```

```{r, echo=FALSE}
fit_pH <- decompose(pH)
plot(fit_pH)
```

```{r}
monthplot(pH)
```

```{r, echo=FALSE}
par(mfrow=c(1,2))
acf(pH, type="correlation", main='')
acf(pH, type="partial", main='')
```

# Eh time series 

```{r}
Eh <- ts(data$Eh, start = c(2004,11), end = c(2019,1), frequency = 12)
plot.ts(Eh, xlab='Time')
```

```{r, echo=FALSE}
fit_Eh <- decompose(Eh)
plot(fit_Eh)
```

```{r}
monthplot(Eh)
```

```{r, echo=FALSE}
par(mfrow=c(1,2))
acf(Eh, type="correlation", main='')
acf(Eh, type="partial", main='')
```

# O2 time series 

```{r}
O2 <- ts(data$O2, start = c(2004,11), end = c(2019,1), frequency = 12)
plot.ts(O2, xlab='Time')
```

```{r, echo=FALSE}
fit_O2 <- decompose(O2)
plot(fit_O2)
```

```{r}
monthplot(O2)
```

```{r, echo=FALSE}
par(mfrow=c(1,2))
acf(O2, type="correlation", main='')
acf(O2, type="partial", main='')
```

# Conductivity time series 

```{r}
cond <- ts(data$Conductivité, start = c(2004,11), end = c(2019,1), frequency = 12)
plot.ts(cond, xlab='Time')
```

```{r, echo=FALSE}
fit_cond <- decompose(cond)
plot(fit_cond)
```

```{r}
monthplot(cond)
```

```{r, echo=FALSE}
par(mfrow=c(1,2))
acf(cond, type="correlation", main='')
acf(cond, type="partial", main='')
```

# Lead time series 

```{r}
lead <- ts(data$Plomb, start = c(2004,11), end = c(2019,1), frequency = 12)
plot.ts(lead, xlab='Time')
```

```{r, echo=FALSE}
fit_lead <- decompose(lead)
plot(fit_lead)
```

```{r}
monthplot(lead)
```

```{r, echo=FALSE}
par(mfrow=c(1,2))
acf(lead, type="correlation", main='')
acf(lead, type="partial", main='')
```

# Aluminium time series 

```{r}
alu <- ts(data$Aluminium, start = c(2004,11), end = c(2019,1), frequency = 12)
plot.ts(alu, xlab='Time')
```

```{r, echo=FALSE}
fit_alu <- decompose(alu)
plot(fit_alu)
```

```{r}
monthplot(alu)
```

```{r, echo=FALSE}
par(mfrow=c(1,2))
acf(alu, type="correlation", main='')
acf(alu, type="partial", main='')
```

# Arséniate time series 

```{r}
arseniate <- ts(data$Arséniate, start = c(2004,11), end = c(2019,1), frequency = 12)
plot.ts(arseniate, xlab='Time')
```

```{r, echo=FALSE}
fit_arseniate <- decompose(arseniate)
plot(fit_arseniate)
```

```{r}
monthplot(arseniate)
```

```{r, echo=FALSE}
par(mfrow=c(1,2))
acf(arseniate, type="correlation", main='')
acf(arseniate, type="partial", main='')
```

# Arsénite time series 

```{r}
arsenite <- ts(data$Arsénite, start = c(2004,11), end = c(2019,1), frequency = 12)
plot.ts(arsenite, xlab='Time')
```

```{r, echo=FALSE}
fit_arsenite <- decompose(arsenite)
plot(fit_arsenite)
```

```{r}
monthplot(arsenite)
```

```{r, echo=FALSE}
par(mfrow=c(1,2))
acf(arsenite, type="correlation", main='')
acf(arsenite, type="partial", main='')
```

# Total Arsenic time series 

```{r}
arsenic <- ts(data$Arsenic_Total, start = c(2004,11), end = c(2019,1), frequency = 12)
plot.ts(arsenic, xlab='Time')
```

```{r, echo=FALSE}
fit_arsenic <- decompose(arsenic)
plot(fit_arsenic)
```

```{r}
monthplot(arsenic)
```

```{r, echo=FALSE}
par(mfrow=c(1,2))
acf(arsenic, type="correlation", main='')
acf(arsenic, type="partial", main='')
```

# Baryum time series 

```{r}
baryum <- ts(data$Baryum, start = c(2004,11), end = c(2019,1), frequency = 12)
plot.ts(baryum, xlab='Time')
```

```{r, echo=FALSE}
fit_baryum <- decompose(baryum)
plot(fit_baryum)
```

```{r}
monthplot(baryum)
```

```{r, echo=FALSE}
par(mfrow=c(1,2))
acf(baryum, type="correlation", main='')
acf(baryum, type="partial", main='')
```

# Sulfate time series 

```{r}
sulfate <- ts(data$Sulfate, start = c(2004,11), end = c(2019,1), frequency = 12)
plot.ts(sulfate, xlab='Time')
```

```{r, echo=FALSE}
fit_sulfate <- decompose(sulfate)
plot(fit_sulfate)
```

```{r}
monthplot(sulfate)
```

```{r, echo=FALSE}
par(mfrow=c(1,2))
acf(sulfate, type="correlation", main='')
acf(sulfate, type="partial", main='')
```

# Zinc time series 

```{r}
zinc <- ts(data$Zinc, start = c(2004,11), end = c(2019,1), frequency = 12)
plot.ts(zinc, xlab='Time')
```

```{r, echo=FALSE}
fit_zinc <- decompose(zinc)
plot(fit_zinc)
```

```{r}
monthplot(zinc)
```

```{r, echo=FALSE}
par(mfrow=c(1,2))
acf(zinc, type="correlation", main='')
acf(zinc, type="partial", main='')
```

# Cobalt time series 

```{r}
cobalt <- ts(data$Cobalt, start = c(2004,11), end = c(2019,1), frequency = 12)
plot.ts(cobalt, xlab='Time')
```

```{r, echo=FALSE}
fit_cobalt <- decompose(cobalt)
plot(fit_cobalt)
```

```{r}
monthplot(cobalt)
```

```{r, echo=FALSE}
par(mfrow=c(1,2))
acf(cobalt, type="correlation", main='')
acf(cobalt, type="partial", main='')
```

# Ferrous Ion time series 

```{r}
ion <- ts(data$Ion_Ferreux, start = c(2004,11), end = c(2019,1), frequency = 12)
plot.ts(ion, xlab='Time')
```

```{r, echo=FALSE}
fit_ion <- decompose(ion)
plot(fit_ion)
```

```{r}
monthplot(ion)
```

```{r, echo=FALSE}
par(mfrow=c(1,2))
acf(ion, type="correlation", main='')
acf(ion, type="partial", main='')
```

# Discharge time series 

```{r}
dis <- ts(data$monthly_discharge, start = c(2004,11), end = c(2019,1), frequency = 12)
plot.ts(dis, xlab='Time')
```

```{r, echo=FALSE}
fit_dis <- decompose(dis)
plot(fit_dis)
```

```{r}
monthplot(dis)
```

```{r, echo=FALSE}
par(mfrow=c(1,2))
acf(dis, type="correlation", main='')
acf(dis, type="partial", main='')
```

# General study

```{r}
fit <- data.frame(
  Dates = data$Dates,
  Aluminium = as.numeric(fit_alu$trend),
  Arséniate = as.numeric(fit_arseniate$trend),
  Arsénite = as.numeric(fit_arsenite$trend),
  Arsenic_Total = as.numeric(fit_arsenic$trend),
  Baryum = as.numeric(fit_baryum$trend),
  Cobalt = as.numeric(fit_cobalt$trend),
  Conductivité = as.numeric(fit_cond$trend),
  Débit = as.numeric(fit_dis$trend),
  Eh = as.numeric(fit_Eh$trend),
  Ion_Ferreux = as.numeric(fit_ion$trend),
  Plomb = as.numeric(fit_lead$trend),
  O2 = as.numeric(fit_O2$trend),
  pH = as.numeric(fit_pH$trend), 
  Pluie = as.numeric(fit_rain$trend),
  Sulfate = as.numeric(fit_sulfate$trend),
  Temperature = as.numeric(fit_temp$trend),
  Zinc = as.numeric(fit_zinc$trend)
)
```

```{r, warning=FALSE}
library(ggplot2)
library(scales)

fit_scaled <- fit %>%
  mutate_at(names(fit[2:18]),
            function(x) rescale(x))
```

```{r}
library(plotly)
p <- ggplot(fit_scaled, aes(x = Dates)) + 
  geom_line(aes(y = Aluminium, color = "Aluminium")) + 
  geom_line(aes(y = Arséniate, color = "As V")) + 
  geom_line(aes(y = Arsénite, color = "As III")) + 
  geom_line(aes(y = Arsenic_Total, color = "Arsenic Total")) + 
  geom_line(aes(y = Baryum, color = "Baryum")) + 
  geom_line(aes(y = Cobalt, color = "Cobalt")) + 
  geom_line(aes(y = Conductivité, color = "Conductivité")) + 
  geom_line(aes(y = Débit, color = "Débit")) + 
  geom_line(aes(y = Eh, color = "Eh")) + 
  geom_line(aes(y = Ion_Ferreux, color = "FeII")) + 
  geom_line(aes(y = O2, color = "O2")) + 
  geom_line(aes(y = pH, color = "pH")) + 
  geom_line(aes(y = Pluie, color = "Pluie")) + 
  geom_line(aes(y = Plomb, color = "Plomb")) +
  geom_line(aes(y = Sulfate, color = "Sulfate")) + 
  geom_line(aes(y = Temperature, color = "Température")) + 
  geom_line(aes(y = Zinc, color = "Zinc")) + 
  scale_x_date(breaks = date_breaks("years"), labels = date_format("%Y")) +
  labs(y = "Scaled Values", color = "Variables") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

g <- ggplotly(p) %>% 
  layout(hovermode = "x unified")
g
```

```{r}
library(htmlwidgets)
saveWidget(g, here("Réunions","conf_interactive_plot.html"), selfcontained = TRUE)
```

