---
title: "Descriptive Statistics on Source S1 with Anduze meteo"
author: "CLETZ Laura"
date: "11/06/2025"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Packages needed:
```{r}
library(data.table)
library(dplyr)
library(tidyverse)
library(lubridate)
library(imputeTS)
here::i_am("R/new_md_imputation.Rmd")
library(here)
```

```{r}
data <- read.table(here("processed","S1_ts.csv"), dec = c(',','.') , sep = ";", header = TRUE) %>% 
  mutate_at(2:26, function(x) as.numeric(as.character(x)))
data$Dates <- as.Date(data$Dates, format="%Y-%m-%d")
```

# Monthly rainfall time series 

```{r}
rain <- ts(data$monthly_rainfall, start = c(2004,11), end = c(2018,6), frequency = 12)
plot.ts(rain, xlab='Time', ylab="Precipitations (in mm)")
# mm = L/m^2
```

```{r, echo=FALSE}
fit_rain <- decompose(rain)
plot(fit_rain)
```

```{r}
monthplot(rain, ylab = "Precipitations (in mm)")
```

```{r, echo=FALSE}
par(mfrow=c(1,2))
acf(rain, type="correlation", main='')
acf(rain, type="partial", main='')
```

# Temperature time series 

```{r}
temp <- ts(data$Temperature, start = c(2004,11), end = c(2018,6), frequency = 12)
plot.ts(temp, xlab='Time')
```

```{r, echo=FALSE}
fit_temp <- decompose(temp)
plot(fit_temp)
```

```{r}
monthplot(temp)
```

```{r, echo=FALSE}
par(mfrow=c(1,2))
acf(temp, type="correlation", main='')
acf(temp, type="partial", main='')
```

# pH time series 

```{r}
pH <- ts(data$pH, start = c(2004,11), end = c(2018,6), frequency = 12)
plot.ts(pH, xlab='Time')
```

```{r, echo=FALSE}
fit_pH <- decompose(pH)
plot(fit_pH)
```

```{r}
monthplot(pH)
```

```{r, echo=FALSE}
par(mfrow=c(1,2))
acf(pH, type="correlation", main='')
acf(pH, type="partial", main='')
```

# Eh time series 

```{r}
Eh <- ts(data$Eh, start = c(2004,11), end = c(2018,6), frequency = 12)
plot.ts(Eh, xlab='Time')
```

```{r, echo=FALSE}
fit_Eh <- decompose(Eh)
plot(fit_Eh)
```

```{r}
monthplot(Eh)
```

```{r, echo=FALSE}
par(mfrow=c(1,2))
acf(Eh, type="correlation", main='')
acf(Eh, type="partial", main='')
```

# O2 time series 

```{r}
O2 <- ts(data$O2, start = c(2004,11), end = c(2018,6), frequency = 12)
plot.ts(O2, xlab='Time')
```

```{r, echo=FALSE}
fit_O2 <- decompose(O2)
plot(fit_O2)
```

```{r}
monthplot(O2)
```

```{r, echo=FALSE}
par(mfrow=c(1,2))
acf(O2, type="correlation", main='')
acf(O2, type="partial", main='')
```

# Conductivity time series 

```{r}
cond <- ts(data$Conductivité, start = c(2004,11), end = c(2018,6), frequency = 12)
plot.ts(cond, xlab='Time')
```

```{r, echo=FALSE}
fit_cond <- decompose(cond)
plot(fit_cond)
```

```{r}
monthplot(cond)
```

```{r, echo=FALSE}
par(mfrow=c(1,2))
acf(cond, type="correlation", main='')
acf(cond, type="partial", main='')
```

# Lead time series 

```{r}
lead <- ts(data$Plomb, start = c(2004,11), end = c(2018,6), frequency = 12)
plot.ts(lead, xlab='Time')
```

```{r, echo=FALSE}
fit_lead <- decompose(lead)
plot(fit_lead)
```

```{r}
monthplot(lead)
```

```{r, echo=FALSE}
par(mfrow=c(1,2))
acf(lead, type="correlation", main='')
acf(lead, type="partial", main='')
```

# Aluminium time series 

```{r}
alu <- ts(data$Aluminium, start = c(2004,11), end = c(2018,6), frequency = 12)
plot.ts(alu, xlab='Time')
```

```{r, echo=FALSE}
fit_alu <- decompose(alu)
plot(fit_alu)
```

```{r}
monthplot(alu)
```

```{r, echo=FALSE}
par(mfrow=c(1,2))
acf(alu, type="correlation", main='')
acf(alu, type="partial", main='')
```

# Antimoine time series 

```{r}
antimoine <- ts(data$Antimoine, start = c(2004,11), end = c(2018,6), frequency = 12)
plot.ts(antimoine, xlab='Time')
```

```{r, echo=FALSE}
fit_antimoine <- decompose(antimoine)
plot(fit_antimoine)
```

```{r}
monthplot(antimoine)
```

```{r, echo=FALSE}
par(mfrow=c(1,2))
acf(antimoine, type="correlation", main='')
acf(antimoine, type="partial", main='')
```

# Arséniate time series 

```{r}
arseniate <- ts(data$Arséniate, start = c(2004,11), end = c(2018,6), frequency = 12)
plot.ts(arseniate, xlab='Time')
```

```{r, echo=FALSE}
fit_arseniate <- decompose(arseniate)
plot(fit_arseniate)
```

```{r}
monthplot(arseniate)
```

```{r, echo=FALSE}
par(mfrow=c(1,2))
acf(arseniate, type="correlation", main='')
acf(arseniate, type="partial", main='')
```

# Arsénite time series 

```{r}
arsenite <- ts(data$Arsénite, start = c(2004,11), end = c(2018,6), frequency = 12)
plot.ts(arsenite, xlab='Time')
```

```{r, echo=FALSE}
fit_arsenite <- decompose(arsenite)
plot(fit_arsenite)
```

```{r}
monthplot(arsenite)
```

```{r, echo=FALSE}
par(mfrow=c(1,2))
acf(arsenite, type="correlation", main='')
acf(arsenite, type="partial", main='')
```

# Total Arsenic time series 

```{r}
arsenic <- ts(data$Arsenic_Total, start = c(2004,11), end = c(2018,6), frequency = 12)
plot.ts(arsenic, xlab='Time')
```

```{r, echo=FALSE}
fit_arsenic <- decompose(arsenic)
plot(fit_arsenic)
```

```{r}
monthplot(arsenic)
```

```{r, echo=FALSE}
par(mfrow=c(1,2))
acf(arsenic, type="correlation", main='')
acf(arsenic, type="partial", main='')
```

# Baryum time series 

```{r}
baryum <- ts(data$Baryum, start = c(2004,11), end = c(2018,6), frequency = 12)
plot.ts(baryum, xlab='Time')
```

```{r, echo=FALSE}
fit_baryum <- decompose(baryum)
plot(fit_baryum)
```

```{r}
monthplot(baryum)
```

```{r, echo=FALSE}
par(mfrow=c(1,2))
acf(baryum, type="correlation", main='')
acf(baryum, type="partial", main='')
```

# Strontium time series 

```{r}
strontium <- ts(data$Strontium, start = c(2004,11), end = c(2018,6), frequency = 12)
plot.ts(strontium, xlab='Time')
```

```{r, echo=FALSE}
fit_strontium <- decompose(strontium)
plot(fit_strontium)
```

```{r}
monthplot(strontium)
```

```{r, echo=FALSE}
par(mfrow=c(1,2))
acf(strontium, type="correlation", main='')
acf(strontium, type="partial", main='')
```

# Sulfate time series 

```{r}
sulfate <- ts(data$Sulfate, start = c(2004,11), end = c(2018,6), frequency = 12)
plot.ts(sulfate, xlab='Time')
```

```{r, echo=FALSE}
fit_sulfate <- decompose(sulfate)
plot(fit_sulfate)
```

```{r}
monthplot(sulfate)
```

```{r, echo=FALSE}
par(mfrow=c(1,2))
acf(sulfate, type="correlation", main='')
acf(sulfate, type="partial", main='')
```

# Thallium time series 

```{r}
thallium <- ts(data$Thallium, start = c(2004,11), end = c(2018,6), frequency = 12)
plot.ts(thallium, xlab='Time')
```

```{r, echo=FALSE}
fit_thallium <- decompose(thallium)
plot(fit_thallium)
```

```{r}
monthplot(thallium)
```

```{r, echo=FALSE}
par(mfrow=c(1,2))
acf(thallium, type="correlation", main='')
acf(thallium, type="partial", main='')
```

# Zinc time series 

```{r}
zinc <- ts(data$Zinc, start = c(2004,11), end = c(2018,6), frequency = 12)
plot.ts(zinc, xlab='Time')
```

```{r, echo=FALSE}
fit_zinc <- decompose(zinc)
plot(fit_zinc)
```

```{r}
monthplot(zinc)
```

```{r, echo=FALSE}
par(mfrow=c(1,2))
acf(zinc, type="correlation", main='')
acf(zinc, type="partial", main='')
```

# Cadmium time series 

```{r}
cadmium <- ts(data$Cadmium, start = c(2004,11), end = c(2018,6), frequency = 12)
plot.ts(cadmium, xlab='Time')
```

```{r, echo=FALSE}
fit_cadmium <- decompose(cadmium)
plot(fit_cadmium)
```

```{r}
monthplot(cadmium)
```

```{r, echo=FALSE}
par(mfrow=c(1,2))
acf(cadmium, type="correlation", main='')
acf(cadmium, type="partial", main='')
```

# Chrome time series 

```{r}
chrome <- ts(data$Chrome, start = c(2004,11), end = c(2018,6), frequency = 12)
plot.ts(chrome, xlab='Time')
```

```{r, echo=FALSE}
fit_chrome <- decompose(chrome)
plot(fit_chrome)
```

```{r}
monthplot(chrome)
```

```{r, echo=FALSE}
par(mfrow=c(1,2))
acf(chrome, type="correlation", main='')
acf(chrome, type="partial", main='')
```

# Fer_Total time series 

```{r}
iron <- ts(data$Fer_Total, start = c(2004,11), end = c(2018,6), frequency = 12)
plot.ts(iron, xlab='Time')
```

```{r, echo=FALSE}
fit_iron <- decompose(iron)
plot(fit_iron)
```

```{r}
monthplot(iron)
```

```{r, echo=FALSE}
par(mfrow=c(1,2))
acf(iron, type="correlation", main='')
acf(iron, type="partial", main='')
```

# Rubidium time series 

```{r}
rub <- ts(data$Rubidium, start = c(2004,11), end = c(2018,6), frequency = 12)
plot.ts(rub, xlab='Time')
```

```{r, echo=FALSE}
fit_rub <- decompose(rub)
plot(fit_rub)
```

```{r}
monthplot(rub)
```

```{r, echo=FALSE}
par(mfrow=c(1,2))
acf(rub, type="correlation", main='')
acf(rub, type="partial", main='')
```

# Cobalt time series 

```{r}
cobalt <- ts(data$Cobalt, start = c(2004,11), end = c(2018,6), frequency = 12)
plot.ts(cobalt, xlab='Time')
```

```{r, echo=FALSE}
fit_cobalt <- decompose(cobalt)
plot(fit_cobalt)
```

```{r}
monthplot(cobalt)
```

```{r, echo=FALSE}
par(mfrow=c(1,2))
acf(cobalt, type="correlation", main='')
acf(cobalt, type="partial", main='')
```

# Copper time series 

```{r}
copper <- ts(data$Cuivre, start = c(2004,11), end = c(2018,6), frequency = 12)
plot.ts(copper, xlab='Time')
```

```{r, echo=FALSE}
fit_copper <- decompose(copper)
plot(fit_copper)
```

```{r}
monthplot(copper)
```

```{r, echo=FALSE}
par(mfrow=c(1,2))
acf(copper, type="correlation", main='')
acf(copper, type="partial", main='')
```

# Nickel time series 

```{r}
nickel <- ts(data$Nickel, start = c(2004,11), end = c(2018,6), frequency = 12)
plot.ts(nickel, xlab='Time')
```

```{r, echo=FALSE}
fit_nickel <- decompose(nickel)
plot(fit_nickel)
```

```{r}
monthplot(nickel)
```

```{r, echo=FALSE}
par(mfrow=c(1,2))
acf(nickel, type="correlation", main='')
acf(nickel, type="partial", main='')
```

# Ferrous Ion time series 

```{r}
ion <- ts(data$Ion_Ferreux, start = c(2004,11), end = c(2018,6), frequency = 12)
plot.ts(ion, xlab='Time')
```

```{r, echo=FALSE}
fit_ion <- decompose(ion)
plot(fit_ion)
```

```{r}
monthplot(ion)
```

```{r, echo=FALSE}
par(mfrow=c(1,2))
acf(ion, type="correlation", main='')
acf(ion, type="partial", main='')
```

# General study

```{r}
fit <- data.frame(
  Dates = data$Dates,
  Aluminium = as.numeric(fit_alu$trend),
  Antimoine = as.numeric(fit_antimoine$trend),
  Arséniate = as.numeric(fit_arseniate$trend),
  Arsénite = as.numeric(fit_arsenite$trend),
  Arsenic_Total = as.numeric(fit_arsenic$trend),
  Baryum = as.numeric(fit_baryum$trend),
  Cadmium = as.numeric(fit_cadmium$trend),
  Chrome = as.numeric(fit_chrome$trend),
  Cobalt = as.numeric(fit_cobalt$trend),
  Conductivité = as.numeric(fit_cond$trend),
  Cuivre = as.numeric(fit_copper$trend),
  Eh = as.numeric(fit_Eh$trend),
  Ion_Ferreux = as.numeric(fit_ion$trend),
  Fer_Total = as.numeric(fit_iron$trend),
  Plomb = as.numeric(fit_lead$trend),
  Nickel = as.numeric(fit_nickel$trend),
  O2 = as.numeric(fit_O2$trend),
  pH = as.numeric(fit_pH$trend), 
  Pluie = as.numeric(fit_rain$trend),
  Rubidium = as.numeric(fit_rub$trend),
  Strontium = as.numeric(fit_strontium$trend),
  Sulfate = as.numeric(fit_sulfate$trend),
  Temperature = as.numeric(fit_temp$trend),
  Thallium = as.numeric(fit_thallium$trend),
  Zinc = as.numeric(fit_zinc$trend)
)
```

```{r, warning=FALSE}
library(ggplot2)
library(scales)

fit_scaled <- fit %>%
  mutate_at(names(fit[2:26]),
            function(x) rescale(x))
```

```{r}
library(plotly)
p <- ggplot(fit_scaled, aes(x = Dates)) + 
  geom_line(aes(y = Aluminium, color = "Aluminium")) + 
  geom_line(aes(y = Antimoine, color = "Antimoine")) + 
  geom_line(aes(y = Arséniate, color = "As V")) + 
  geom_line(aes(y = Arsénite, color = "As III")) + 
  geom_line(aes(y = Arsenic_Total, color = "Arsenic Total")) + 
  geom_line(aes(y = Baryum, color = "Baryum")) + 
  geom_line(aes(y = Cadmium, color = "Cadmium")) + 
  geom_line(aes(y = Chrome, color = "Chrome")) + 
  geom_line(aes(y = Cobalt, color = "Cobalt")) + 
  geom_line(aes(y = Conductivité, color = "Conductivité")) + 
  geom_line(aes(y = Cuivre, color = "Cuivre")) + 
  geom_line(aes(y = Eh, color = "Eh")) + 
  geom_line(aes(y = Ion_Ferreux, color = "FeII")) + 
  geom_line(aes(y = Fer_Total, color = "Fer Total")) +
  geom_line(aes(y = Nickel, color = "Nickel")) + 
  geom_line(aes(y = O2, color = "O2")) + 
  geom_line(aes(y = pH, color = "pH")) + 
  geom_line(aes(y = Pluie, color = "Pluie")) + 
  geom_line(aes(y = Rubidium, color = "Rubidium")) + 
  geom_line(aes(y = Strontium, color = "Strontium")) + 
  geom_line(aes(y = Plomb, color = "Plomb")) +
  geom_line(aes(y = Sulfate, color = "Sulfate")) + 
  geom_line(aes(y = Temperature, color = "Température")) + 
  geom_line(aes(y = Thallium, color = "Thallium")) + 
  geom_line(aes(y = Zinc, color = "Zinc")) + 
  scale_x_date(breaks = date_breaks("years"), labels = date_format("%Y")) +
  labs(y = "Scaled Values", color = "Variables") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

g <- ggplotly(p) %>% 
  layout(hovermode = "x unified")
g
```

```{r}
library(htmlwidgets)
saveWidget(g, here("Réunions","S1_interactive_plot.html"), selfcontained = TRUE)
```

